### 2018-02-05 ms/genetic_cultural_fitness.tex
protected polymorphism in constant environment

\documentclass[12pt]{article} % Default font size of the document, change to 10pt to fit more text

\usepackage{geometry}
\geometry{
a4paper,
total={170mm,257mm},
left=20mm,
top=20mm,
headheight=12pt
}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=CadetBlue,citecolor=black,linkcolor=black]{hyperref}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
%\usepackage{nameref} % included in hyperref
\usepackage{subcaption}

\usepackage{titlesec}
% \titleformat{command}[shape]{format}{label}{sep}{before}[after]
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}[]
\titleformat{\subsection}[runin]{\normalfont\large\bfseries}{\thesubsection}{1em}{}[]
\titleformat{\paragraph}[runin]{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}[]

% left margin	| space before (vertical) | space after (horizontal)
\titlespacing{\section}{0em}{0\baselineskip}{0\baselineskip}
\titlespacing{\subsection}{0em}{0\baselineskip}{1em}
\titlespacing{\paragraph}{0em}{0\baselineskip}{1em}

%SetFonts
% newtxtext+newtxmath
\usepackage{newtxtext} %loads helv for ss, txtt for tt
\usepackage{amsmath}
\usepackage[bigdelims]{newtxmath}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
%SetFonts

% math commands
\newcommand*{\tr}{^\intercal}
\let\vec\mathbf
\newcommand{\matrx}[1]{{\left[ \stackrel{}{#1}\right]}}
\newcommand{\diag}[1]{\mbox{diag}\matrx{#1}}
\newcommand{\goesto}{\rightarrow}
\newcommand{\dspfrac}[2]{\frac{\displaystyle #1}{\displaystyle #2} }
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{result}{Result}
\renewcommand\qedsymbol{} % no square at end of proof
\newcommand{\cl}{\mathbf{L}}
\newcommand{\cj}{\mathbf{J}}
\newcommand{\ci}{I}
\newcommand{\E}{\mathbf{E}}

% NatBib
\usepackage[round,colon,authoryear]{natbib}

% Title page
\title{Oblique Transmission with Genetic and Cultural Fitnesses}

% authors http://ftp.cc.uoc.gr/mirrors/CTAN/macros/latex/contrib/preprint/authblk.pdf
\usepackage[blocks,auth-lg,affil-it]{authblk}
\author{Yoav Ram}
\author{Marcus W. Feldman}
\affil{Department of Biology, Stanford University, Stanford, CA}

\date{\today}

% Document
\begin{document}
\maketitle

\section*{Model}

Consider a population with one trait and two trait values $A$ and $B$, with frequencies $x$ and $1-x$.
Phenotypes are transmitted either by learning from a parent with probability $\rho$ or by learning from a random adult model (e.g. teacher) with probability $1-\rho$.
The two phenotypes have distinct genetic ($w_A$ and $w_B$) and cultural ($v_A$ and $v_B$) fitness values~\citep{ElMouden2014}, defined by relative reproduction success and cultural influence, respectively.
\begin{equation} \label{eq:table}
\begin{array}{ccccc}
\hbox{ phenotype}&A&B\\
\hbox{ frequency}&x&1-x\\
\hbox{ genetic fitness}&w_A&w_B\\
\hbox{ cultural fitness}&v_A&v_B\\
\end{array}
\end{equation}

Given the frequency $x$ of phenotype $A$ in the current generation, the frequency in the next generation $x'$ is given by
\begin{equation} \label{eq:recurrence}
x' = \rho x \frac{w_A}{\bar w} + (1-\rho) x \frac{v_A}{\bar v} 
\end{equation}
where $\bar w = x w_A + (1-x) w_B$ is the population mean genetic fitness and $\bar v = x v_A + (1-x) v_B$ is the population mean cultural fitness. 
Note that if $v_A = v_B$ then \eqref{eq:recurrence} simplifies to 
$x' = \rho x \frac{w_A}{\bar w} + (1-\rho) x$,
which is eq.~2 from~\citet{Ram2018}.
So setting $v_A = v_B = \rho$ or even to so other parameter $v_A = v_B = \tau$ will not provide new insight.

\section*{Constant environment}

In a constant environment we will assume, without loss of generality, that $w_A > w_B$.
We will now find the equilibrium phenotype frequencies.
Setting $x'=x$ in~\eqref{eq:recurrence} and rearranging, we get a cubic polynomial $ax^3+bx^2+cx$.
%\begin{equation}
%\begin{aligned}
%G(x) &= a x^3 + b x^2 +c x, \quad \text{where} \\
%a &= (w_A - w_B) (v_A - v_B) \\ 
%b &= \rho(w_A v_B - w_B v_A) - (w_A - 2 w_B)(v_A - v_B) \\
%c &= -w_B(v_A - v_B) - \rho (w_A v_B - w_B v_A).  
%\end{aligned}
%\end{equation}
Two roots are easy to verify: $x=0 \Rightarrow x'=0$ and $x=1 \Rightarrow x'=\rho \frac{w_A}{w_A} + (1-\rho) \frac{v_A}{v_A} = 1$. 
Therefore, the cubic polynomial is $x (1-x) (x^*-x)$ where
\begin{equation}
%\begin{aligned}
x^* = 
%\frac{w_B v_B - (1-\rho) w_B v_A - \rho w_A v_B}{w_A v_A - w_A v_B - w_B v_A + w_B v_B} = \\
\rho \frac{w_B v_A - w_A v_B}{(w_A - w_B)(v_A - v_B)} - \frac{w_B}{w_A-w_B}.
%\end{aligned}
\end{equation}
%The conditions for $0 < x^* < 1$ are:
%\begin{equation}
%\begin{aligned}
%\frac{w_B (v_A - v_B)}{w_B v_A - w_A v_B}  < 
%\rho < 
%\frac{w_A (v_A - v_B)}{w_B v_A - w_A v_B} \\
%\end{aligned}
%\end{equation}

Now, $x=0$ is unstable if
\begin{equation}
%\frac{x'}{x} \eqsim 
\rho \frac{w_A}{w_B} + (1-\rho) \frac{v_A}{v_B} > 1,
\end{equation}
and $x=1$ is unstable if
\begin{equation}
%\frac{(1-x)'}{1-x} \eqsim 
\rho \frac{w_B}{w_A} + (1-\rho) \frac{v_B}{v_A} > 1.
\end{equation}

So both $x=0$ and $x=1$ are unstable if
\begin{equation} \label{eq:protected_poly_const}
w_B(v_B - v_A) < \rho (w_A v_B - w_B v_A) < w_A(v_B - v_A),
\end{equation}
in which case there exists a protected phenotypic polymorphism $0 < x^* < 1$.

There are three possibilities (disregarding equalities):
\begin{enumerate}
\item If $v_A > v_B$ and $\frac{w_A}{w_B} > \frac{v_A}{v_B}$ then condition~\eqref{eq:protected_poly_const} cannot apply as it requires $\rho<0$.
\item If $v_A > v_B$ and $\frac{w_A}{w_B} < \frac{v_A}{v_B}$ then condition~\eqref{eq:protected_poly_const} cannot apply as it requires $\rho>1$.
\item If $v_B > v_A$ then $w_A v_B - w_B v_A > 0$, $w_B(v_B - v_A)<w_A v_B - w_B v_A$, and $w_A v_B - w_B v_A > w_A(v_B - v_A)$ and condition~\eqref{eq:protected_poly_const} applies, ensuring a protected phenotypic polymorphism, if
\begin{equation}
\frac{w_B(v_B - v_A)}{w_A v_B - w_B v_A} < \rho < \frac{w_A(v_B - v_A)}{w_A v_B - w_B v_A},
\end{equation}
which can also be written as 
\begin{equation}
\frac{-s_v}{s_w-s_v} 
< \rho < 
(1+s_w) \cdot \frac{-s_v}{s_w-s_v},
\end{equation}
where $s_w=\frac{w_A}{w_B}-1>0$ and $s_v=\frac{v_A}{v_B}-1<0$.
\end{enumerate}

%\section*{Acknowledgements.}

%%%%%%%%%%%%%%%%%%%
\bibliographystyle{unsrtnat}
%\bibliography{/Users/yoavram/Documents/library}
\bibliography{genetic_cultural_fitness}
%%%%%%%%%%%%%%%%%%%
\end{document}

### 2018-03-13 ms/fluctuating_oblique.tex
first results of fluc trans

\documentclass[12pt]{extarticle} %twocolumn
\usepackage{geometry}
\geometry{
a4paper,
total={170mm,257mm},
left=20mm,
top=20mm,
headheight=12pt
}

%\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx} % Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
% TeX will automatically convert eps --> pdf in pdflatex

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=CadetBlue,citecolor=black,linkcolor=black]{hyperref}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[]{authblk}

%SetFonts
% newtxtext+newtxmath
\usepackage{newtxtext} %loads helv for ss, txtt for tt
\usepackage{amsmath}
\usepackage[bigdelims]{newtxmath}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
%SetFonts

% less space before sections 
% \@startsection {NAME}{LEVEL}{INDENT}{BEFORESKIP}{AFTERSKIP}{STYLE} 
%            optional * [ALTHEADING]{HEADING} 
\makeatletter
 \renewcommand\section{\@startsection {section}{1}{\z@}%
     {-2.5ex \@plus -1ex \@minus -.2ex}%
     {1.3ex \@plus.2ex}%
    {\Large\bfseries}}

% Yoav & Lee commands
\newcommand*{\tr}{^\intercal}
\let\vec\mathbf
\newcommand{\matrx}[1]{{\left[ \stackrel{}{#1}\right]}}
\newcommand{\diag}[1]{\mbox{diag}\matrx{#1}}
\newcommand{\goesto}{\rightarrow}
\newcommand{\dspfrac}[2]{\frac{\displaystyle #1}{\displaystyle #2} }
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{result}{Result}
\renewcommand\qedsymbol{} % no square at end of proof
\newcommand{\cl}{\mathbf{L}}
\newcommand{\cj}{\mathbf{J}}
\newcommand{\ci}{I}

% NatBib
\usepackage[round,colon,authoryear]{natbib}

% Title page
\title{Vertical and oblique transmission with fluctuating transmission}

\author[a]{Yoav Ram}
\author[b]{Uri Liberman}
\author[a]{Marcus W. Feldman}
\affil[a]{Department of Biology, Stanford University, Stanford, CA}
\affil[b]{School of Mathematical Sciences, Tel Aviv University, Israel}

\date{\today}

% Document
\begin{document}
\maketitle

% Abstract
%\begin{abstract}
%\end{abstract}

% Introduction
%\section*{Introduction}

% Model
\section*{Model}

Consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ and fitness values $1+s$ and $1$, respectively.
Phenotypes are transmitted vertically with probability $\rho$ or diagonally with probability $1-\rho$.
Given $x$ the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is
\begin{equation} \label{eq:recurrence}
x' = \rho \frac{1+s}{\overline w} x + (1-\rho)x,
\end{equation}
where $\overline w = 1 + xs$ is the population mean fitness.

% Periodic fluctuating transmission
\subsection*{Periodic fluctuating transmission}

Suppose that the vertical transmission rate $\rho$ is fluctuating, such that $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
Following \eqref{eq:recurrence}, the recurrence equations for two generations is then
\begin{equation}\begin{aligned} \label{eq:recurrence_two_generations}
x' = \rho_1 \frac{1+s}{\overline w} x + (1-\rho_1)x, \\
x'' = \rho_2 \frac{1+s}{\overline w} x' + (1-\rho_2)x'.
\end{aligned}\end{equation}

To find the equilibrium $x^*$, that is, solutions of $x''=x$, we note that both fixations (i.e. $x=0$ and $x=1$) are equilibria. 
Therefore, $x^*$ is a root of $G(x) = \frac{x''-x}{x(1-x)} \gamma$, where
\begin{align}
G(x) = ax^2+bx+c, \\
\gamma = \frac{1+xs}{s} \Big( 1 + xs(2+s\rho_1) + x^2 s^2 (1-\rho_1) \Big) > 0, \\
a = s^2 (1-\rho_1) (\rho_1 + \rho_2(1-\rho_1)) > 0, \\
b = s(2+s\rho_1)(\rho_1 + \rho_2(1-\rho_1)) > 0, \\
c = \rho_1 + \rho_1 \rho_2 s + \rho_2 > 0.
\end{align} 
Note that 
\begin{align}
G(0) = c > 0,\\ 
G'(0) = b > 0,\\
G(1)=a+b+c>c>0,\\
G'(1)=2a+b>b>0,
\end{align}
and therefore $G(x)=0$ is not attained for any $x \in [0,1]$, and there is no steady polymorphism in \eqref{eq:recurrence_two_generations}.
Also, the local stability of $x=0$ is determined by $(1+s\rho_1)(1+s\rho_2)$.
Therefore, when $A$ is favored and $s>0$, $x=0$ is unstable and $x=1$ is stable, whereas the opposite applies when $B$ is favored, so that fixation of the favored phenotype is locally stable.

Overall, we find that fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype without any protected polymorphisms.

% Periodic fluctuating transmission and selection
\subsection*{Periodic fluctuating transmission and selection}

Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate.
Following \eqref{eq:recurrence}, the recurrence equations for two generations is then
\begin{equation}\begin{aligned} \label{eq:recurrence_two_generations_fluc_s}
x' = \rho_1 \frac{1+s}{1+sx} x + (1-\rho_1)x, \\
x'' = \rho_2 \frac{1}{1+s-sx} x' + (1-\rho_2)x'.
\end{aligned}\end{equation}

To find the equilibrium $x^*$, that is, solutions of $x''=x$, we note that both fixations (i.e. $x=0$ and $x=1$) are equilibria. 
Therefore, $x^*$ is a root of $G(x) = \frac{x''-x}{x(1-x)} \gamma$, where
\begin{align} \label{eq:Gx_fluctuating_selection_transmission_k1}
G(x) = ax^2+bx+c, \\
\gamma = -\frac{1+xs}{s} \Big( 1 + s + s^2x (1-\rho) +s^2x^2(1-\rho) \Big) > 0, \\
a = s^2 \rho_1 (1-\rho_1) (1 - \rho_2 + \rho_2/\rho_1)) > 0, \\
b = s(1-\rho_1)(2\rho_2 - s\rho_1(1-\rho_2)), \\
c = \rho_2 - \rho_1 - s\rho_1(1-\rho_2).
\end{align}

Note that 
\begin{align}
G(0) = c > 0  \Leftrightarrow \rho_2 > \frac{\rho_1(1+s)}{1+\rho_1s}, \\
G'(0) = b > 0 \Leftrightarrow \rho_2 > \frac{\rho_1 s}{2 + \rho_1s}, \\
G(1) = a + b + c = (1+s)[\rho_2 - \rho_1 + s \rho_2 (1-\rho_1)], \\
G'(1) = 2a + b = s(1-\rho_1)(2(1+s)\rho_2 + s\rho_1(1-\rho_2)) > 0.
\end{align} 
From analysis of the quadratic $G(x)$ we can then determine that it has a root in $(0, 1)$, and that a protected polymorphism exists $x^*$ (i.e. $G(x^*)=0$ and $0<x^*<1$), if and only if 
\begin{equation}
-s\rho_2(1-\rho_1) < \rho_2 - \rho_1 < s\rho_1(1-\rho_2),
\end{equation}
Therefore, a protected polymorphism exists only if the difference in the transmission rate between consecutive generation is not too large, and, surprisingly, the stronger selection is, the larger the difference can be that still allows a protected polymorphism.
Equivalently, a protected polymorphism exists if
\begin{equation} \label{eq:rho2_fluctuating_selection_transmission_k1}
\frac{\rho_1}{1+s(1-\rho_1)} < \rho_2 < \frac{(1+s)\rho_1}{1+s\rho_1}.
\end{equation}

\begin{figure*}[ht]
\centering
\includegraphics[width=\linewidth]{../figures/{rho1_rho2_k=1}.pdf}
\caption{
\textbf{Protected polymorphism with fluctuating transmission and selection.}
The positive root $x^*$ of $G(x)$, see eq.~\ref{eq:Gx_fluctuating_selection_transmission_k1}.
Dashed lines represent, from left to right, $\rho_2=\frac{\rho_1}{1+s(1-\rho_1)}$, $\rho_2=\rho_1$, and $\rho_2=\frac{(1+s)\rho_1}{1+s\rho_1}$, see eq.~\ref{eq:rho2_fluctuating_selection_transmission_k1}.
Here, $\rho_1=0.5$, $s=0.1$.}
\label{fig:rho1_rho2_k=1}
\end{figure*}

% Randomly fluctuating transmission and selection
\subsection*{Randomly fluctuating transmission and selection}

We now consider the case that both transmission and selection fluctuate randomly.
Rewrite eq.~28 from \citet{Ram2018} so that $\rho$ is also a random variable
\begin{equation}
x_t = x_t \frac{1 + \rho_t s_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
\end{equation}
where $s_t$ are independent and identically distributed and also $\rho_t$ are independent and identically distributed ($t=0,1,2,\ldots$), $P(-1+C<s_t<D)=1$ for some positive $C$ and $D$, and $0<\rho_t<1$.
Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D$ and from Result~6 and 7 of \citet{Ram2018} we have:
\begin{itemize}
\item Suppose $E[log(1+\rho_t s_t)]>0$. Then $x^*=0$ is not stochastically locally stable and in fact $P(lim_{t \to \infty} x_t=0) = 0$.
\item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=1$ is stochastically locally stable and in particular if $E[\rho_t s_t] \le 0$ then $x^*=0$ is stochastically locally stable.
\end{itemize}

% Results
%\section*{Results}

% Discussion
%\section*{Discussion}

% Acknowledgements
%{\small
%\section*{Acknowledgements}
%
%This work was supported in part by 
%the Stanford Center for Computational, Evolutionary and Human Genomics, 
%and the Morrison Institute for Population and Resources Studies, Stanford University.
%}

\bibliographystyle{agsm}
%\bibliography{/Users/yoavram/Documents/library}
\bibliography{fluctuating_oblique}

\end{document}

### 2018-03-13 ms/fluctuating_oblique.tex
finite pop; beta dist rho; AkBk polymorphism

@@
 Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate.
+We start with the simple case in which the environment changes every generations, causing fluctuations in selection and transmission.
 Following \eqref{eq:recurrence}, the recurrence equations for two generations is then
@@
 \end{align} 
-From analysis of the quadratic $G(x)$ we can then determine that it has a root in $(0, 1)$, and that a protected polymorphism exists $x^*$ (i.e. $G(x^*)=0$ and $0<x^*<1$), if and only if 
+From analysis of the quadratic $G(x)$ we can then determine that it has a root in $(0, 1)$, and that a stable polymorphism exists $x^*$ (i.e. $G(x^*)=0$ and $0<x^*<1$), if and only if 
 \begin{equation}
@@
 \end{equation}
-Therefore, a protected polymorphism exists only if the difference in the transmission rate between consecutive generation is not too large, and, surprisingly, the stronger selection is, the larger the difference can be that still allows a protected polymorphism.
-Equivalently, a protected polymorphism exists if
+Therefore, a stable polymorphism exists only if the difference in the transmission rate between consecutive generations is not too large, and, surprisingly, the stronger selection is, the larger the difference can be that still allows a stable polymorphism.
+Equivalently, a stable polymorphism exists if
 \begin{equation} \label{eq:rho2_fluctuating_selection_transmission_k1}
@@
 
+More generally, if the environment changes every $k$ generations, so that for the first $k$ generations $A$ is favored and the transmission rate is $\rho_1$ and for the second $k$ generations $B$ is favored and the transmission rate is $\rho_2$, then, similar to eq.~(12) in~\citet{Ram2018}, the local stability of the fixation of $B$ ($x^*=0$) is determined by the product $[f_A(0)]^k[f_B(0)]^k$, and the local stability of fixation of $A$ ($x^*=1$) is determined by the product  $[f_A(1)]^k[f_B(1)]^k$, with
+\begin{equation}
+f_A(x) = \rho_1 \frac{1+s}{1+sx} + 1 - \rho_1, \quad \text{and} \quad 
+f_B(x) = \rho_2 \frac{1}{1+s-sx} + 1 - \rho_2.
+\end{equation}
+For both fixations to be locally unstable -- thus guaranteeing a protected polymorphism -- we require $[f_A(0)]^k[f_B(0)]^l>1$ and $[f_A(1)]^k[f_B(1)]^l>1$, which is guaranteed by \eqref{eq:rho2_fluctuating_selection_transmission_k1}.
+
 % Randomly fluctuating transmission and selection
@@
 \end{itemize}
+
+For example, if $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is likely when $A$ if favored ($s>0$) and oblique transmission is likely when $B$ is favored ($s<0$), then ($Beta$ is the beta function)
+\begin{equation}
+E[log(1+\rho_t s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{\rho^s \log{(1+\rho s)}}{Beta(1+s, 1)}  \; d\rho}\; ds} \approx 0.2525,
+\end{equation}
+and $B$ almost surely doesn't fix, and
+\begin{equation}
+E[log(1-\rho_t s_t / (1+s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{(1-\rho)^{-s/(1+s)} \log{(1-\rho s / (1+s))}}{Beta(1, 1-s/(1+s)}  \; d\rho}\; ds} \approx -0.076,
+\end{equation}
+so fixation of $A$ is stochastically locally stable.
+The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$): in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
+
+% Finite population size
+\subsection*{Finite population size}
+
+To include the effect of random drift due to finite population size in the above deterministic model, we follow \citet{Ram2018} and develop a diffusion equation approximation.
+In~\citet{Ram2018} only selection fluctuated via $s_t$, but here we also transmission to fluctuate via $\rho_t$.
+
+We find a result similar to Result 11 from~\citet{Ram2018}:
+The mean $\mu(x)$ and variance $\sigma^2(x)$ of the change in the frequency of $A$ in one generation for the diffusion approximation in the case of a cycling environment $AkBl$, where $k+l=n$, are
+\begin{equation}
+\mu(x) = S_n x(1-x), \quad \text{and} \quad \sigma^2(x) = n x (1-x)
+\end{equation}
+where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$, which is a weighted average of the selection coefficients in favor of phenotype $A$, weighted by the vertical transmission rate.
 

### 2018-03-14 ms/fluctuating_oblique.tex
reorganized and cleaned up math

@@
 Consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ and fitness values $1+s$ and $1$, respectively.
-Phenotypes are transmitted vertically with probability $\rho$ or diagonally with probability $1-\rho$.
+Phenotypes are transmitted vertically with probability $\rho$ or obliquely with probability $1-\rho$.
 Given $x$ the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is
@@
 Suppose that the vertical transmission rate $\rho$ is fluctuating, such that $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
-Following \eqref{eq:recurrence}, the recurrence equations for two generations is then
+From \eqref{eq:recurrence}, the recurrence equations for two generations is then
 \begin{equation}\begin{aligned} \label{eq:recurrence_two_generations}
@@
 
-To find the equilibrium $x^*$, that is, solutions of $x''=x$, we note that both fixations (i.e. $x=0$ and $x=1$) are equilibria. 
-Therefore, $x^*$ is a root of $G(x) = \frac{x''-x}{x(1-x)} \gamma$, where
-\begin{align}
-G(x) = ax^2+bx+c, \\
-\gamma = \frac{1+xs}{s} \Big( 1 + xs(2+s\rho_1) + x^2 s^2 (1-\rho_1) \Big) > 0, \\
-a = s^2 (1-\rho_1) (\rho_1 + \rho_2(1-\rho_1)) > 0, \\
-b = s(2+s\rho_1)(\rho_1 + \rho_2(1-\rho_1)) > 0, \\
-c = \rho_1 + \rho_1 \rho_2 s + \rho_2 > 0.
-\end{align} 
-Note that 
-\begin{align}
-G(0) = c > 0,\\ 
-G'(0) = b > 0,\\
-G(1)=a+b+c>c>0,\\
-G'(1)=2a+b>b>0,
-\end{align}
-and therefore $G(x)=0$ is not attained for any $x \in [0,1]$, and there is no steady polymorphism in \eqref{eq:recurrence_two_generations}.
-Also, the local stability of $x=0$ is determined by $(1+s\rho_1)(1+s\rho_2)$.
-Therefore, when $A$ is favored and $s>0$, $x=0$ is unstable and $x=1$ is stable, whereas the opposite applies when $B$ is favored, so that fixation of the favored phenotype is locally stable.
-
-Overall, we find that fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype without any protected polymorphisms.
+First, fixations of $A$ and $B$ ($x^*=0$ and $x^*=1$, respectively) are both equilibria, as they solve $x''=x$.
+Second, because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ and $x''/x = 1+\frac{\rho_2 s (1-x)}{1+sx} > 1$, the frequency of $A$ increases every generation and $x^*=1$ is globally stable.
+Therefore, we find that fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype without any stable polymorphisms.
 
@@
 
-Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate.
-We start with the simple case in which the environment changes every generations, causing fluctuations in selection and transmission.
-Following \eqref{eq:recurrence}, the recurrence equations for two generations is then
-\begin{equation}\begin{aligned} \label{eq:recurrence_two_generations_fluc_s}
-x' = \rho_1 \frac{1+s}{1+sx} x + (1-\rho_1)x, \\
-x'' = \rho_2 \frac{1}{1+s-sx} x' + (1-\rho_2)x'.
+Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together.
+
+If the environment changes every $k$ generations, so that for the first $k$ generations $A$ is favored and the transmission rate is $\rho_1$ and for the second $k$ generations $B$ is favored and the transmission rate is $\rho_2$.
+The change in the frequency of $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively, where
+\begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
+F_A(x) = \rho_1 \frac{1+s}{1+sx} x + (1-\rho_1)x, \\
+F_B(x) = \rho_2 \frac{1}{1+s-sx'} x' + (1-\rho_2)x'.
 \end{aligned}\end{equation}
+Fixations of $A$ ($x^*=1$) and $B$ ($x^*=0$) are locally stable if, respectively,
+\begin{equation}\begin{aligned}
+\big(F'_A(1) F'_B(1)\big)^k < 1, \\
+\big(F'_A(0) F'_B(0)\big)^k < 1. \\
+\end{aligned}\end{equation}
+and if these both of these conditions are not met then there exists a protected polymorphism.
+Therefore, for a protected polymorphism we require
+\begin{equation}
+1 < F'_A(1) F'_B(1) = \big(1-\rho_1\frac{s}{1+s}\big)\big(1+\rho_2 s\big) 
+= 1-\frac{1}{1+s}\big(\rho_2(1+s-\rho_1)-\rho_1\big),
+\end{equation}
+and
+\begin{equation}
+1 < F'_A(0) F'_B(0) = \big(1+\rho_1 s\big)\big(1-\rho_2 \frac{s}{1+s}\big) 
+= 1+\frac{s}{1+s}\big(\rho_1(1+s) -\rho_2(1+\rho_1 s)\big),
+\end{equation}
+which can be summarized as a condition on $\rho_2$
+\begin{equation} \label{eq:poly_condition_periodic_fluc}
+\frac{\rho_1}{1+s(1-\rho_1)} < \rho_2 < \frac{(1+s)\rho_1}{1+s\rho_1},
+\end{equation}
+or as a condition on the difference between $\rho_1$ and $\rho_2$
+\begin{equation}
+-s\rho_2(1-\rho_1) < \rho_2 - \rho_1 < s\rho_1(1-\rho_2).
+\end{equation}
 
-To find the equilibrium $x^*$, that is, solutions of $x''=x$, we note that both fixations (i.e. $x=0$ and $x=1$) are equilibria. 
-Therefore, $x^*$ is a root of $G(x) = \frac{x''-x}{x(1-x)} \gamma$, where
-\begin{align} \label{eq:Gx_fluctuating_selection_transmission_k1}
+\begin{figure*}[hbt]
+\centering
+\includegraphics[width=\linewidth]{../figures/{rho1_rho2_k=1}.pdf}
+\caption{
+\textbf{Protected polymorphism with fluctuating transmission and selection.}
+The positive root $x^*$ of $G(x)$, see eq.~\ref{eq:xstar_periodic_fluc_k=1}.
+Dashed lines represent, from left to right, $\rho_2=\frac{\rho_1}{1+s(1-\rho_1)}$, $\rho_2=\rho_1$, and $\rho_2=\frac{(1+s)\rho_1}{1+s\rho_1}$, see eq.~\ref{eq:poly_condition_periodic_fluc}.
+Here, $\rho_1=0.5$, $s=0.1$.}
+\label{fig:rho1_rho2_k=1}
+\end{figure*}
+
+\autoref{fig:rho1_rho2_k=1} shows the solution for $F_B(F_A(x))=x$ for $k=1$, highlighting the area of the parameter space in which a protected polymorphism exists.
+Surprisingly, the stronger selection is, the larger the difference can be that still allows a stable polymorphism.
+
+In the case of $k=1$ we can also find the frequency of $A$ at the protected polymorphism, $x^*$.
+Denote $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
+\begin{align} \label{eq:xstar_periodic_fluc_k=1}
 G(x) = ax^2+bx+c, \\
@@
 b = s(1-\rho_1)(2\rho_2 - s\rho_1(1-\rho_2)), \\
-c = \rho_2 - \rho_1 - s\rho_1(1-\rho_2).
+c = \rho_2 - \rho_1 - s\rho_1(1-\rho_2),
 \end{align} 
+and then $x^*$ is a solution of the quadratic $G(x)=0$, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
+The conditions for existence of the protected polymorphism \eqref{eq:poly_condition_periodic_fluc} is equivalent to $c<0$.% and $a+b+c>0$.
+Therefore, $\sqrt{b^2-4ac} > b$, and if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$ (\autoref{fig:rho1_rho2_k=1}).
 
-Note that 
-\begin{align}
-G(0) = c > 0  \Leftrightarrow \rho_2 > \frac{\rho_1(1+s)}{1+\rho_1s}, \\
-G'(0) = b > 0 \Leftrightarrow \rho_2 > \frac{\rho_1 s}{2 + \rho_1s}, \\
-G(1) = a + b + c = (1+s)[\rho_2 - \rho_1 + s \rho_2 (1-\rho_1)], \\
-G'(1) = 2a + b = s(1-\rho_1)(2(1+s)\rho_2 + s\rho_1(1-\rho_2)) > 0.
-\end{align} 
-From analysis of the quadratic $G(x)$ we can then determine that it has a root in $(0, 1)$, and that a stable polymorphism exists $x^*$ (i.e. $G(x^*)=0$ and $0<x^*<1$), if and only if 
-\begin{equation}
--s\rho_2(1-\rho_1) < \rho_2 - \rho_1 < s\rho_1(1-\rho_2),
-\end{equation}
-Therefore, a stable polymorphism exists only if the difference in the transmission rate between consecutive generations is not too large, and, surprisingly, the stronger selection is, the larger the difference can be that still allows a stable polymorphism.
-Equivalently, a stable polymorphism exists if
-\begin{equation} \label{eq:rho2_fluctuating_selection_transmission_k1}
-\frac{\rho_1}{1+s(1-\rho_1)} < \rho_2 < \frac{(1+s)\rho_1}{1+s\rho_1}.
-\end{equation}
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=\linewidth]{../figures/{rho1_rho2_k=1}.pdf}
-\caption{
-\textbf{Protected polymorphism with fluctuating transmission and selection.}
-The positive root $x^*$ of $G(x)$, see eq.~\ref{eq:Gx_fluctuating_selection_transmission_k1}.
-Dashed lines represent, from left to right, $\rho_2=\frac{\rho_1}{1+s(1-\rho_1)}$, $\rho_2=\rho_1$, and $\rho_2=\frac{(1+s)\rho_1}{1+s\rho_1}$, see eq.~\ref{eq:rho2_fluctuating_selection_transmission_k1}.
-Here, $\rho_1=0.5$, $s=0.1$.}
-\label{fig:rho1_rho2_k=1}
-\end{figure*}
-
-More generally, if the environment changes every $k$ generations, so that for the first $k$ generations $A$ is favored and the transmission rate is $\rho_1$ and for the second $k$ generations $B$ is favored and the transmission rate is $\rho_2$, then, similar to eq.~(12) in~\citet{Ram2018}, the local stability of the fixation of $B$ ($x^*=0$) is determined by the product $[f_A(0)]^k[f_B(0)]^k$, and the local stability of fixation of $A$ ($x^*=1$) is determined by the product  $[f_A(1)]^k[f_B(1)]^k$, with
-\begin{equation}
-f_A(x) = \rho_1 \frac{1+s}{1+sx} + 1 - \rho_1, \quad \text{and} \quad 
-f_B(x) = \rho_2 \frac{1}{1+s-sx} + 1 - \rho_2.
-\end{equation}
-For both fixations to be locally unstable -- thus guaranteeing a protected polymorphism -- we require $[f_A(0)]^k[f_B(0)]^l>1$ and $[f_A(1)]^k[f_B(1)]^l>1$, which is guaranteed by \eqref{eq:rho2_fluctuating_selection_transmission_k1}.
+%\begin{align}
+%G(0) = c > 0  \Leftrightarrow \rho_2 > \frac{\rho_1(1+s)}{1+\rho_1s}, \\
+%G'(0) = b > 0 \Leftrightarrow \rho_2 > \frac{\rho_1 s}{2 + \rho_1s}, \\
+%G(1) = a + b + c = (1+s)[\rho_2 - \rho_1 + s \rho_2 (1-\rho_1)], \\
+%G'(1) = 2a + b = s(1-\rho_1)(2(1+s)\rho_2 + s\rho_1(1-\rho_2)) > 0.
+%\end{align}
 

### 2018-03-14 src/get_bibtex.py
copy from another project; remove urls

@@ -1,0 +1,50 @@
+import re
+
+from bibtexparser import load as load_bib
+from bibtexparser.bwriter import BibTexWriter
+from bibtexparser.bibdatabase import BibDatabase
+import click
+
+species_name_pattern = re.compile(
+	r"^(.*?)\{\\textless\}i\{\\textgreater\}(.*?)\{\\textless\}\/i\{\\textgreater\}(.*?)$")
+
+@click.command()
+@click.argument('keys_filename', type=click.Path(exists=True, file_okay=True, dir_okay=False, readable=True))
+@click.argument('bibtex_filename', type=click.Path(exists=True, file_okay=True, dir_okay=False, readable=True))
+@click.argument('output_filename', type=click.Path(file_okay=True, dir_okay=False, writable=True))
+@click.option('-v/-V', '--verbose/--no-verbose', default=False)
+def main(keys_filename, bibtex_filename, output_filename, verbose):
+	with open(keys_filename) as f:
+		citation_keys = [line.strip() for line in f.readlines()]
+	if verbose:
+		print("Read {} keys from {}".format(len(citation_keys), keys_filename))
+
+	with open(bibtex_filename) as f:
+		main_bib = load_bib(f)
+	if verbose:
+		print("Read {} entries from {}".format(len(main_bib.entries), bibtex_filename))
+
+	out_bib = BibDatabase()
+	for key in citation_keys:
+		e = main_bib.entries_dict[key]
+		out_bib.entries.append(e)
+	
+	if verbose:
+		print("Writing {} entries to {}".format(len(out_bib.entries), output_filename))
+
+	for ent in out_bib.entries:
+		m = species_name_pattern.match(ent['title'])
+		if m:
+			prefix, species, postfix = m.groups()
+			ent['title'] = prefix + r"\emph{" + species + r"}" + postfix
+		for key in ['file', 'abstract', 'mendeley-tags', 'keyword', 'URL']:
+			if key in ent:
+				ent.pop(key)		
+
+	writer = BibTexWriter()
+	with open(output_filename, 'w') as f:
+	    f.write(writer.write(out_bib))
+
+
+if __name__ == '__main__':
+	main()

### 2018-03-14 ms/fluctuating_oblique.tex
update

@@
 \begin{itemize}
-\item Suppose $E[log(1+\rho_t s_t)]>0$. Then $x^*=0$ is not stochastically locally stable and in fact $P(lim_{t \to \infty} x_t=0) = 0$.
-\item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=1$ is stochastically locally stable and in particular if $E[\rho_t s_t] \le 0$ then $x^*=0$ is stochastically locally stable.
+\item Suppose $E[log(1+\rho_t s_t)]>0$. Then $x^*=0$ is not stochastically locally stable and in fact $P(lim_{t \to \infty} x_t=0) = 0$, i.e., fixation of $B$ almost surely does not occur.
+\item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=0$ is stochastically locally stable and 
+\item Similarly, $E[log(1-\rho_t s_t/(1+s_t)]<0$ leads to stochastic local stability of $x^*=1$, and $E[log(1-\rho_t s_t/(1+s_t))]>0$ assures that fixation of $A$ almost surely does not occur.
+\item In particular, if $E[\rho_t s_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-\rho_t s_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
 \end{itemize}
 
-For example, if $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is likely when $A$ if favored ($s>0$) and oblique transmission is likely when $B$ is favored ($s<0$), then ($Beta$ is the beta function)
+We will now present a number of examples.
+First, if $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then both $E[log(1+\rho_t s_t)]$ and $E[log(1-\rho_t s_t/(1+s_t))]$ are negative, and therefore both fixations ($A$ and $B$) are stochastically locally stable.
+It is important to note that without fluctuations in $\rho$ (i.e. fixed $\rho$) it is not possible to have both fixations stable~\citep{Ram2018}.
+
+Second, if $s_t$ is symmetric around zero ($E[s_t]=E[s_t/(1+s_t)]=0$), and $\rho_t$ and $s_t$ are positively (negatively) correlated, such that vertical transmission is more likely when $A$ ($B$) is favored, then $E[-\rho_t s_t/(1+s_t)]<0$ ($E[-\rho_t s_t]<0$) and therefore fixation of $A$ ($B$) is stochastically locally stable.
+For example, let $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$), then ($Beta$ being the beta function),
 \begin{equation}
@@
 so fixation of $A$ is stochastically locally stable.
-The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$): in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
+The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$); in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
+
+Third, it is also possible that both fixations are stochastically locally unstable, but, similar to the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
+For example, \autoref{fig:rho1_rho2_stoch_p} shows the expected outcome when $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
+The blue and red areas denote expected fixation of $A$ and $B$, respectively (i.e.
+stochastic local stability) and the white area shows expected polymorphism (both fixations are not stochastically locally stable).
+
+\begin{figure*}[hbt]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/{rho1_rho2_stoch_p}.pdf}
+\caption{
+\textbf{Stochastic local stability.}
+Here, $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$, with $s=0.05$ and $p=0.505$, a value combination that allowed a stochastic polymorphism with a fixed vertical transmission rate $\rho=0.1$ in~\citet[Fig.~2]{Ram2018}.}
+\label{fig:rho1_rho2_stoch_p}
+\end{figure*}
 

### 2018-03-15 ms/fluctuating_oblique.tex
clean typos; added beta figure

@@
 and then $x^*$ is a solution of the quadratic $G(x)=0$, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
-The conditions for existence of the protected polymorphism \eqref{eq:poly_condition_periodic_fluc} is equivalent to $c<0$.% and $a+b+c>0$.
+The condition \eqref{eq:poly_condition_periodic_fluc} guarantees that $x^* \in (0,1)$ and is equivalent to $c<0$.% and $a+b+c>0$.
 Therefore, $\sqrt{b^2-4ac} > b$, and if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$ (\autoref{fig:rho1_rho2_k=1}).
@@
 
-Second, if $s_t$ is symmetric around zero ($E[s_t]=E[s_t/(1+s_t)]=0$), and $\rho_t$ and $s_t$ are positively (negatively) correlated, such that vertical transmission is more likely when $A$ ($B$) is favored, then $E[-\rho_t s_t/(1+s_t)]<0$ ($E[-\rho_t s_t]<0$) and therefore fixation of $A$ ($B$) is stochastically locally stable.
-For example, let $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$), then ($Beta$ being the beta function),
+Second, 
+%if $s_t$ is symmetric around zero ($E[s_t]=0$), and $\rho_t$ and $s_t$ are negatively correlated, such that vertical transmission is more likely when $B$ is favored, then $E[-\rho_t s_t]<0$ and therefore fixation of $B$ is stochastically locally stable.
+let $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$ (\autoref{fig:beta}), such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$), then ($Beta$ being the beta function),
 \begin{equation}
@@
 The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$); in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
+
+\begin{figure*}[hbt]
+\centering
+\includegraphics[width=\linewidth]{../figures/{beta}.pdf}
+\caption{
+\textbf{Beta distribution.}
+\textbf{(A)} Distribution of $\rho_t \sim Beta(1+s_t, 1)$ for different values of $s_t$.
+It can be seen that for $s_t>0$ vertical transmission is more likely, and for $s_t<0$ oblique transmission is more likely.
+\textbf{(B)} The compound distribution of $\rho_t \sim Beta(1+s_t, 1)$ for $s_t \sim U(-1,1)$, demonstrating that, overall, oblique transmission is more likely than vertical transmission.
+\textbf{(C)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation.
+}
+\label{fig:beta}
+\end{figure*}
 

### 2018-03-15 ms/fluctuating_oblique.tex
merged notebooks; fixed UL comments

--- 2018-03-14 ms/fluctuating_oblique.tex
@@
 \item In particular, if $E[\rho_t s_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-\rho_t s_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
+\item Note that it is not possible that both $E[log(1+\rho_t s_t)]$ and $E[log(1-\rho_t s_t/(1+s_t)]$ are negative, as their sum is positive
+\begin{multline}
+E[log(1+\rho_t s_t)] + E[log(1-\rho_t s_t/(1+s_t)] = \\
+E[log(1+\rho_t s_t) + log(1-\rho_t s_t/(1+s_t)]= \\
+E[log\big((1+\rho_t s_t)(1-\rho_t s_t/(1+s_t)\big)]= \\
+E[log\big( 1+\rho_t(1-\rho_t)(s_t^2)/(1+s_t) \big)] > 0,
+\end{multline}
+and therefore it is not possible that both fixations are stochastically locally stable.
 \end{itemize}
@@
 We will now present a number of examples.
-First, if $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then both $E[log(1+\rho_t s_t)]$ and $E[log(1-\rho_t s_t/(1+s_t))]$ are negative, and therefore both fixations ($A$ and $B$) are stochastically locally stable.
-It is important to note that without fluctuations in $\rho$ (i.e. fixed $\rho$) it is not possible to have both fixations stable~\citep{Ram2018}.
+First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)$) and $s_t$ is symmetric around zero, then $E[\rho_t s_t]=0$ and $E[-\rho_t s_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
+Therefore, fixation of $B$ is stochastically locally stable and fixation of $A$ almost surely doesn't occur.
+For example, let $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then $E[log(1+\rho_t s_t)]\approx -0.07315$ and $E[log(1-\rho_t s_t/(1+s_t))]\approx 0.2337$.
+However, note that symmetry of $s_t$ around zero provides an advantage to phenotype $B$ -- using Jensen's inequality, $E[w_A/w_B] = E[1+s_t] = 1 \le E[1/(1+s_t)] = E[w_B/w_A]$.
+Therefore, if we take $w_{A,t}, w_{B,t} \sim U(0,1)$ independently, and $s_t=(w_A-w_B)/w_B$, then neither $A$ nor $B$ has an advantage, on average (i.e.
+$E[w_A/w_B]=E[w_B/w_A]$, see \autoref{fig:beta}A), and both $E[\rho_t s_t]$ and $E[-\rho_t s_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable.
 
-Second, if $s_t$ is symmetric around zero ($E[s_t]=E[s_t/(1+s_t)]=0$), and $\rho_t$ and $s_t$ are positively (negatively) correlated, such that vertical transmission is more likely when $A$ ($B$) is favored, then $E[-\rho_t s_t/(1+s_t)]<0$ ($E[-\rho_t s_t]<0$) and therefore fixation of $A$ ($B$) is stochastically locally stable.
-For example, let $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$), then ($Beta$ being the beta function),
-\begin{equation}
-E[log(1+\rho_t s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{\rho^s \log{(1+\rho s)}}{Beta(1+s, 1)}  \; d\rho}\; ds} \approx 0.2525,
-\end{equation}
-and $B$ almost surely doesn't fix, and
-\begin{equation}
-E[log(1-\rho_t s_t / (1+s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{(1-\rho)^{-s/(1+s)} \log{(1-\rho s / (1+s))}}{Beta(1, 1-s/(1+s)}  \; d\rho}\; ds} \approx -0.076,
-\end{equation}
-so fixation of $A$ is stochastically locally stable.
-The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$); in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
+If instead we do not assume independence of $s_t$ and $\rho_t$, we can get fixation.
+Let $\rho_t \sim \beta(1+s, 1)$, where $s_t=(w_A-w_B)/w_B$ and $w_A, w_B \sim U(0,1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$; \autoref{fig:beta}).
+Then $E[log(1+\rho_t s_t)] >0$ and $B$ almost surely doesn't fix, and $E[log(1-\rho_t s_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
+The opposite occurs if $\rho_t \sim \beta(1, 1+s)$ and the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -4$).
+In that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
 
-Third, it is also possible that both fixations are stochastically locally unstable, but, similar to the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
+\begin{figure*}[hbt]
+\centering
+\includegraphics[width=\linewidth]{../figures/{beta}.png}
+\caption{
+\textbf{Covariance of selection and transmission.}
+\textbf{(A)} Histogram of $w_A/w_B$ where $w_A$ and $w_B$ are identically and independently distributed uniform random variables $U(0,1)$.
+\textbf{(B)} Histogram of $s_t = (w_A-w_B)/w_B$.
+\textbf{(C)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
+\textbf{(C)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
+}
+\label{fig:beta}
+\end{figure*}
+
+Third, it is also possible that both fixations are not stochastically locally stable even if $s_t$ and $\rho_t$ covary, but, similar to the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
 For example, \autoref{fig:rho1_rho2_stoch_p} shows the expected outcome when $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.

### 2018-03-15 ms/fluctuating_oblique.tex
first draft of intro

@@
 % Introduction
-%\section*{Introduction}
+\section*{Introduction}
+
+Cultural transmission comprises of both vertical transmission (from parent to offspring) and non-vertical transmission, namely horizontal (between peers) and oblique (from non-parental adults to offspring).
+\citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific trait is transmitted either vertically or obliquely.
+Others models include different forms of cultural transmission that include both vertical and oblique transmission; usually the transmission strategy of each individual is fixed: for example, \citet{McElreath2008} and \cite{Aoki2005} focused on competitions between individual, vertical (innate), and oblique or horizontal learners; \citet{Fogarty2017} compared the effect of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
+In many of these models the effects of fluctuating selection and environmental changes was studied~\citep[reviewed in][]{Aoki2014}. 
+For example, we recently studied a model in which each individual can learn his phenotype either from a parent, with probability $\rho$, or from a non-parental adult, with probability $1-\rho$; see \autoref{fig:transmission}.
+We found that if selection fluctuates between favoring two phenotypes, but on average favored both phenotypes for similar time periods, then a polymorphism can be maintained~\citep{Ram2018}.
+
+However, the case in which transmission also fluctuates has not received much attention.
+Nevertheless, if we assume that social learning can be affected by weather, food availability, population density, and even political atmosphere, then it is only logical that the parameters of transmission will fluctuate over time.
+For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults, therefore learn obliquely.
+
+Here, we analyze a model in which the probability $\rho$ that individuals learn from their parents (vertical transmission), rather than from non-parental adults (oblique transmission), fluctuates over time.
+We find that a polymorphic population cannot be maintained if unless these fluctuations are very small and only weakly correlated with fluctuations in selection.
+We suggest that the effect of environmental changes can affect cultural evolution not only through their effect of cultural selection, but also on cultural transmission.
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width=0.5\linewidth]{../figures/{transmission}.png}
+\caption{
+\textbf{Cultural transmission with mixed vertical and oblique transmission.}
+When a newborn matures, she will copy her phenotype -- color -- from her mother with probability $\rho$, therefore becoming blue, or from some other female with probability $1-\rho$, in which case her color will depend on the frequency of blue and red adult females.}
+\label{fig:transmission}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=\linewidth]{../figures/{rho1_rho2_k=1}.pdf}
+\includegraphics[width=0.65\linewidth]{../figures/{rho1_rho2_k=1}.pdf}
 \caption{
@@
 \bibliographystyle{agsm}
-%\bibliography{/Users/yoavram/Documents/library}
-\bibliography{fluctuating_oblique}
+\bibliography{/Users/yoavram/Documents/library}
+%\bibliography{fluctuating_oblique}
 

### 2018-03-16 ms/fluctuating_oblique.tex
intro revision by MWF

--- 2018-03-15 ms/fluctuating_oblique.tex
@@
 
-%\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
+\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
 \usepackage{graphicx} % Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
@@
 % Introduction
-%\section*{Introduction}
+\section*{Introduction}
+
+Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from non-parental adults to offspring) transmission. 
+\citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific trait is transmitted either vertically or obliquely.
+Mathematical models of cultural evolution may include one or more modes of cultural transmission and in these models the mode of transmission by individuals is fixed. 
+For example, \cite{Aoki2005} and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
+
+Many of these models included fluctuating selection envisaged as due to environmental changes~\citep[reviewed in][]{Aoki2014}. 
+For example, we recently studied a model in which each individual can learn a dichotomous phenotype either from a parent, with probability $\rho$, or from a non-parental adult, with probability $1-\rho$; see \autoref{fig:transmission}.
+We found that if selection fluctuates between favoring each of the two phenotypes, but on average favored both phenotypes for similar time periods, then a phenotypic polymorphism may be maintained~\citep{Ram2018}.
+
+There has been much less theory developed for fluctuating transmission.
+Nevertheless, if we assume that social learning can be affected by weather, food availability, population density, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
+
+Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from non-parental adults with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time~(\autoref{fig:transmission}).
+We find that a phenotypic polymorphism cannot be maintained unless fluctuations in the mode of transmission are very small and only weakly correlated with fluctuations in selection.
+We suggest that cultural evolution is affected by environmental changes that may affect both selection and transmission, and that further empirical and theoretical studies are required to disentangle these effects.
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width=0.5\linewidth]{../figures/{transmission}.png}
+\caption{
+\textbf{Cultural transmission with mixed vertical and oblique transmission.}
+When a newborn matures, she will copy her phenotype -- color -- from her mother with probability $\rho$, therefore becoming blue, or from some other female with probability $1-\rho$, in which case her color will depend on the frequency of blue and red adult females.}
+\label{fig:transmission}
+\end{figure*}
 
@@
 
-Consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ and fitness values $1+s$ and $1$, respectively.
+Consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ and fitness values $w_A=1+s$ and $w_B=1$, respectively.
 Phenotypes are transmitted vertically with probability $\rho$ or obliquely with probability $1-\rho$.
@@
 
-We now consider the case that both transmission and selection fluctuate randomly.
+We now allow both transmission and selection to fluctuate randomly.
 Rewrite eq.~28 from \citet{Ram2018} so that $\rho$ is also a random variable
@@
 \end{equation}
-where $s_t$ are independent and identically distributed and also $\rho_t$ are independent and identically distributed ($t=0,1,2,\ldots$), $P(-1+C<s_t<D)=1$ for some positive $C$ and $D$, and $0<\rho_t<1$.
+where $s_t$ are i.i.d (independent and identically distributed), $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$, $\rho_t$ are i.i.d, and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
 Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D$ and from Result~6 and 7 of \citet{Ram2018} we have:
@@
 \item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=0$ is stochastically locally stable and 
-\item Similarly, $E[log(1-\rho_t s_t/(1+s_t)]<0$ leads to stochastic local stability of $x^*=1$, and $E[log(1-\rho_t s_t/(1+s_t))]>0$ assures that fixation of $A$ almost surely does not occur.
+\item Similarly, if $E[log(1-\rho_t s_t/(1+s_t)]<0$, then $x^*=1$ is stochastically locally stable, and if $E[log(1-\rho_t s_t/(1+s_t))]>0$, then fixation of $A$ almost surely does not occur.
 \item In particular, if $E[\rho_t s_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-\rho_t s_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
@@
 First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)$) and $s_t$ is symmetric around zero, then $E[\rho_t s_t]=0$ and $E[-\rho_t s_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
-Therefore, fixation of $B$ is stochastically locally stable and fixation of $A$ almost surely doesn't occur.
+Therefore, fixation of $B$ is stochastically locally stable and fixation of $A$ almost surely does not occur.
 For example, let $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then $E[log(1+\rho_t s_t)]\approx -0.07315$ and $E[log(1-\rho_t s_t/(1+s_t))]\approx 0.2337$.
@@
 
-If instead we do not assume independence of $s_t$ and $\rho_t$, we can get fixation.
-Let $\rho_t \sim \beta(1+s, 1)$, where $s_t=(w_A-w_B)/w_B$ and $w_A, w_B \sim U(0,1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$; \autoref{fig:beta}).
+Second, if $s_t$ and $\rho_t$ are not independent, we can get fixation.
+Let $\rho_t \sim \beta(1+s, 1)$, where $s_t=(w_A-w_B)/w_B$ and $w_A, w_B \sim U(0,1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$); that is, vertical transmission is more likely when $A$ is favored (i.e.
+$s_t>0$) and oblique transmission is more  likely when $B$ is favored (i.e.
+$s_t<0$; \autoref{fig:beta}).
 Then $E[log(1+\rho_t s_t)] >0$ and $B$ almost surely doesn't fix, and $E[log(1-\rho_t s_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
@@
 
-Third, it is also possible that both fixations are not stochastically locally stable even if $s_t$ and $\rho_t$ covary, but, similar to the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
+Third, it is also possible that both fixations are not stochastically locally stable even if $s_t$ and $\rho_t$ covary, but, as in the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
 For example, \autoref{fig:rho1_rho2_stoch_p} shows the expected outcome when $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
 The blue and red areas denote expected fixation of $A$ and $B$, respectively (i.e.
-stochastic local stability) and the white area shows expected polymorphism (both fixations are not stochastically locally stable).
+stochastic local stability) and the white area shows expected polymorphism (neither fixation is stochastically locally stable).
 
@@
 \textbf{Stochastic local stability.}
-Here, $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$, with $s=0.05$ and $p=0.505$, a value combination that allowed a stochastic polymorphism with a fixed vertical transmission rate $\rho=0.1$ in~\citet[Fig.~2]{Ram2018}.}
+Here, $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$, with $s=0.05$ and $p=0.505$, a combination that allows a stochastic polymorphism with a fixed vertical transmission rate $\rho=0.1$ in~\citet[Fig.~2]{Ram2018}.}
 \label{fig:rho1_rho2_stoch_p}

### 2018-03-16 ms/fluctuating_oblique.tex
k-l result

@@
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by weather, food availability, population density, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+Nevertheless, if we assume that social learning can be affected by weather~\citep{Phithakkitnukoon2012}, food availability, population density, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
 For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
@@
 Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from non-parental adults with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time~(\autoref{fig:transmission}).
-We find that a phenotypic polymorphism cannot be maintained unless fluctuations in the mode of transmission are very small and only weakly correlated with fluctuations in selection.
+We find that $\ldots$ %TODO
 We suggest that cultural evolution is affected by environmental changes that may affect both selection and transmission, and that further empirical and theoretical studies are required to disentangle these effects.
@@
 
-% Model
-\section*{Model}
+% Model & Results
+\section*{Model \& Results}
 
 Consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ and fitness values $w_A=1+s$ and $w_B=1$, respectively.
-Phenotypes are transmitted vertically with probability $\rho$ or obliquely with probability $1-\rho$.
+Phenotypes are transmitted vertically with probability $\rho$ or obliquely with probability $1-\rho$~(\autoref{fig:transmission}).
 Given $x$ the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is
@@
 \begin{equation}\begin{aligned}
-\big(F'_A(1) F'_B(1)\big)^k < 1, \\
-\big(F'_A(0) F'_B(0)\big)^k < 1. \\
+\big[F'_A(1) F'_B(1)\big]^k < 1, \\
+\big[F'_A(0) F'_B(0)\big]^k < 1. \\
 \end{aligned}\end{equation}
-and if these both of these conditions are not met then there exists a protected polymorphism.
+and if both of these conditions are not met then there exists a protected polymorphism.
 Therefore, for a protected polymorphism we require
 \begin{equation}
-1 < F'_A(1) F'_B(1) = \big(1-\rho_1\frac{s}{1+s}\big)\big(1+\rho_2 s\big) 
-= 1-\frac{1}{1+s}\big(\rho_2(1+s-\rho_1)-\rho_1\big),
+1 < F'_A(1) F'_B(1) = \Big(1-\rho_1\frac{s}{1+s}\Big)\Big(1+\rho_2 s\Big) 
+= 1-\frac{1}{1+s}\big[\rho_2(1+s-\rho_1)-\rho_1\big],
 \end{equation}
@@
 \begin{equation}
-1 < F'_A(0) F'_B(0) = \big(1+\rho_1 s\big)\big(1-\rho_2 \frac{s}{1+s}\big) 
-= 1+\frac{s}{1+s}\big(\rho_1(1+s) -\rho_2(1+\rho_1 s)\big),
+1 < F'_A(0) F'_B(0) = \Big(1+\rho_1 s\Big)\Big(1-\rho_2 \frac{s}{1+s}\Big) 
+= 1+\frac{s}{1+s}\big[\rho_1(1+s) -\rho_2(1+\rho_1 s)\big],
 \end{equation}
@@
 %\end{align}
+
+In the more general case, phenotype $A$ is favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle such that $\rho=\rho_1$ when $A$ is favored and $\rho=\rho_2$ when $B$ is favored.
+The requirements for a protected polymorphism is now
+\begin{equation}\begin{aligned}
+1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b-\Delta \rho s/(1+s)\big)^l,\\
+1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a+\Delta \rho s)^l,
+\end{aligned}\end{equation}
+where $a=1+\rho_1 s>1$, $b=1-\rho_1\frac{s}{1+s}<1$, and $\Delta \rho = \rho_2 - \rho_1$.
+This leads to a condition similar to eq.~20 in~\citet{Ram2018}, but more complex due to the addition of $\Delta \rho$:
+\begin{equation} \label{eq:poly_condition_periodic_fluc_k_l}
+\frac{-\log{b}}{\log{\big(a+\Delta \rho s\big)}} < 
+\frac{l}{k} < 
+\frac{\log{a}}{-\log{\big(b-\Delta \rho s/(1+s)\big)}}.
+\end{equation}
+Note that as $\Delta \rho$ increases, both the RHS and the LHS of \eqref{eq:poly_condition_periodic_fluc_k_l} decrease.
+Therefore, increasing the vertical transmission rate specifically when $B$ is favored ($\Delta \rho>0$) decreases the environmental period ratio $l/k$ that permits protected polymorphism; reducing the vertical transmission rate when $B$ is favored ($\Delta \rho<0$) will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
 
@@
 
-% Results
-%\section*{Results}
+% Discussion
+\section*{Discussion}
 
-% Discussion
-%\section*{Discussion}
+Models of cultural transmission mostly assume a fixed relative rate at which different modes of transmission -- vertical, horizontal, or oblique -- occur.
+Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly. 
+Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that transmission fluctuations constrain the maintenance of phenotypic polymorphism.
+%Specifically, if fluctuations are not small then a single phenotype is likely to fix in the population~(\autoref{eq:poly_condition_periodic_fluc}, \autoref{fig:rho1_rho2_k=1}, \autoref{fig:rho1_rho2_stoch_p}), and
+
+% Face book rural internet
+% National service - Israel, Singapore
+% Immigrtion - draught
+% Opening schools, universities
+% sending students abroad, bringing experts from abroad
+% foreign labor
+% refugees
+% hunter-gatherers/chimps increase habitat in years with less food production
+% religion - dark ages etc.
 

### 2018-03-20 ms/fluctuating_oblique.tex
citation, todo

@@
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by weather~\citep{Phithakkitnukoon2012}, food availability, population density, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+Nevertheless, if we assume that social learning can be affected by frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, food availability, population density~\citep{Aureli1997a}, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
 For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
@@
 
+% TODO plot \Delta \rho vs s and show the boundaries
+
 % Randomly fluctuating transmission and selection

### 2018-03-20 ms/fluctuating_oblique.tex
UL addition to constant selection

@@
 First, fixations of $A$ and $B$ ($x^*=0$ and $x^*=1$, respectively) are both equilibria, as they solve $x''=x$.
-Second, because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ and $x''/x = 1+\frac{\rho_2 s (1-x)}{1+sx} > 1$, the frequency of $A$ increases every generation and $x^*=1$ is globally stable.
+Second, when $s>0$ ($s<0$) because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ ($<1$) and $x''/x = 1+\frac{\rho_2 s (1-x)}{1+sx} > 1$ ($<1$), the frequency of $A$ ($B$) increases every generation and $x^*=1$ ($x^*=0$) is globally stable.
 Therefore, we find that fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype without any stable polymorphisms.

### 2018-03-20 ms/fluctuating_oblique.tex
new discussion paragraphs; citations

@@
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, food availability, population density~\citep{Aureli1997a}, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+Nevertheless, if we assume that social learning can be affected by frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, stress~\citep{Farine2015}, food availability, population density~\citep{Aureli1997a}, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
 For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
@@
 
-% Face book rural internet
+One caveat of our model is the use of the ``phenotypic gambit'' -- the assumption that the transmission mode itself is strictly vertically transmitted.
+Although there is evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption may be a challenged: at least in humans, individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
+It is also possible that the transmission mode is regulated.
+For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
+Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only some individuals do so.
+
+Another extension of our model could incorporate more sophisticated oblique transmission schemes.
+For example, conformity -- preference for learning a frequent phenotype -- has been demonstrated in wild vervet monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
+We suggest that the oblique transmission mode can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions.
+
+% Facebook rural internet
 % National service - Israel, Singapore
@@
 \bibliographystyle{agsm}
-\bibliography{/Users/yoavram/Documents/library}
-%\bibliography{fluctuating_oblique}
+%\bibliography{/Users/yoavram/Documents/library}
+\bibliography{fluctuating_oblique}
 

### 2018-03-21 ms/fluctuating_oblique.tex
revision; more citations

--- 2018-03-20 ms/fluctuating_oblique.tex
@@
 \citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific trait is transmitted either vertically or obliquely.
-Mathematical models of cultural evolution may include one or more modes of cultural transmission and in these models the mode of transmission by individuals is fixed. 
+Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed. 
 For example, \cite{Aoki2005} and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
@@
 Many of these models included fluctuating selection envisaged as due to environmental changes~\citep[reviewed in][]{Aoki2014}. 
-For example, we recently studied a model in which each individual can learn a dichotomous phenotype either from a parent, with probability $\rho$, or from a non-parental adult, with probability $1-\rho$; see \autoref{fig:transmission}.
+We recently studied a model in which each individual can learn a dichotomous phenotype either from a parent, with probability $\rho$, or from a non-parental adult, with probability $1-\rho$ (\autoref{fig:transmission}).
 We found that if selection fluctuates between favoring each of the two phenotypes, but on average favored both phenotypes for similar time periods, then a phenotypic polymorphism may be maintained~\citep{Ram2018}.
+Furthermore, we found that if the environment changes very rapidly then lower $\rho$ values are likely to evolve, that is, oblique transmission is favored over vertical transmission.
 
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, food availability, population density~\citep{Aureli1997a}, or even political atmosphere, then it is reasonable that the mode and rate of transmission will fluctuate over time.
-For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
+Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size~\citep{Fischer2015} or density~\citep{Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases.
+%For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
 
@@
 We find that $\ldots$ %TODO
-We suggest that cultural evolution is affected by environmental changes that may affect both selection and transmission, and that further empirical and theoretical studies are required to disentangle these effects.
+We suggest that cultural evolution can be affected by environmental changes via both selection and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
 
@@
 
-First, fixations of $A$ and $B$ ($x^*=0$ and $x^*=1$, respectively) are both equilibria, as they solve $x''=x$.
-Second, because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ and $x''/x = 1+\frac{\rho_2 s (1-x)}{1+sx} > 1$, the frequency of $A$ increases every generation and $x^*=1$ is globally stable.
+First, fixations of $A$ and $B$ ($x^*=1$ and $x^*=0$, respectively) are both equilibria, as they solve $x''=x$.
+Second, when $s>0$ ($s<0$) because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ ($<1$) and $x''/x' = 1+\frac{\rho_2 s (1-x')}{1+sx'} > 1$ ($<1$), the frequency of $A$ ($B$) increases every generation and $x^*=1$ ($x^*=0$) is globally stable.
 Therefore, we find that fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype without any stable polymorphisms.
@@
 
-Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together.
-
-If the environment changes every $k$ generations, so that for the first $k$ generations $A$ is favored and the transmission rate is $\rho_1$ and for the second $k$ generations $B$ is favored and the transmission rate is $\rho_2$.
+Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$ when $B$ is favored the transmission rate is $\rho_B$.
 The change in the frequency of $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively, where
 \begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
-F_A(x) = \rho_1 \frac{1+s}{1+sx} x + (1-\rho_1)x, \\
-F_B(x) = \rho_2 \frac{1}{1+s-sx'} x' + (1-\rho_2)x'.
+F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, \\
+F_B(x) = \rho_B \frac{1}{1+s-sx'} x' + (1-\rho_B)x'.
 \end{aligned}\end{equation}
-Fixations of $A$ ($x^*=1$) and $B$ ($x^*=0$) are locally stable if, respectively,
+
+\paragraph{Symmetric periods: $AkBk$.}
+We first consider environments which periodically fluctuate between favoring $A$ and $B$ every $k$ generations.
+Using a linear approximation, fixations of $A$ ($x^*=1$) and $B$ ($x^*=0$) are locally stable if, respectively,
 \begin{equation}\begin{aligned}
@@
 and if both of these conditions are not met then there exists a protected polymorphism.
-Therefore, for a protected polymorphism we require
+So for a protected polymorphism we require
 \begin{equation}
-1 < F'_A(1) F'_B(1) = \Big(1-\rho_1\frac{s}{1+s}\Big)\Big(1+\rho_2 s\Big) 
-= 1-\frac{1}{1+s}\big[\rho_2(1+s-\rho_1)-\rho_1\big],
+1 < F'_A(1) F'_B(1) = \Big(1-\rho_A\frac{s}{1+s}\Big)\Big(1+\rho_B s\Big) 
+= 1-\frac{1}{1+s}\big[\rho_B(1+s-\rho_A)-\rho_A\big],
 \end{equation}
@@
 \begin{equation}
-1 < F'_A(0) F'_B(0) = \Big(1+\rho_1 s\Big)\Big(1-\rho_2 \frac{s}{1+s}\Big) 
-= 1+\frac{s}{1+s}\big[\rho_1(1+s) -\rho_2(1+\rho_1 s)\big],
+1 < F'_A(0) F'_B(0) = \Big(1+\rho_A s\Big)\Big(1-\rho_B \frac{s}{1+s}\Big) 
+= 1+\frac{s}{1+s}\big[\rho_A(1+s) -\rho_B(1+\rho_A s)\big],
 \end{equation}
-which can be summarized as a condition on $\rho_2$
+which can be summarized as a condition on $\rho_B$,
 \begin{equation} \label{eq:poly_condition_periodic_fluc}
-\frac{\rho_1}{1+s(1-\rho_1)} < \rho_2 < \frac{(1+s)\rho_1}{1+s\rho_1},
+\frac{\rho_A}{1+s(1-\rho_A)} < \rho_B < \frac{(1+s)\rho_A}{1+s\rho_A},
 \end{equation}
-or as a condition on the difference between $\rho_1$ and $\rho_2$
+or as a condition on the difference between $\rho_B$ and $\rho_A$
 \begin{equation}
--s\rho_2(1-\rho_1) < \rho_2 - \rho_1 < s\rho_1(1-\rho_2).
+-s\rho_B(1-\rho_A) < \rho_B - \rho_A < s\rho_A(1-\rho_B).
 \end{equation}
+Surprisingly, the stronger selection is, the larger the difference can be that still allows a stable polymorphism.
 
-\begin{figure*}[hbt]
+\paragraph{Period of one: $A1B1$.}
+In the case of $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
+Denote $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
+\begin{equation} \begin{aligned} \label{eq:xstar_periodic_fluc_k=1}
+G(x) &= ax^2+bx+c, \\
+\gamma &= -\frac{1 + sx}{s} \Big[ 1 + s + s^2 x (1-\rho)(1+x)\Big], \\
+a &= s^2 \rho_A (1-\rho_A) (1 - \rho_B + \rho_B/\rho_A)) > 0, \\
+b &= s(1-\rho_A)(2\rho_B - s\rho_A(1-\rho_B)), \\
+c &= \rho_B - \rho_A - s\rho_A(1-\rho_B),
+\end{aligned} \end{equation}
+% $a+b+c>0$.
+and then $x^*$ is a solution of the quadratic $G(x)=0$. %, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
+The condition \eqref{eq:poly_condition_periodic_fluc}, which guarantees that $0 < x^* < 1$, is equivalent to $c<0$, and therefore it also guarantees that $\sqrt{b^2-4ac} > b$.
+So, if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$.
+\autoref{fig:rho1_rho2_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
+
+%\begin{align}
+%G(0) = c > 0  \Leftrightarrow \rho_B > \frac{\rho_A(1+s)}{1+\rho_As}, \\
+%G'(0) = b > 0 \Leftrightarrow \rho_B > \frac{\rho_A s}{2 + \rho_As}, \\
+%G(1) = a + b + c = (1+s)[\rho_B - \rho_A + s \rho_B (1-\rho_A)], \\
+%G'(1) = 2a + b = s(1-\rho_A)(2(1+s)\rho_B + s\rho_A(1-\rho_B)) > 0.
+%\end{align}
+
+\begin{figure*}[htb]
 \centering
@@
 \textbf{Protected polymorphism with fluctuating transmission and selection.}
-The positive root $x^*$ of $G(x)$, see eq.~\ref{eq:xstar_periodic_fluc_k=1}.
-Dashed lines represent, from left to right, $\rho_2=\frac{\rho_1}{1+s(1-\rho_1)}$, $\rho_2=\rho_1$, and $\rho_2=\frac{(1+s)\rho_1}{1+s\rho_1}$, see eq.~\ref{eq:poly_condition_periodic_fluc}.
-Here, $\rho_1=0.5$, $s=0.1$.}
+The positive root $x^*$ of $G(x)$, see~\eqref{eq:xstar_periodic_fluc_k=1}.
+Dashed lines represent, from left to right, $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$, $\rho_B=\rho_A$, and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, see~\eqref{eq:poly_condition_periodic_fluc}.
+Here, $\rho_A=0.5$, $s=0.1$.}
 \label{fig:rho1_rho2_k=1}
@@
 
-\autoref{fig:rho1_rho2_k=1} shows the solution for $F_B(F_A(x))=x$ for $k=1$, highlighting the area of the parameter space in which a protected polymorphism exists.
-Surprisingly, the stronger selection is, the larger the difference can be that still allows a stable polymorphism.
-
-In the case of $k=1$ we can also find the frequency of $A$ at the protected polymorphism, $x^*$.
-Denote $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
-\begin{align} \label{eq:xstar_periodic_fluc_k=1}
-G(x) = ax^2+bx+c, \\
-\gamma = -\frac{1+xs}{s} \Big( 1 + s + s^2x (1-\rho) +s^2x^2(1-\rho) \Big) > 0, \\
-a = s^2 \rho_1 (1-\rho_1) (1 - \rho_2 + \rho_2/\rho_1)) > 0, \\
-b = s(1-\rho_1)(2\rho_2 - s\rho_1(1-\rho_2)), \\
-c = \rho_2 - \rho_1 - s\rho_1(1-\rho_2),
-\end{align} 
-and then $x^*$ is a solution of the quadratic $G(x)=0$, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
-The condition \eqref{eq:poly_condition_periodic_fluc} guarantees that $x^* \in (0,1)$ and is equivalent to $c<0$.% and $a+b+c>0$.
-Therefore, $\sqrt{b^2-4ac} > b$, and if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$ (\autoref{fig:rho1_rho2_k=1}).
-
-%\begin{align}
-%G(0) = c > 0  \Leftrightarrow \rho_2 > \frac{\rho_1(1+s)}{1+\rho_1s}, \\
-%G'(0) = b > 0 \Leftrightarrow \rho_2 > \frac{\rho_1 s}{2 + \rho_1s}, \\
-%G(1) = a + b + c = (1+s)[\rho_2 - \rho_1 + s \rho_2 (1-\rho_1)], \\
-%G'(1) = 2a + b = s(1-\rho_1)(2(1+s)\rho_2 + s\rho_1(1-\rho_2)) > 0.
-%\end{align}
-
-In the more general case, phenotype $A$ is favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle such that $\rho=\rho_1$ when $A$ is favored and $\rho=\rho_2$ when $B$ is favored.
-The requirements for a protected polymorphism is now
+\paragraph{Asymmetric period: $AkBl$.}
+In the more general case, phenotype $A$ is favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle such that $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
+The requirements for a protected polymorphism are now
 \begin{equation}\begin{aligned}
-1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b-\Delta \rho s/(1+s)\big)^l,\\
-1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a+\Delta \rho s)^l,
+1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b - (\rho_B - \rho_A) s/(1+s)\big)^l,\\
+1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a + (\rho_B - \rho_A) s)^l,
 \end{aligned}\end{equation}
-where $a=1+\rho_1 s>1$, $b=1-\rho_1\frac{s}{1+s}<1$, and $\Delta \rho = \rho_2 - \rho_1$.
+where $a=1+\rho_A s>1$, $b=1-\rho_A\frac{s}{1+s}<1$, and $\Delta \rho = \rho_B - \rho_A$.
 This leads to a condition similar to eq.~20 in~\citet{Ram2018}, but more complex due to the addition of $\Delta \rho$:
 \begin{equation} \label{eq:poly_condition_periodic_fluc_k_l}
-\frac{-\log{b}}{\log{\big(a+\Delta \rho s\big)}} < 
+\frac{-\log{b}}{\log{\big(a+(\rho_B - \rho_A) s\big)}} < 
 \frac{l}{k} < 
-\frac{\log{a}}{-\log{\big(b-\Delta \rho s/(1+s)\big)}}.
+\frac{\log{a}}{-\log{\big(b-(\rho_B - \rho_A) s/(1+s)\big)}}.
 \end{equation}
-Note that as $\Delta \rho$ increases, both the RHS and the LHS of \eqref{eq:poly_condition_periodic_fluc_k_l} decrease.
-Therefore, increasing the vertical transmission rate specifically when $B$ is favored ($\Delta \rho>0$) decreases the environmental period ratio $l/k$ that permits protected polymorphism; reducing the vertical transmission rate when $B$ is favored ($\Delta \rho<0$) will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
+Therefore, increasing $\rho_B$ the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$ will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
 
-% TODO plot \Delta \rho vs s and show the boundaries
+% TODO plot \rho_B - \rho_A vs s and show the boundaries
 
@@
 where $s_t$ are i.i.d (independent and identically distributed), $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$, $\rho_t$ are i.i.d, and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
-Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D$ and from Result~6 and 7 of \citet{Ram2018} we have:
+Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D)$, and from Results~6 and 7 of \citet{Ram2018} we have:
 \begin{itemize}
 \item Suppose $E[log(1+\rho_t s_t)]>0$. Then $x^*=0$ is not stochastically locally stable and in fact $P(lim_{t \to \infty} x_t=0) = 0$, i.e., fixation of $B$ almost surely does not occur.
-\item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=0$ is stochastically locally stable and 
+\item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=0$ is stochastically locally stable. 
 \item Similarly, if $E[log(1-\rho_t s_t/(1+s_t)]<0$, then $x^*=1$ is stochastically locally stable, and if $E[log(1-\rho_t s_t/(1+s_t))]>0$, then fixation of $A$ almost surely does not occur.
 \item In particular, if $E[\rho_t s_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-\rho_t s_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
-\item Note that it is not possible that both $E[log(1+\rho_t s_t)]$ and $E[log(1-\rho_t s_t/(1+s_t)]$ are negative, as their sum is positive
+\item It is not possible that $E[log(1+\rho_t s_t)]$ and $E[log(1-\rho_t s_t/(1+s_t)]$ are both negative, as their sum is positive
 \begin{multline}
@@
 E[log\big((1+\rho_t s_t)(1-\rho_t s_t/(1+s_t)\big)]= \\
-E[log\big( 1+\rho_t(1-\rho_t)(s_t^2)/(1+s_t) \big)] > 0,
+E[log\big( 1+\rho_t(1-\rho_t)s_t^2/(1+s_t) \big)] > 0,
 \end{multline}
@@
 
-We will now present a number of examples.
-First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)$) and $s_t$ is symmetric around zero, then $E[\rho_t s_t]=0$ and $E[-\rho_t s_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
+\paragraph{Examples.}
+First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)=0$) and $s_t$ is symmetric around zero, then $E[\rho_t s_t]=0$ and $E[-\rho_t s_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
 Therefore, fixation of $B$ is stochastically locally stable and fixation of $A$ almost surely does not occur.
 For example, let $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then $E[log(1+\rho_t s_t)]\approx -0.07315$ and $E[log(1-\rho_t s_t/(1+s_t))]\approx 0.2337$.
-However, note that symmetry of $s_t$ around zero provides an advantage to phenotype $B$ -- using Jensen's inequality, $E[w_A/w_B] = E[1+s_t] = 1 \le E[1/(1+s_t)] = E[w_B/w_A]$.
-Therefore, if we take $w_{A,t}, w_{B,t} \sim U(0,1)$ independently, and $s_t=(w_A-w_B)/w_B$, then neither $A$ nor $B$ has an advantage, on average (i.e.
-$E[w_A/w_B]=E[w_B/w_A]$, see \autoref{fig:beta}A), and both $E[\rho_t s_t]$ and $E[-\rho_t s_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable.
 
-Second, if $s_t$ and $\rho_t$ are not independent, we can get fixation.
-Let $\rho_t \sim \beta(1+s, 1)$, where $s_t=(w_A-w_B)/w_B$ and $w_A, w_B \sim U(0,1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$); that is, vertical transmission is more likely when $A$ is favored (i.e.
+However, note that symmetry of $s_t$ around zero provides an advantage to phenotype $B$: using Jensen's inequality, $E[w_A/w_B] = E[1+s_t] = 1 \le E[1/(1+s_t)] = E[w_B/w_A]$.
+Therefore, if we take the i.i.d fitness random variables for $A$ and $B$ to be $w_{A,t}, w_{B,t} \sim U(0,1)$, respectively, and define $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$, then neither $A$ nor $B$ has an advantage, on average (i.e.
+$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[\rho_t s_t]$ and $E[-\rho_t s_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to converge to a polymorphic distribution.
+
+Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, we can get fixation.
+Let $w_A, w_B \sim U(0,1)$, $s_t=(w_A-w_B)/w_B$ and $\rho_t \sim \beta(1+s_t, 1)$. 
+The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$); that is, vertical transmission is more likely when $A$ is favored (i.e.
 $s_t>0$) and oblique transmission is more  likely when $B$ is favored (i.e.
@@
 Then $E[log(1+\rho_t s_t)] >0$ and $B$ almost surely doesn't fix, and $E[log(1-\rho_t s_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
-The opposite occurs if $\rho_t \sim \beta(1, 1+s)$ and the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -4$).
+The opposite occurs if $\rho_t \sim \beta(1, 1+s_t)$ and the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -4$).
 In that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.
@@
 
-To include the effect of random drift due to finite population size in the above deterministic model, we follow \citet{Ram2018} and develop a diffusion equation approximation.
-In~\citet{Ram2018} only selection fluctuated via $s_t$, but here we also transmission to fluctuate via $\rho_t$.
+To include the effects of random genetic drift due to finite population size in the above deterministic model, we follow \citet{Ram2018} and develop a diffusion equation approximation.
+In~\citet{Ram2018} only selection fluctuated via $s_t$, but here we also let transmission to fluctuate via $\rho_t$.
 
 We find a result similar to Result 11 from~\citet{Ram2018}:
-The mean $\mu(x)$ and variance $\sigma^2(x)$ of the change in the frequency of $A$ in one generation for the diffusion approximation in the case of a cycling environment $AkBl$, where $k+l=n$, are
+The mean $\mu(x)$ and variance $\sigma^2(x)$ of the change in the frequency of $A$ in the case of a cycling environment $AkBl$, where $k+l=n$, are
 \begin{equation}
-\mu(x) = S_n x(1-x), \quad \text{and} \quad \sigma^2(x) = n x (1-x)
+\mu(x) = S_n x(1-x)
+\quad \text{and} \quad
+\sigma^2(x) = n x (1-x),
 \end{equation}
-where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$, which is a weighted average of the selection coefficients in favor of phenotype $A$, weighted by the vertical transmission rate.
+where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$, which the sum of selection coefficients in favor of phenotype $A$ over $n$ generations weighted by the vertical transmission rate $\rho_t$.
 
@@
 Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly. 
-Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that transmission fluctuations constrain the maintenance of phenotypic polymorphism.
+Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ % TODO
 %Specifically, if fluctuations are not small then a single phenotype is likely to fix in the population~(\autoref{eq:poly_condition_periodic_fluc}, \autoref{fig:rho1_rho2_k=1}, \autoref{fig:rho1_rho2_stoch_p}), and
 
-% Face book rural internet
+One caveat of our model is the use of the ``phenotypic gambit'' -- the assumption that the transmission mode itself is strictly vertically transmitted.
+Although there is evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
+Indeed, it has been demonstrated that guppy fish are more likely to learn from others if previous social experiences provided benefits~\citep{Leris2016}.
+It is also possible that the transmission mode is regulated.
+For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
+Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only some individuals do so.
+
+An extension of our model could incorporate more sophisticated oblique transmission schemes.
+For example, conformity -- preference for learning a frequent phenotype -- has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
+We suggest that the specific mode of oblique transmission can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions.
+Additional work will be required to understand how such fluctuations affect the population dynamics.
+
+% Facebook rural internet 
 % National service - Israel, Singapore

### 2018-03-26 ms/fluctuating_oblique.tex
better figure for k=1

@@
 So, if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$.
-\autoref{fig:rho1_rho2_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
+\autoref{fig:rho1_rho2_phase_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
+The figure demonstrates that the stronger selection is (x-axis), the greater the fluctuations in $\rho$ can be (y-axis) while still maintaining a polymorphic population (area between the dashed lines).
 
@@
 \centering
-\includegraphics[width=0.65\linewidth]{../figures/{rho1_rho2_k=1}.pdf}
+\includegraphics[width=0.65\linewidth]{../figures/{rho1_rho2_phase_k=1}.pdf}
 \caption{
-\textbf{Protected polymorphism with fluctuating transmission and selection.}
-The positive root $x^*$ of $G(x)$, see~\eqref{eq:xstar_periodic_fluc_k=1}.
-Dashed lines represent, from left to right, $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$, $\rho_B=\rho_A$, and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, see~\eqref{eq:poly_condition_periodic_fluc}.
-Here, $\rho_A=0.5$, $s=0.1$.}
-\label{fig:rho1_rho2_k=1}
+\textbf{Protected polymorphism.}
+The stable equilibrium of the frequency of phenotype $A$ (\eqref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$, x-axis) and difference between vertical transmission rates ($\rho_B-\rho_A$, y-axis) when both selection and transmission fluctuate every generation ($k=1$).
+Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ that allow a protected polymorphism (\eqref{eq:poly_condition_periodic_fluc}).
+Here, $\rho_A=0.5$.}
+\label{fig:rho1_rho2_phase_k=1}
 \end{figure*}
@@
 Therefore, increasing $\rho_B$ the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$ will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
-
-% TODO plot \rho_B - \rho_A vs s and show the boundaries
 
@@
 \textbf{Stochastic local stability.}
-Here, $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$, with $s=0.05$ and $p=0.505$, a combination that allows a stochastic polymorphism with a fixed vertical transmission rate $\rho=0.1$ in~\citet[Fig.~2]{Ram2018}.}
+Here, $s_t=0.05$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-0.05$ and $\rho_t=\rho_2$ with probability $1-p=0.495$, a combination that allows a stochastic polymorphism with a fixed vertical transmission rate $\rho=0.1$ in~\citet[Fig.~2]{Ram2018}.}
 \label{fig:rho1_rho2_stoch_p}

### 2018-04-24 ms/fluctuating_oblique.tex
TODOs
revised; fixed MWF's comments

@@
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size~\citep{Fischer2015} or density~\citep{Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size or density~\citep{Fischer2015,Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission will fluctuate over time.
 For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases.
@@
 \textbf{Cultural transmission with mixed vertical and oblique transmission.}
-When a newborn matures, she will copy her phenotype -- color -- from her mother with probability $\rho$, therefore becoming blue, or from some other female with probability $1-\rho$, in which case her color will depend on the frequency of blue and red adult females.}
+When a newborn matures, she copies her phenotype -- color -- from her mother with probability $\rho$, therefore becoming blue, or from some other female with probability $1-\rho$, in which case her color will depend on the frequency of blue and red adult females.}
 \label{fig:transmission}
@@
 
-Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$ when $B$ is favored the transmission rate is $\rho_B$.
-The change in the frequency of $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively, where
+Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$, and when $B$ is favored the transmission rate is $\rho_B$.
+The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively,
 \begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
 F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, \\
-F_B(x) = \rho_B \frac{1}{1+s-sx'} x' + (1-\rho_B)x'.
+F_B(x) = \rho_B \frac{1}{1+s-sx} x' + (1-\rho_B)x.
 \end{aligned}\end{equation}
 
-\paragraph{Symmetric periods: $AkBk$.}
+\paragraph{Symmetric periods -- $AkBk$.}
 We first consider environments which periodically fluctuate between favoring $A$ and $B$ every $k$ generations.
@@
 \end{aligned}\end{equation}
-and if both of these conditions are not met then there exists a protected polymorphism.
+and if both of these conditions are not met then there exists a \emph{protected polymorphism}.
 So for a protected polymorphism we require
@@
 
-\paragraph{Period of one: $A1B1$.}
+\paragraph{Period of one -- $A1B1$.}
 In the case of $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
@@
 % $a+b+c>0$.
-and then $x^*$ is a solution of the quadratic $G(x)=0$. %, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
+then $x^*$ is a solution of the quadratic $G(x)=0$. %, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
 The condition \eqref{eq:poly_condition_periodic_fluc}, which guarantees that $0 < x^* < 1$, is equivalent to $c<0$, and therefore it also guarantees that $\sqrt{b^2-4ac} > b$.
-So, if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$.
+So, if $0<s<1$ (such that $b$ is guaranteed to be positive given LHS of eq.~\ref{eq:poly_condition_periodic_fluc}), then 
+\begin{equation}
+x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}.
+\end{equation}
 \autoref{fig:rho1_rho2_phase_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
@@
 \textbf{Protected polymorphism.}
-The stable equilibrium of the frequency of phenotype $A$ (\eqref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$, x-axis) and difference between vertical transmission rates ($\rho_B-\rho_A$, y-axis) when both selection and transmission fluctuate every generation ($k=1$).
-Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ that allow a protected polymorphism (\eqref{eq:poly_condition_periodic_fluc}).
+The stable equilibrium of the frequency of phenotype $A$ \eqref{eq:xstar_periodic_fluc_k=1} for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
+Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ that allow a protected polymorphism \eqref{eq:poly_condition_periodic_fluc}.
 Here, $\rho_A=0.5$.}
@@
 
-\paragraph{Asymmetric period: $AkBl$.}
+\paragraph{Asymmetric period -- $AkBl$.}
 In the more general case, phenotype $A$ is favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle such that $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
@@
 \begin{equation} \label{eq:poly_condition_periodic_fluc_k_l}
-\frac{-\log{b}}{\log{\big(a+(\rho_B - \rho_A) s\big)}} < 
+\frac{-\log{b}}{\log{\big(a + s \Delta \rho \big)}} < 
 \frac{l}{k} < 
-\frac{\log{a}}{-\log{\big(b-(\rho_B - \rho_A) s/(1+s)\big)}}.
+\frac{\log{a}}{-\log{\big(b - s \Delta \rho/(1+s)\big)}}.
 \end{equation}
-Therefore, increasing $\rho_B$ the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$ will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
+Therefore, increasing the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$  will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
 
@@
 Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, we can get fixation.
-Let $w_A, w_B \sim U(0,1)$, $s_t=(w_A-w_B)/w_B$ and $\rho_t \sim \beta(1+s_t, 1)$. 
-The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$); that is, vertical transmission is more likely when $A$ is favored (i.e.
+Let $w_{A,t}, w_{B,t} \sim U(0,1)$, $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$ and $\rho_t \sim Beta(1+s_t, 1)$ (a beta distribution with parameters $1+s_t$ and $1$). 
+The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$; estimated by averaging $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
 $s_t>0$) and oblique transmission is more  likely when $B$ is favored (i.e.
@@
 \textbf{Covariance of selection and transmission.}
-\textbf{(A)} Histogram of $w_A/w_B$ where $w_A$ and $w_B$ are identically and independently distributed uniform random variables $U(0,1)$.
-\textbf{(B)} Histogram of $s_t = (w_A-w_B)/w_B$.
+\textbf{(A)} Histogram of $w_{A,t}/w_{B,t}$ where $w_{A,t}$ and $w_{B,t}$ are identically and independently distributed uniform random variables $U(0,1)$.
+\textbf{(B)} Histogram of $s_t = (w_{A,t}-w_{B,t})/w_{B,t}$.
 \textbf{(C)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
@@
 \textbf{Stochastic local stability.}
-Here, $s_t=0.05$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-0.05$ and $\rho_t=\rho_2$ with probability $1-p=0.495$, a combination that allows a stochastic polymorphism with a fixed vertical transmission rate $\rho=0.1$ in~\citet[Fig.~2]{Ram2018}.}
+Here, $s_t=0.05$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-0.05$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
+The diagonal represents the case of no transmission fluctuations; \citet[Fig.~2]{Ram2018} demonstrated that with a constant transmission rate $\rho=0.1$ and the above distribution of $s_t$, both fixations are not stochastically stable.
+}
 \label{fig:rho1_rho2_stoch_p}
 \end{figure*}
+
+% TODO full correlation between rho and s?
 
@@
 We find a result similar to Result 11 from~\citet{Ram2018}:
-The mean $\mu(x)$ and variance $\sigma^2(x)$ of the change in the frequency of $A$ in the case of a cycling environment $AkBl$, where $k+l=n$, are
-\begin{equation}
+The mean $\mu(x)$ and variance $\sigma^2(x)$ of the change in the frequency $x$ of phenotype $A$ in the case of a cycling environment $AkBl$, where $k+l=n$, are
+\begin{equation} \label{eq:drift_diffusion_terms}
 \mu(x) = S_n x(1-x)
@@
 \end{equation}
-where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$, which the sum of selection coefficients in favor of phenotype $A$ over $n$ generations weighted by the vertical transmission rate $\rho_t$.
+where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$.
+Furthermore, combining \eqref{eq:drift_diffusion_terms} with eqs.~46-47 from \citep{Ram2018}, we find that the fixation probability of phenotype $A$, starting with frequency $x$ of phenotype $A$, is
+\begin{equation}
+u(x) = \frac{1 - e^{-2 \frac{S_n}{n} x}}{1 - e^{-2 \frac{S_n}{n}}}.
+\end{equation}
+From Result 10 in \citep{Ram2018}, $u(x)$ is monotone increasing in $S_n/n$ which is the average selection coefficient of $A$ weighted by the vertical transmission rates $\rho_t$.
+Therefore, if $s_t$ and $\rho_t$ are positively (negatively) correlated, $S_n/n$ increases (decreases), and the fixation probability $u(x)$ increases (decreases). 
+This occurs because selection only affects those individuals that transmit their phenotype to their own offspring (i.e.
+vertically), and a fraction $1-\rho_t$ of the population individuals is effectively masked (for better or worse) from selection at each generation.
 

### 2018-04-30 ms/fluctuating_oblique.tex
migration model

@@
 Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from non-parental adults with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time~(\autoref{fig:transmission}).
-We find that $\ldots$ %TODO
+We find that $\ldots$ (to be continued)  %TODO
 We suggest that cultural evolution can be affected by environmental changes via both selection and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
@@
 
+% Fluctuations in space
+\subsection*{Fluctuations in space}
+
+We now describe a model in which fluctuations in selection and transmission occur in space, rather then time; that is, we model a population divided to two demes.
+Selection (e.g.
+reproduction) and transmission (e.g.
+learning, development) occur within the demes, followed by migration of sub-adults -- individuals that already acquired their phenotype but have yet to reproduce.
+The frequency of phenotype $A$ in deme $j$ is marked by $x_j$, and therefore after selection and transmission the frequencies are 
+\begin{equation} \label{eq:migration_model_selection_transmission}
+x_j' = \rho_j \frac{w_j}{\overline{w}_j} x_j + (1-\rho_j) x_j,
+\end{equation}
+where $w_j$ is the fitness of phenotype $A$ in deme $j$ relative to the fitness of phenotype $B$; $\overline{w}_j=w_j x_j + (1-x_j)$ is the population mean fitness in deme $j$; and $\rho_j$ is the vertical transmission rate in deme $j$.
+
+Following migration, the frequencies of $A$ in the two demes are
+\begin{equation} \label{eq:migration_model}
+\begin{aligned}
+x_1'' = (1-m_1) x_1' + m_2 x_2', \\
+x_2'' = m_1 x_1' + (1-m_2) x_2',
+\end{aligned}
+\end{equation}
+where $m_j$ is the migration rate out of deme $j$.
+
+\autoref{fig:migration} shows the stable frequencies of phenotype $A$ (i.e.
+the frequencies that solve $x_1=x_1''$ and $x_2=x_2''$) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar intensities (see figure legend).
+Notably, without oblique transmission (left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
+With some oblique transmission, but equal in both demes, results are similar (not shown).
+However, if oblique transmission is stronger in deme 2 compared to deme 1 (middle and right columns), the stable frequency of $A$ increases in both demes.
+Therefore, the mean fitness in deme 1 decreases to a lesser extend than in deme 2, and even increases when the migration rate is high enough.
+Furthermore, the polymorphism that is protected when transmission rates are equal disappears when transmission rates differ and migration is strong enough, and phenotype $A$ fixes (see right side of right column).
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/migration.pdf}
+\caption{
+\textbf{Oblique transmission and migration.} 
+The figure shows $x^*_i$ the stable frequency of $A$ (top row) and $\bar{w}^*_i$ population mean fitness (bottom row) in the two demes.
+Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
+The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.8$, and $0.25$, in the left, middle, and right columns, respectively.
+Migration rate $m$ is on the x-axis.
+}
+\label{fig:migration}
+\end{figure*}
+
 % Discussion
@@
 Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly. 
-Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ % TODO
+Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ (to be continued) % TODO
 %Specifically, if fluctuations are not small then a single phenotype is likely to fix in the population~(\autoref{eq:poly_condition_periodic_fluc}, \autoref{fig:rho1_rho2_k=1}, \autoref{fig:rho1_rho2_stoch_p}), and

### 2018-05-06 ms/genetic_cultural_fitness.tex
edit

@@
-\documentclass[12pt]{article} % Default font size of the document, change to 10pt to fit more text
-
+\documentclass[14pt]{extarticle}
 \usepackage{geometry}
@@
 }
+
 \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
@@
 $x' = \rho x \frac{w_A}{\bar w} + (1-\rho) x$,
-which is eq.~2 from~\citet{Ram2018}.
-So setting $v_A = v_B = \rho$ or even to so other parameter $v_A = v_B = \tau$ will not provide new insight.
+which is eq.~2 from~\citet{Ram2018}, so we hereafter assume that $v_A \ne v_B$.
 
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Constant environment}
+\subsection*{Equilibrium.}
 
 In a constant environment we will assume, without loss of generality, that $w_A > w_B$.
-We will now find the equilibrium phenotype frequencies.
-Setting $x'=x$ in~\eqref{eq:recurrence} and rearranging, we get a cubic polynomial $ax^3+bx^2+cx$.
+To find the phenotype equilibrium frequencies,
+set $x'=x$ in~\eqref{eq:recurrence} and rearrange to get a cubic polynomial $a_3x^3+a_2x^2+a_1x$.
 %\begin{equation}
@@
 %\end{equation}
-Two roots are easy to verify: $x=0 \Rightarrow x'=0$ and $x=1 \Rightarrow x'=\rho \frac{w_A}{w_A} + (1-\rho) \frac{v_A}{v_A} = 1$. 
-Therefore, the cubic polynomial is $x (1-x) (x^*-x)$ where
-\begin{equation}
-%\begin{aligned}
-x^* = 
+Two roots are easy to verify: $x=0$ and $x=1$
+% \Rightarrow x'=0$ and $x=1 \Rightarrow x'=\rho \frac{w_A}{w_A} + (1-\rho) \frac{v_A}{v_A} = 1$. 
+Therefore, the cubic polynomial is $x (1-x) (x^*-x)$ and the three roots are
+\begin{align} \label{eq:roots}
+&x = 0 \\
+&x = 1 \\
+&x^* = 
 %\frac{w_B v_B - (1-\rho) w_B v_A - \rho w_A v_B}{w_A v_A - w_A v_B - w_B v_A + w_B v_B} = \\
-\rho \frac{w_B v_A - w_A v_B}{(w_A - w_B)(v_A - v_B)} - \frac{w_B}{w_A-w_B}.
-%\end{aligned}
-\end{equation}
-%The conditions for $0 < x^* < 1$ are:
-%\begin{equation}
-%\begin{aligned}
-%\frac{w_B (v_A - v_B)}{w_B v_A - w_A v_B}  < 
-%\rho < 
-%\frac{w_A (v_A - v_B)}{w_B v_A - w_A v_B} \\
-%\end{aligned}
-%\end{equation}
-
-Now, $x=0$ is unstable if
-\begin{equation}
-%\frac{x'}{x} \eqsim 
-\rho \frac{w_A}{w_B} + (1-\rho) \frac{v_A}{v_B} > 1,
-\end{equation}
-and $x=1$ is unstable if
-\begin{equation}
-%\frac{(1-x)'}{1-x} \eqsim 
-\rho \frac{w_B}{w_A} + (1-\rho) \frac{v_B}{v_A} > 1.
+%\rho \frac{w_B v_A - w_A v_B}{(w_A - w_B)(v_A - v_B)} - \frac{w_B}{w_A-w_B} = 
+\frac{\rho s_w + (1-\rho) s_v}{s_w s_v},
+\end{align}
+where $1+s_w=\frac{w_A}{w_B}>1$ and $1+s_v=\frac{v_A}{v_B}>0$ (note this assumes that $s_w, s_v > -1$).
+%Note that $x^*$ might be larger than 1 or less than 0, in which case it is not an equilibrium frequency.
+%Note that $x^*<1$ iff $\rho \frac{w_A}{w_A-w_B} > (1-\rho) \frac{v_A}{v_B-v_A}$, 
+%and $x^*>0$ iff $\frac{\rho}{1-\rho} > |\frac{s_v}{s_w}|$.
+For $x^*$ to be a valid equilibrium frequency, it must be between 0 and 1, which implies the following conditions
+\begin{equation} \label{eq:xstar_conditions}
+\begin{cases}
+0 < \rho < \frac{-s_v}{s_w - s_v},& \quad \text{if}\; s_w > 0 > s_v > -1,\\
+0 < \rho < 1,& \quad \text{if}\; s_w > s_v > 0 \\
+0 < \rho < \frac{s_v}{s_v - s_w},& \quad \text{if}\; s_v > s_w > 0
+\end{cases}
 \end{equation}
 
-So both $x=0$ and $x=1$ are unstable if
-\begin{equation} \label{eq:protected_poly_const}
-w_B(v_B - v_A) < \rho (w_A v_B - w_B v_A) < w_A(v_B - v_A),
+Now, $x=0$ is locally stable if 
+\begin{equation}
+\frac{x'}{x} \eqsim 
+\rho (1+s_w) + (1-\rho) (1+s_v)
 \end{equation}
-in which case there exists a protected phenotypic polymorphism $0 < x^* < 1$.
+is less than 1 and unstable if it is larger than 1.
+Similarly, $x=1$ is locally stable if 
+\begin{equation}
+\frac{(1-x)'}{1-x} \eqsim 
+\rho \frac{1}{1+s_w} + (1-\rho) \frac{1}{1+s_v}
+\end{equation}
+is less than 1 and unstable if it is larger than 1.
+This implies the following conditions for the instability of both $x=0$ and $x=1$:
+\begin{equation} \label{eq:fixation_instability_conditions}
+\begin{cases}
+\frac{-s_v}{s_w - s_v} < \rho < (1+s_w) \frac{-s_v}{s_w - s_v},& \quad \text{if}\; s_w > 0 > s_v > -1,\\
+\text{n.a.},& \quad \text{if}\; s_w > s_v > 0 \\
+\text{n.a.},& \quad \text{if}\; s_v > s_w > 0
+\end{cases}
+\end{equation}
 
-There are three possibilities (disregarding equalities):
-\begin{enumerate}
-\item If $v_A > v_B$ and $\frac{w_A}{w_B} > \frac{v_A}{v_B}$ then condition~\eqref{eq:protected_poly_const} cannot apply as it requires $\rho<0$.
-\item If $v_A > v_B$ and $\frac{w_A}{w_B} < \frac{v_A}{v_B}$ then condition~\eqref{eq:protected_poly_const} cannot apply as it requires $\rho>1$.
-\item If $v_B > v_A$ then $w_A v_B - w_B v_A > 0$, $w_B(v_B - v_A)<w_A v_B - w_B v_A$, and $w_A v_B - w_B v_A > w_A(v_B - v_A)$ and condition~\eqref{eq:protected_poly_const} applies, ensuring a protected phenotypic polymorphism, if
+Therefore, although a polymorphism fixed point exists~\eqref{eq:xstar_conditions}, it cannot be stable, because whenever it exists at least one of the fixations (in either $A$ or $B$) is stable~\eqref{eq:fixation_instability_conditions}.
+
+%TODO - when is x=1 stable? When is x=0 stable?
+
+\subsection*{Finite population.}
+To model the effect of random drift due to a finite population size $N$, we use a Wright-Fisher model.
+Let $X_t$ denote the number of $A$ individuals at generation $t$ ($X_t \in \{0, 1, \ldots, N\}$) so that $x=X_t/N$.
+$x'$ still represents the frequency of $A$ in the next generation as in~\eqref{eq:recurrence}.
+According to the Wright-Fisher model~\citep{Ewens2004}, the number of $A$ individuals in generation $(t+1)$ is binomially distributed:
 \begin{equation}
-\frac{w_B(v_B - v_A)}{w_A v_B - w_B v_A} < \rho < \frac{w_A(v_B - v_A)}{w_A v_B - w_B v_A},
+X_{t+1} \sim Bin(N, x').
 \end{equation}
-which can also be written as 
+The stochastic process $X_t$ has two absorbing states $X_t=0$ and $X_t=N$, corresponding to the two fixations in $A$ and $B$, respectively.
+When one of the fixations is globally stable in the deterministic model~\eqref{eq:recurrence}, we are interested in the fixation probability $u(x)$ and time $T(x)$ starting from an arbitrary frequency $x$ and towards the stable absorbing state (either $x=0$ or $x=1$).
+When a protected polymorphism exists, we are interested in the probabilities $u_0(x)$ and $u_1(x)$ for fixation towards $x=0$ or $x=1$, respectively, and the time for loss of polymorphism $T(x)$, or the fixation in either $x=0$ or $x=1$.
+
+To proceed, we use a diffusion approximation, which requires that the differential selection (genetic or cultural) is small, or formally~\citep{Karlin1974a}
 \begin{equation}
-\frac{-s_v}{s_w-s_v} 
-< \rho < 
-(1+s_w) \cdot \frac{-s_v}{s_w-s_v},
+w_A - w_B \eqsim \frac{\eta_w}{N}, \quad v_A - v_B \eqsim \frac{\eta_v}{N},
 \end{equation}
-where $s_w=\frac{w_A}{w_B}-1>0$ and $s_v=\frac{v_A}{v_B}-1<0$.
-\end{enumerate}
+up to terms of order smaller than $1/N$.
+Therefore, the infinitesimal mean displacement is
+\begin{equation}
+\mu(x)=\frac{1}{N}\E[X_{t+1}-X_t]=x(1-x) \eta
+\end{equation}
+where $\eta = \rho \eta_w + (1-\rho) \eta_v$.
+The infinitesimal variance $\sigma^2(x)$ is
+\begin{equation}
+\sigma^2(x)=\frac{1}{N^2}Var[X_{t+1}-X_t]=x(1-x),
+\end{equation}
+up to terms of order smaller than $1/N$.
+
+%We start with the case that $x=1$ is a stable equilibrium.
+%, that is, either $\eta>0$ and $\eta_v>0$ or $\eta_w>0$, $\eta_v<0$, and $\rho > (1+s_w)\frac{-s_v}{s_w-s_v}$.
+Therefore, the probability for fixation of $A$ (i.e. at $x=1$), starting with frequency $x$, is
+\begin{equation} \label{eq:uA(x)}
+u_A(x) = \frac{1-e^{-2 \eta x}}{1-e^{-2\eta}},
+\end{equation}
+and $u_B(x) = 1-u_A(x)$ is the probability for fixation of $B$ (i.e. at $x=0$).
+The time to fixation at either $A$ or $B$ (starting with frequency $x$) is 
+\begin{equation} \label{eq:T(x)}
+T(x) = 
+\frac{1-u(x)}{\eta} \int_0^x{\frac{e^{2 \eta z} - 1}{z(1-z)} dz} + 
+\frac{u(x)}{\eta} \int_x^1{\frac{1-e^{-2 \eta (1-z)}}{z(1-z)} dz}
+\end{equation}
+where $u_A(x)$ is given in~\eqref{eq:uA(x)}, and $T(x)$ is measured in units of $1/N$.
+
+For the fixation probability $u(x)$, we have the following result.
+First, note that $u(x)$ is monotone increasing in $\eta$~\cite[see][Appendix E]{Ram2018}. 
+Next, $\frac{d\eta}{d\rho} = \eta_w - \eta_v$ and therefore $u(x)$ is monotone increasing (decreasing) in $\rho$ when $\eta_w > \eta_v$ ($\eta_w < \eta_v$). 
+That is, if the genetic fitness differential is bigger than the cultural fitness differential, then the probability of fixation increases with the vertical transmission rate; if the cultural fitness differential is bigger, than the fixation probability decreases with the vertical transmission rate.
+
+We are specifically interested in the probability for stochastic fixation of $A$ (rather than $B$) starting at the (unstable) equilibrium polymorphism $x=x^*$ in~\eqref{eq:roots} (i.e.
+assuming $\eta_w>0$, $\eta_v<0$, and~\eqref{eq:fixation_instability_conditions}).
+
+First, assuming $w_B \eqsim v_B \eqsim 1$ we have $\eta_w \eqsim s_w$ and $\eta_v \eqsim s_v$, so $\eta<0$ and $x^*=\eta / \eta_v \eta_w$~(from \eqref{eq:roots}) and we have
+\begin{equation} \label{eq:uA(x*)}
+u_A(x^*) = \frac{1-e^{-2 \eta^2/\eta_v \eta_w}}{1-e^{-2\eta}},
+\end{equation}
+Similarly, the time for loss of polymorphism when starting at the stable polymorphism $x=x^*$ under the same assumptions is
+\begin{equation} \label{eq:T(x*)}
+T(x^*) = 
+\frac{1-u(x^*)}{\eta} \int_0^{x^*}{\frac{e^{2 \eta z} - 1}{z(1-z)} dz} + 
+\frac{u(x^*)}{\eta} \int_{x^*}^1{\frac{1-e^{-2 \eta (1-z)}}{z(1-z)} dz}
+\end{equation}
 
@@
 \bibliographystyle{unsrtnat}
-%\bibliography{/Users/yoavram/Documents/library}
-\bibliography{genetic_cultural_fitness}
+\bibliography{/Users/yoavram/Documents/library}
+%\bibliography{genetic_cultural_fitness}
 %%%%%%%%%%%%%%%%%%%

### 2018-05-13 ms/ms.tex
rename
analytic results for migration model

--- 2018-04-30 ms/fluctuating_oblique.tex
@@
-\documentclass[12pt]{extarticle} %twocolumn
+\documentclass[11pt]{extarticle}
 \usepackage{geometry}
@@
 \begin{aligned}
-x_1'' = (1-m_1) x_1' + m_2 x_2', \\
-x_2'' = m_1 x_1' + (1-m_2) x_2',
+x_1'' &= (1-m_1) x_1' + m_2 x_2', \\
+x_2'' &= m_1 x_1' + (1-m_2) x_2',
 \end{aligned}
 \end{equation}
-where $m_j$ is the migration rate out of deme $j$.
+where $0 \le m_j \le 1/2$ is the migration rate out of deme $j$.
 
-\autoref{fig:migration} shows the stable frequencies of phenotype $A$ (i.e.
-the frequencies that solve $x_1=x_1''$ and $x_2=x_2''$) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar intensities (see figure legend).
-Notably, without oblique transmission (left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
+For the simplified case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} and symmetric selection $w_1=1/w_2=w>1$, the equilibrium $(x_1^*, x_2^*)$ solves
+\begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
+x_1 &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big), \\
+x_2 &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big).
+\end{aligned}
+\end{equation}
+
+With only oblique transmission ($\rho_1=\rho_2=0$), the solutions to \eqref{eq:migration_model_unconditional_symmetric} are the neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
+With only vertical transmission ($\rho_1=\rho_2=1$), the solutions to \eqref{eq:migration_model_unconditional_symmetric} are $(0,0)$, $(1,1)$ and the polymorphic solution
+\begin{equation}
+\begin{aligned} \label{eq:migration_model_unconditional_symmetric_solution}
+x_1^* &= \frac{w-1-m(w+1) + \sqrt{\Delta}}{2(w-1)}, \\
+x_2^* &= 1-x_1^*,
+\end{aligned}
+\end{equation}
+where $\Delta = m^2(w+1)^2+(1-2m)(w-1)^2$.
+Examining the stability of these equilibria, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $a \lambda^2 + b \lambda + c$ with
+\begin{equation}
+\begin{aligned}
+a &= 1, \quad
+b &= -\frac{(1-m)(w^2+1)}{w}, \quad
+c &=  1-2m.
+\end{aligned}
+\end{equation}
+Since $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is positive, both equilibria are unstable, and the protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_solution}.
+
+With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), the stability of $(0,0)$ is determined by a characteristic polynomial with coefficients
+\begin{equation}
+\begin{aligned}
+a &= 1, \quad
+b &= -\frac{(1-m)(w+1-\rho+\rho/w)}{w}, \quad
+c &=  (1-2m)(w(1-\rho)+\rho).
+\end{aligned}
+\end{equation}
+Again, $a>0$ and $c>0$   and stability can be analyzed by the sign of $a+b+c$.
+For $\rho=1$ we already have $a+b+c<0$.
+For $\rho=0$ we have $a+b+c=-m(w-1)<0$.
+Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any valid $\rho$, the leading eigenvalue is positive, and $(0,0)$ is unstable.
+Regarding the stability of $(1,1)$, which is determined by a characteristic polynomial  with coefficients
+\begin{equation}
+\begin{aligned}
+a &= 1, \quad
+b &= -\frac{1-m}{w}(1 + w(1-\rho) +w^2 \rho), \quad
+c &= \frac{1-2m}{w} (w\rho+1-\rho).
+\end{aligned}
+\end{equation}
+Once more, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
+However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur at $(0,1)$ and the fixation is stable.
+We can also determine the rate $\hat \rho$ for which $a+b+c=0$, such that for
+\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
+\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)}
+\end{equation}
+the fixation $(1,1)$ is stable if and only if $2a+b>0$.
+We already found that when $\rho=0$ we have $2a+b>0$.
+When $\rho = \hat \rho$, we have $2a+b=\frac{m(w+1)+w-1}{w(1-m)(w-1)}>0$.
+Since $2a+b$ is linear in $\rho$, we find that $2a+b>0$ for $\hat \rho \le \rho \le 0$ and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_rho} guarantees the stable fixation of $A$.
+
+A similar condition can be found for $m$: the value of $a+b+c$ at $m=0$ and $m=1$ is $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when
+\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
+m > \hat m = \frac{\rho (w-1)}{\rho (w-1) +1-\rho}.
+\end{equation}
+In addition, $2a+b$ is linear in $m$ and equal to 2 when $m=1$; it is equal to zero when $m=(w-1)(w\rho-1)/(w(w-1)\rho+w+1) < \hat m$, and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_m} is sufficient for stability of fixation of $A$.
+
+\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$ (i.e.
+the frequencies that solve \eqref{eq:migration_model_unconditional_symmetric_solution}) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar intensities (see figure legend).
+Notably, without oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
 With some oblique transmission, but equal in both demes, results are similar (not shown).
-However, if oblique transmission is stronger in deme 2 compared to deme 1 (middle and right columns), the stable frequency of $A$ increases in both demes.
-Therefore, the mean fitness in deme 1 decreases to a lesser extend than in deme 2, and even increases when the migration rate is high enough.
-Furthermore, the polymorphism that is protected when transmission rates are equal disappears when transmission rates differ and migration is strong enough, and phenotype $A$ fixes (see right side of right column).
+However, if oblique transmission is stronger in deme 2 compared to deme 1 (\autoref{fig:migration_m}, middle and right columns), the stable frequency of $A$ increases in both demes.
+Therefore, the mean fitness in deme 1 decreases to a lesser extent than in deme 2, and even increases when the migration rate is high enough.
+
+The polymorphism $(x_1^*, x_2^*)$ is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, and phenotype $A$ fixes. 
+For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the stronger migration is, the less oblique transmission is required to fix $A$, see shaded areas in \autoref{fig:migration_rho}.
+Similarly, with enough migration ($m > \hat m$), phenotype $A$ fixes, and the more oblique transmission in deme 2, the less migration is needed to fix $A$; see shaded area in \autoref{fig:migration_m}.
 
@@
 \centering
-\includegraphics[width=0.75\linewidth]{../figures/migration.pdf}
+\includegraphics[width=0.75\linewidth]{../figures/migration_rho.pdf}
 \caption{
-\textbf{Oblique transmission and migration.} 
+\textbf{Oblique transmission and migration: effect of transmission.} 
 The figure shows $x^*_i$ the stable frequency of $A$ (top row) and $\bar{w}^*_i$ population mean fitness (bottom row) in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
-The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.8$, and $0.25$, in the left, middle, and right columns, respectively.
+The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2$ (x-axis) in deme 2.
+Migration rate $m$ is 0.05, 0.1, or 0.25 I the left, middle, and right columns, respectively.
+The shaded area denotes stables fixation of phenotype $A$ according to~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
+}
+\label{fig:migration_rho}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/migration_m.pdf}
+\caption{
+\textbf{Oblique transmission and migration: effect of migration.} 
+The figure shows $x^*_i$ the stable frequency of $A$ (top row) and $\bar{w}^*_i$ population mean fitness (bottom row) in the two demes.
+Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
+The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.4$, and $0.2$, in the left, middle, and right columns, respectively, in deme 2.
 Migration rate $m$ is on the x-axis.
+The shaded area denotes stables fixation of phenotype $A$ according to~\eqref{eq:migration_model_unconditional_symmetric_condition_m}.
 }
-\label{fig:migration}
+\label{fig:migration_m}
 \end{figure*}
@@
 \bibliographystyle{agsm}
-%\bibliography{/Users/yoavram/Documents/library}
-\bibliography{fluctuating_oblique}
+\bibliography{/Users/yoavram/Documents/library}
+%\bibliography{ms}
 

### 2018-05-13 ms/ms.tex
revised migration results description

@@
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size or density~\citep{Fischer2015,Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission will fluctuate over time.
+Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size or density~\citep{Fischer2015,Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission will fluctuate over time and/or space.
 For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases.
@@
 
-Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from non-parental adults with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time~(\autoref{fig:transmission}).
+Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from non-parental adults with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space~(\autoref{fig:transmission}).
 We find that $\ldots$ (to be continued)  %TODO
-We suggest that cultural evolution can be affected by environmental changes via both selection and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
+We suggest that cultural evolution can be affected by environmental heterogeneity via both selection and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
 
@@
 
-For the simplified case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} and symmetric selection $w_1=1/w_2=w>1$, the equilibrium $(x_1^*, x_2^*)$ solves
+\paragraph{Unconditional migration and symmetric selection.}
+
+For the simplified case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} and symmetric selection $w_1=1/w_2=w>1$, the model is given by
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
-x_1 &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big), \\
-x_2 &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big).
+x_1'' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big), \\
+x_2'' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big).
 \end{aligned}
@@
 
-With only oblique transmission ($\rho_1=\rho_2=0$), the solutions to \eqref{eq:migration_model_unconditional_symmetric} are the neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
-With only vertical transmission ($\rho_1=\rho_2=1$), the solutions to \eqref{eq:migration_model_unconditional_symmetric} are $(0,0)$, $(1,1)$ and the polymorphic solution
+we have the following results:
+\begin{itemize}
+\item With only oblique transmission ($\rho_1=\rho_2=0$), the only equilibria are the neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
+\item With only vertical transmission ($\rho_1=\rho_2=1$), the fixation equilibria  $(0,0)$, $(1,1)$ are unstable and there exists a protected polymorphism
 \begin{equation}
-\begin{aligned} \label{eq:migration_model_unconditional_symmetric_solution}
+\begin{aligned} \label{eq:migration_model_unconditional_symmetric_polymorphism}
 x_1^* &= \frac{w-1-m(w+1) + \sqrt{\Delta}}{2(w-1)}, \\
@@
 where $\Delta = m^2(w+1)^2+(1-2m)(w-1)^2$.
-Examining the stability of these equilibria, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $a \lambda^2 + b \lambda + c$ with
+\item With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), fixation of $B$ is unstable, and fixation of $A$ is stable if and only if the vertical transmission rate in deme 2 is below $\hat \rho$, where
+\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
+\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)},
+\end{equation}
+or if the migration rate is above $\hat m$, where
+\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
+m > \hat m = \frac{\rho (w-1)}{\rho (w-1) +1-\rho}.
+\end{equation}
+\end{itemize}
+
+\paragraph{Proof.}
+
+Examining the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $a \lambda^2 + b \lambda + c$ with
 \begin{equation}
@@
 \end{equation}
-Since $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is positive, both equilibria are unstable, and the protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_solution}.
+Since $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is positive, both equilibria are unstable, and the protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
 
-With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), the stability of $(0,0)$ is determined by a characteristic polynomial with coefficients
+When $\rho_1=1$ and $\rho_2=\rho$, the stability of $(0,0)$ is determined by a characteristic polynomial with coefficients
 \begin{equation}
@@
 Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any valid $\rho$, the leading eigenvalue is positive, and $(0,0)$ is unstable.
-Regarding the stability of $(1,1)$, which is determined by a characteristic polynomial  with coefficients
+
+The stability of $(1,1)$ is determined by a characteristic polynomial  with coefficients
 \begin{equation}
@@
 Once more, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
-However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur at $(0,1)$ and the fixation is stable.
-We can also determine the rate $\hat \rho$ for which $a+b+c=0$, such that for
-\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
+However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and the fixation is stable.
+
+For any given $\rho$, $a+b+c>0$ if and only if
+\begin{equation}
 \rho < \hat \rho = \frac{m}{m+(1-m)(w-1)}
 \end{equation}
-the fixation $(1,1)$ is stable if and only if $2a+b>0$.
+and then $(1,1)$ is stable if and only if $2a+b>0$.
 We already found that when $\rho=0$ we have $2a+b>0$.
@@
 
-A similar condition can be found for $m$: the value of $a+b+c$ at $m=0$ and $m=1$ is $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when
-\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
+For $m$, the value of $a+b+c$ at $m=0$ and $m=1$ is $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when 
+\begin{equation}
 m > \hat m = \frac{\rho (w-1)}{\rho (w-1) +1-\rho}.
@@
 
-\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$ (i.e.
-the frequencies that solve \eqref{eq:migration_model_unconditional_symmetric_solution}) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar intensities (see figure legend).
+\paragraph{Examples.}
+
+\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar intensities.
 Notably, without oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
@@
 
-The polymorphism $(x_1^*, x_2^*)$ is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, and phenotype $A$ fixes. 
+The polymorphism $(x_1^*, x_2^*)$ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, in which case fixation of phenotype $A$ is stable.
 For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the stronger migration is, the less oblique transmission is required to fix $A$, see shaded areas in \autoref{fig:migration_rho}.
@@
 Migration rate $m$ is 0.05, 0.1, or 0.25 I the left, middle, and right columns, respectively.
-The shaded area denotes stables fixation of phenotype $A$ according to~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
+The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
+Lines drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differed by less than $10^{-4}$, starting with equal frequencies.
 }
@@
 Migration rate $m$ is on the x-axis.
-The shaded area denotes stables fixation of phenotype $A$ according to~\eqref{eq:migration_model_unconditional_symmetric_condition_m}.
+The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
+Lines drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differed by less than $10^{-4}$, starting with equal frequencies.
 }
@@
 Models of cultural transmission mostly assume a fixed relative rate at which different modes of transmission -- vertical, horizontal, or oblique -- occur.
-Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly. 
+Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly, or over space.
 Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ (to be continued) % TODO
@@
 
+\newpage
+
 \bibliographystyle{agsm}
-\bibliography{/Users/yoavram/Documents/library}
-%\bibliography{ms}
+%\bibliography{/Users/yoavram/Documents/library}
+\bibliography{ms}
 

### 2018-05-13 src/collect_figures.py
collect figures from ms.tex - copied and modified from Milpitas

@@ -1,0 +1,41 @@
+# import tempfile
+import shutil
+import os.path
+from os.path import join as join_path
+from os.path import split as split_path
+import re
+from zipfile import ZipFile
+
+if __name__ == '__main__':
+	ms_dir = '.'
+	figures_dir = '../figures'
+	ms_figures_dir = join_path(ms_dir, 'figures')
+	zip_fname = '/Users/yoavram/Dropbox/Sunnyvale/figures.zip'
+	tex_files = ['ms.tex']
+	pattern = re.compile(r'\\includegraphics(?:\[.*\])?\{(.*\.\w{3})\}')
+
+	if os.path.exists(ms_figures_dir):
+		shutil.rmtree(ms_figures_dir)
+		os.mkdir(ms_figures_dir)
+	if not os.path.exists(ms_figures_dir):
+		os.mkdir(ms_figures_dir)
+
+	figures = []
+	for fn in tex_files:
+		with open(join_path(ms_dir, fn)) as f:
+			matches = (pattern.match(line) for line in f)
+			matches = (m for m in matches if m is not None)
+			filenames = (m.groups()[0] for m in matches)
+			filenames = (split_path(fn)[-1] for fn in filenames)
+			filenames = (join_path(figures_dir, fn) for fn in filenames)
+			figures.extend(filenames)
+
+	with ZipFile(zip_fname, 'w') as z:
+		for fn in figures:
+			fn = fn.replace('{', '').replace('}', '')
+			print(fn)
+			shutil.copy(fn, join_path(ms_figures_dir, split_path(fn)[-1]))
+			z.write(fn)
+
+	print("{} figures copied to {} and zipped to {}".format(
+		len(figures), ms_figures_dir, zip_fname))

### 2018-05-15 ms/ms.tex
MWF comments

@@
-\documentclass[11pt]{extarticle}
+\documentclass[14pt]{extarticle}
 \usepackage{geometry}
@@
 \citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific trait is transmitted either vertically or obliquely.
-Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed. 
+Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time.
 For example, \cite{Aoki2005} and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
+In these studies, the cultural transmission rule did not have during the evolutionary process.
 
-Many of these models included fluctuating selection envisaged as due to environmental changes~\citep[reviewed in][]{Aoki2014}. 
+Several of these models included fluctuating selection, which could be due to environmental changes~\citep[reviewed in][]{Aoki2014}. 
 We recently studied a model in which each individual can learn a dichotomous phenotype either from a parent, with probability $\rho$, or from a non-parental adult, with probability $1-\rho$ (\autoref{fig:transmission}).
@@
 We find that $\ldots$ (to be continued)  %TODO
-We suggest that cultural evolution can be affected by environmental heterogeneity via both selection and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
+We suggest that cultural evolution can be affected by environmental changes in both fitness and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
 
@@
 
-% Periodic fluctuating transmission
-\subsection*{Periodic fluctuating transmission}
+% Periodically fluctuating transmission
+\subsection*{Periodically fluctuating transmission}
 
-Suppose that the vertical transmission rate $\rho$ is fluctuating, such that $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
+Suppose that the vertical transmission rate $\rho$ fluctuates, such that $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
 From \eqref{eq:recurrence}, the recurrence equations for two generations is then
@@
 Second, when $s>0$ ($s<0$) because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ ($<1$) and $x''/x' = 1+\frac{\rho_2 s (1-x')}{1+sx'} > 1$ ($<1$), the frequency of $A$ ($B$) increases every generation and $x^*=1$ ($x^*=0$) is globally stable.
-Therefore, we find that fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype without any stable polymorphisms.
+Therefore, fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype, and there can be no stable polymorphisms.
 
-% Periodic fluctuating transmission and selection
-\subsection*{Periodic fluctuating transmission and selection}
+% Periodically fluctuating transmission and selection
+\subsection*{Periodically fluctuating transmission and selection}
 
-Now suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$, and when $B$ is favored the transmission rate is $\rho_B$.
+Suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$, and when $B$ is favored the transmission rate is $\rho_B$.
 The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively,
@@
 \paragraph{Symmetric periods -- $AkBk$.}
-We first consider environments which periodically fluctuate between favoring $A$ and $B$ every $k$ generations.
+We first consider environments which fluctuate periodically every $k$ generations between favoring $A$ and $B$.
 Using a linear approximation, fixations of $A$ ($x^*=1$) and $B$ ($x^*=0$) are locally stable if, respectively,
@@
 \end{aligned}\end{equation}
-and if both of these conditions are not met then there exists a \emph{protected polymorphism}.
+and if neither of these conditions are met then there exists a \emph{protected polymorphism}.
 So for a protected polymorphism we require
@@
 \end{equation}
-Surprisingly, the stronger selection is, the larger the difference can be that still allows a stable polymorphism.
+Surprisingly, the stronger the selection (i.e. the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.
 
@@
 b &= s(1-\rho_A)(2\rho_B - s\rho_A(1-\rho_B)), \\
-c &= \rho_B - \rho_A - s\rho_A(1-\rho_B),
+c &= \rho_B - \rho_A - s\rho_A(1-\rho_B).
 \end{aligned} \end{equation}
 % $a+b+c>0$.
-then $x^*$ is a solution of the quadratic $G(x)=0$. %, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
+Then $x^*$ is a solution of the quadratic $G(x)=0$. %, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
 The condition \eqref{eq:poly_condition_periodic_fluc}, which guarantees that $0 < x^* < 1$, is equivalent to $c<0$, and therefore it also guarantees that $\sqrt{b^2-4ac} > b$.
-So, if $0<s<1$ (such that $b$ is guaranteed to be positive given LHS of eq.~\ref{eq:poly_condition_periodic_fluc}), then 
+So, if $0<s<1$ (so that $b$ is guaranteed to be positive if LHS of eq.~\ref{eq:poly_condition_periodic_fluc}), then 
 \begin{equation}
@@
 \autoref{fig:rho1_rho2_phase_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
-The figure demonstrates that the stronger selection is (x-axis), the greater the fluctuations in $\rho$ can be (y-axis) while still maintaining a polymorphic population (area between the dashed lines).
+The figure demonstrates that the stronger the selection (x-axis), the greater the fluctuations in $\rho$ can be (y-axis) that still allow a polymorphic population (area between the dashed lines).
 
@@
 The stable equilibrium of the frequency of phenotype $A$ \eqref{eq:xstar_periodic_fluc_k=1} for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
-Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ that allow a protected polymorphism \eqref{eq:poly_condition_periodic_fluc}.
+Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from~\eqref{eq:poly_condition_periodic_fluc} that allow protected polymorphism.
 Here, $\rho_A=0.5$.}
@@
 \paragraph{Asymmetric period -- $AkBl$.}
-In the more general case, phenotype $A$ is favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle such that $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
+More generally, phenotype $A$ could be favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle with $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
 The requirements for a protected polymorphism are now
 \begin{equation}\begin{aligned}
-1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b - (\rho_B - \rho_A) s/(1+s)\big)^l,\\
-1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a + (\rho_B - \rho_A) s)^l,
+1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b - \Delta \rho s/(1+s)\big)^l,\\
+1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a + \Delta \rho s)^l,
 \end{aligned}\end{equation}
 where $a=1+\rho_A s>1$, $b=1-\rho_A\frac{s}{1+s}<1$, and $\Delta \rho = \rho_B - \rho_A$.
-This leads to a condition similar to eq.~20 in~\citet{Ram2018}, but more complex due to the addition of $\Delta \rho$:
+This leads to a condition similar to eq.~20 in~\citet{Ram2018}, but more complex due to the inclusion of $\Delta \rho$:
 \begin{equation} \label{eq:poly_condition_periodic_fluc_k_l}
@@
 \end{equation}
-Therefore, increasing the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$  will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
+Therefore, for a given value of $\rho_A$, increasing the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$  will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
 
@@
 
-We now allow both transmission and selection to fluctuate randomly.
+Now suppose that both selection and transmission fluctuate randomly.
 Rewrite eq.~28 from \citet{Ram2018} so that $\rho$ is also a random variable
@@
 where $s_t$ are i.i.d (independent and identically distributed), $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$, $\rho_t$ are i.i.d, and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
-Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D)$, and from Results~6 and 7 of \citet{Ram2018} we have:
+Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D)$. 
+From \citet[][results~6 and 7]{Ram2018} we have:
 \begin{itemize}
@@
 
-Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, we can get fixation.
+Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, there can be fixation.
 Let $w_{A,t}, w_{B,t} \sim U(0,1)$, $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$ and $\rho_t \sim Beta(1+s_t, 1)$ (a beta distribution with parameters $1+s_t$ and $1$). 
@@
 $s_t<0$; \autoref{fig:beta}).
-Then $E[log(1+\rho_t s_t)] >0$ and $B$ almost surely doesn't fix, and $E[log(1-\rho_t s_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
+Then $E[log(1+\rho_t s_t)] >0$ and $B$ almost surely doesn't fix. 
+Also, $E[log(1-\rho_t s_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
 The opposite occurs if $\rho_t \sim \beta(1, 1+s_t)$ and the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -4$).
@@
 \textbf{(C)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
-\textbf{(C)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
+\textbf{(D)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
 }
@@
 Third, it is also possible that both fixations are not stochastically locally stable even if $s_t$ and $\rho_t$ covary, but, as in the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
-For example, \autoref{fig:rho1_rho2_stoch_p} shows the expected outcome when $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
-The blue and red areas denote expected fixation of $A$ and $B$, respectively (i.e.
-stochastic local stability) and the white area shows expected polymorphism (neither fixation is stochastically locally stable).
+For example, \autoref{fig:rho1_rho2_stoch_p} shows the expected outcome when $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$, while $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
+The blue and red areas show expected fixation of $A$ and $B$, respectively (i.e.
+stochastic local stability) and the white area shows expected protected polymorphism (i.e.
+neither fixation is stochastically locally stable).
 
@@
 Here, $s_t=0.05$ and $\rho_t=\rho_1$ with probability $p=0.505$ and $s_t=-0.05$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
-The diagonal represents the case of no transmission fluctuations; \citet[Fig.~2]{Ram2018} demonstrated that with a constant transmission rate $\rho=0.1$ and the above distribution of $s_t$, both fixations are not stochastically stable.
+The diagonal represents the case of no transmission fluctuations; \citet[Fig.~2]{Ram2018} demonstrated that with a constant transmission rate $\rho=0.1$ and the above distribution of $s_t$, neither fixation is stochastically stable.
 }
@@
 To include the effects of random genetic drift due to finite population size in the above deterministic model, we follow \citet{Ram2018} and develop a diffusion equation approximation.
-In~\citet{Ram2018} only selection fluctuated via $s_t$, but here we also let transmission to fluctuate via $\rho_t$.
+In~\citet{Ram2018} only selection fluctuated via $s_t$, but here we also have transmission fluctuating via $\rho_t$.
 
-We find a result similar to Result 11 from~\citet{Ram2018}:
+We obtain a result similar to result 11 from~\citet{Ram2018}:
 The mean $\mu(x)$ and variance $\sigma^2(x)$ of the change in the frequency $x$ of phenotype $A$ in the case of a cycling environment $AkBl$, where $k+l=n$, are
@@
 where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$.
-Furthermore, combining \eqref{eq:drift_diffusion_terms} with eqs.~46-47 from \citep{Ram2018}, we find that the fixation probability of phenotype $A$, starting with frequency $x$ of phenotype $A$, is
+Furthermore, combining \eqref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the fixation probability of phenotype $A$, starting from a frequency $x$ of phenotype $A$, is
 \begin{equation}
@@
 \end{equation}
-From Result 10 in \citep{Ram2018}, $u(x)$ is monotone increasing in $S_n/n$ which is the average selection coefficient of $A$ weighted by the vertical transmission rates $\rho_t$.
+From result 10 in \citep{Ram2018}, $u(x)$ is monotone increasing in $S_n/n$ which is the average selection coefficient of $A$ weighted by the vertical transmission rates $\rho_t$.
 Therefore, if $s_t$ and $\rho_t$ are positively (negatively) correlated, $S_n/n$ increases (decreases), and the fixation probability $u(x)$ increases (decreases). 
@@
 learning, development) occur within the demes, followed by migration of sub-adults -- individuals that already acquired their phenotype but have yet to reproduce.
-The frequency of phenotype $A$ in deme $j$ is marked by $x_j$, and therefore after selection and transmission the frequencies are 
+The frequency of phenotype $A$ in deme $j$ is denoted by $x_j$, and therefore after selection and transmission the frequencies are 
 \begin{equation} \label{eq:migration_model_selection_transmission}
@@
 
-For the simplified case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} and symmetric selection $w_1=1/w_2=w>1$, the model is given by
+For the simple case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} and symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 
-we have the following results:
+We have the following results:
 \begin{itemize}
@@
 where $\Delta = m^2(w+1)^2+(1-2m)(w-1)^2$.
-\item With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), fixation of $B$ is unstable, and fixation of $A$ is stable if and only if the vertical transmission rate in deme 2 is below $\hat \rho$, where
+\item With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), fixation of $B$ is unstable, and fixation of $A$ is stable if and only if the vertical transmission rate in deme 2 is below $\hat \rho$; that is,
 \begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
@@
 \end{equation}
-or if the migration rate is above $\hat m$, where
+or if the migration rate is above $\hat m$; that is,
 \begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
@@
 
-Examining the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $a \lambda^2 + b \lambda + c$ with
+For the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $a \lambda^2 + b \lambda + c$ with
 \begin{equation}
@@
 \end{equation}
-Since $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is positive, both equilibria are unstable, and the protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
+Since $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is greater than one, both equilibria are unstable, and a protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
 
-When $\rho_1=1$ and $\rho_2=\rho$, the stability of $(0,0)$ is determined by a characteristic polynomial with coefficients
+When $\rho_1=1$ and $\rho_2=\rho$, the stability of $(0,0)$ (i.e. fixation in $B$) is determined by the characteristic polynomial with coefficients
 \begin{equation}
@@
 For $\rho=0$ we have $a+b+c=-m(w-1)<0$.
-Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any valid $\rho$, the leading eigenvalue is positive, and $(0,0)$ is unstable.
+Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any valid $\rho$, the leading eigenvalue is greater than one, and $(0,0)$ is unstable.
 
-The stability of $(1,1)$ is determined by a characteristic polynomial  with coefficients
+The stability of $(1,1)$ (i.e. fixation in $A$) is determined by the characteristic polynomial with coefficients
 \begin{equation}
@@
 b &= -\frac{1-m}{w}(1 + w(1-\rho) +w^2 \rho), \quad
-c &= \frac{1-2m}{w} (w\rho+1-\rho).
+c &= \frac{1-2m}{w} (w\rho+1-\rho),
 \end{aligned}
 \end{equation}
-Once more, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
-However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and the fixation is stable.
+and real roots. 
+Again, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
+However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and fixation is stable.
 
-For any given $\rho$, $a+b+c>0$ if and only if
+For any $\rho$, $a+b+c>0$ if and only if
 \begin{equation}
-\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)}
+\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)},
 \end{equation}
 and then $(1,1)$ is stable if and only if $2a+b>0$.
-We already found that when $\rho=0$ we have $2a+b>0$.
+We have already seen that when $\rho=0$ we have $2a+b>0$.
 When $\rho = \hat \rho$, we have $2a+b=\frac{m(w+1)+w-1}{w(1-m)(w-1)}>0$.
-Since $2a+b$ is linear in $\rho$, we find that $2a+b>0$ for $\hat \rho \le \rho \le 0$ and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_rho} guarantees the stable fixation of $A$.
+Since $2a+b$ is linear in $\rho$, we find that $2a+b>0$ for $\hat \rho \ge \rho \ge 0$ and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_rho} guarantees the stable fixation of $A$.
 
-For $m$, the value of $a+b+c$ at $m=0$ and $m=1$ is $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when 
+The values of $a+b+c$ at $m=0$ and $m=1$ are $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when 
 \begin{equation}
@@
 
-\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar intensities.
-Notably, without oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
+\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
+Notably, in the absence of oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
 With some oblique transmission, but equal in both demes, results are similar (not shown).
-However, if oblique transmission is stronger in deme 2 compared to deme 1 (\autoref{fig:migration_m}, middle and right columns), the stable frequency of $A$ increases in both demes.
+However, if oblique transmission is stronger in deme 2 than in deme 1 (\autoref{fig:migration_m}, middle and right columns), the stable frequency of $A$ increases in both demes.
 Therefore, the mean fitness in deme 1 decreases to a lesser extent than in deme 2, and even increases when the migration rate is high enough.
@@
 \textbf{Oblique transmission and migration: effect of transmission.} 
-The figure shows $x^*_i$ the stable frequency of $A$ (top row) and $\bar{w}^*_i$ population mean fitness (bottom row) in the two demes.
+The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
@@
 The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
-Lines drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differed by less than $10^{-4}$, starting with equal frequencies.
+Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
 }
@@
 \textbf{Oblique transmission and migration: effect of migration.} 
-The figure shows $x^*_i$ the stable frequency of $A$ (top row) and $\bar{w}^*_i$ population mean fitness (bottom row) in the two demes.
+The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
@@
 The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
-Lines drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differed by less than $10^{-4}$, starting with equal frequencies.
+Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
 }
@@
 
-Models of cultural transmission mostly assume a fixed relative rate at which different modes of transmission -- vertical, horizontal, or oblique -- occur.
+Most models of cultural transmission assume a fixed relative rate at which different modes of transmission -- vertical, horizontal, or oblique -- occur.
 Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly, or over space.

### 2018-05-27 ms/ms_sunnyvale.tex
MWF comments

--- 2018-05-15 ms/ms.tex
@@
 Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time.
-For example, \cite{Aoki2005} and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
-In these studies, the cultural transmission rule did not have during the evolutionary process.
+For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
+In these studies, the cultural transmission rule did not change during the evolutionary process.
 
@@
 There has been much less theory developed for fluctuating transmission.
-Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size or density~\citep{Fischer2015,Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission will fluctuate over time and/or space.
+Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size or density~\citep{Fischer2015,Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission may fluctuate over time and/or space.
 For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases.
+%[MWF-NOTE: non-parental: adults in the parental generation]
 %For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
@@
 \textbf{Cultural transmission with mixed vertical and oblique transmission.}
-When a newborn matures, she copies her phenotype -- color -- from her mother with probability $\rho$, therefore becoming blue, or from some other female with probability $1-\rho$, in which case her color will depend on the frequency of blue and red adult females.}
+When a newborn matures, she copies the phenotype---color---of her mother with probability $\rho$, therefore becoming blue, or from some other female with probability $1-\rho$, in which case her color will depend on the frequency of blue and red adult females.}
 \label{fig:transmission}
@@
 
-% Model & Results
-\section*{Model \& Results}
+% Model and Results
+\section*{Model and Results}
 
@@
 
-\paragraph{Symmetric periods -- $AkBk$.}
-We first consider environments which fluctuate periodically every $k$ generations between favoring $A$ and $B$.
+\paragraph{Symmetric periods---$AkBk$.}
+We first consider environments that fluctuate periodically every $k$ generations between favoring $A$ and $B$.
 Using a linear approximation, fixations of $A$ ($x^*=1$) and $B$ ($x^*=0$) are locally stable if, respectively,
@@
 and if neither of these conditions are met then there exists a \emph{protected polymorphism}.
-So for a protected polymorphism we require
+Thus for a protected polymorphism, we require
+% TODO check RHS
 \begin{equation}
 1 < F'_A(1) F'_B(1) = \Big(1-\rho_A\frac{s}{1+s}\Big)\Big(1+\rho_B s\Big) 
-= 1-\frac{1}{1+s}\big[\rho_B(1+s-\rho_A)-\rho_A\big],
+= 1+\frac{s}{1+s}\big[\rho_B(1+s-\rho_A)-\rho_A\big],
 \end{equation}
@@
 
-\paragraph{Period of one -- $A1B1$.}
+\paragraph{Period of one---$A1B1$.}
 In the case of $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
-Denote $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
+Set $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
 \begin{equation} \begin{aligned} \label{eq:xstar_periodic_fluc_k=1}
@@
 The condition \eqref{eq:poly_condition_periodic_fluc}, which guarantees that $0 < x^* < 1$, is equivalent to $c<0$, and therefore it also guarantees that $\sqrt{b^2-4ac} > b$.
-So, if $0<s<1$ (so that $b$ is guaranteed to be positive if LHS of eq.~\ref{eq:poly_condition_periodic_fluc}), then 
+So, if $0<s<1$ (so that $b$ is guaranteed to be positive if LHS of eq.~\ref{eq:poly_condition_periodic_fluc} holds), then 
 \begin{equation}
@@
 \autoref{fig:rho1_rho2_phase_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
-The figure demonstrates that the stronger the selection (x-axis), the greater the fluctuations in $\rho$ can be (y-axis) that still allow a polymorphic population (area between the dashed lines).
+The figure demonstrates that the stronger the selection (x-axis), the greater the fluctuations in $\rho$  (y-axis) that still allow a polymorphic population (area between the dashed lines).
 
@@
 \textbf{Protected polymorphism.}
-The stable equilibrium of the frequency of phenotype $A$ \eqref{eq:xstar_periodic_fluc_k=1} for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
+The stable equilibrium of the frequency of phenotype $A$ (from eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
 Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from~\eqref{eq:poly_condition_periodic_fluc} that allow protected polymorphism.
@@
 
-\paragraph{Asymmetric period -- $AkBl$.}
-More generally, phenotype $A$ could be favored for $k$ generations and phenotype $B$ for $l$ generations, and the transmission rate follows the same cycle with $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
+\paragraph{Asymmetric period---$AkBl$.}
+More generally, phenotype $A$ could be favored for $k$ generations and phenotype $B$ for $l$ generations, with the transmission rate following the same cycle with $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
 The requirements for a protected polymorphism are now
@@
 \end{equation}
-Therefore, for a given value of $\rho_A$, increasing the vertical transmission rate specifically when $B$ is favored decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$  will have the opposite effect, increasing the ratio $l/k$ that permits a protected polymorphism.
+Therefore, for a given value of $\rho_A$, if $\rho_B>\rho_A$ then increasing the vertical transmission rate of $B$ decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$  will have the opposite effect; it increases the ratio $l/k$ that permits a protected polymorphism.
 
@@
 \begin{equation}
-x_t = x_t \frac{1 + \rho_t s_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
+x_t = x_t \frac{1 + z_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
 \end{equation}
@@
 \begin{itemize}
-\item Suppose $E[log(1+\rho_t s_t)]>0$. Then $x^*=0$ is not stochastically locally stable and in fact $P(lim_{t \to \infty} x_t=0) = 0$, i.e., fixation of $B$ almost surely does not occur.
-\item Suppose $E[log(1+\rho_t s_t)]<0$. Then $x^*=0$ is stochastically locally stable. 
-\item Similarly, if $E[log(1-\rho_t s_t/(1+s_t)]<0$, then $x^*=1$ is stochastically locally stable, and if $E[log(1-\rho_t s_t/(1+s_t))]>0$, then fixation of $A$ almost surely does not occur.
-\item In particular, if $E[\rho_t s_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-\rho_t s_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
-\item It is not possible that $E[log(1+\rho_t s_t)]$ and $E[log(1-\rho_t s_t/(1+s_t)]$ are both negative, as their sum is positive
+\item Suppose $E[log(1+z_t)]>0$. Then $x^*=0$ is not stochastically locally stable and in fact $P(lim_{t \to \infty} x_t=0) = 0$, i.e., fixation of $B$ almost surely does not occur.
+\item Suppose $E[log(1+z_t)]<0$. Then $x^*=0$ is stochastically locally stable. 
+\item Similarly, if $E[log(1-z_t/(1+s_t)]<0$, then $x^*=1$ is stochastically locally stable, and if $E[log(1-z_t/(1+s_t))]>0$, then fixation of $A$ almost surely does not occur.
+\item In particular, if $E[z_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-z_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
+\item It is not possible that $E[log(1+z_t)]$ and $E[log(1-z_t/(1+s_t)]$ are both negative, as their sum is positive
 \begin{multline}
-E[log(1+\rho_t s_t)] + E[log(1-\rho_t s_t/(1+s_t)] = \\
-E[log(1+\rho_t s_t) + log(1-\rho_t s_t/(1+s_t)]= \\
-E[log\big((1+\rho_t s_t)(1-\rho_t s_t/(1+s_t)\big)]= \\
+E[log(1+z_t)] + E[log(1-z_t/(1+s_t)] = \\
+E[log(1+z_t) + log(1-z_t/(1+s_t)]= \\
+E[log\big((1+z_t)(1-z_t/(1+s_t)\big)]= \\
 E[log\big( 1+\rho_t(1-\rho_t)s_t^2/(1+s_t) \big)] > 0,
@@
 \paragraph{Examples.}
-First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)=0$) and $s_t$ is symmetric around zero, then $E[\rho_t s_t]=0$ and $E[-\rho_t s_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
+First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)=0$) and $s_t$ is symmetric around zero, then $E[z_t]=0$ and $E[-z_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
 Therefore, fixation of $B$ is stochastically locally stable and fixation of $A$ almost surely does not occur.
-For example, let $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then $E[log(1+\rho_t s_t)]\approx -0.07315$ and $E[log(1-\rho_t s_t/(1+s_t))]\approx 0.2337$.
+For example, let $s_t \sim U(-1, 1)$ and $\rho_t \sim U(0,1)$ independently (in particular, $cov(\rho_t, s_t)=0$), then $E[log(1+z_t)]\approx -0.07315$ and $E[log(1-z_t/(1+s_t))]\approx 0.2337$.
 
@@
 Therefore, if we take the i.i.d fitness random variables for $A$ and $B$ to be $w_{A,t}, w_{B,t} \sim U(0,1)$, respectively, and define $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$, then neither $A$ nor $B$ has an advantage, on average (i.e.
-$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[\rho_t s_t]$ and $E[-\rho_t s_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to converge to a polymorphic distribution.
+$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to converge to a polymorphic distribution.
 
@@
 $s_t<0$; \autoref{fig:beta}).
-Then $E[log(1+\rho_t s_t)] >0$ and $B$ almost surely doesn't fix. 
-Also, $E[log(1-\rho_t s_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
+Then $E[log(1+z_t)] >0$ and $B$ almost surely doesn't fix. 
+Also, $E[log(1-z_t / (1+s_t)] < 0$, so fixation of $A$ is stochastically locally stable.
 The opposite occurs if $\rho_t \sim \beta(1, 1+s_t)$ and the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -4$).
@@
 
-To include the effects of random genetic drift due to finite population size in the above deterministic model, we follow \citet{Ram2018} and develop a diffusion equation approximation.
+To include the effects of random genetic drift due to finite population size in the above deterministic model, we follow \citet{Ram2018} and develop a diffusion  approximation.
 In~\citet{Ram2018} only selection fluctuated via $s_t$, but here we also have transmission fluctuating via $\rho_t$.
@@
 \end{equation}
-where $S_n = \sum_{t=1}^{n}{\rho_t s_t}$.
+where $S_n = \sum_{t=1}^{n}{z_t}$.
 Furthermore, combining \eqref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the fixation probability of phenotype $A$, starting from a frequency $x$ of phenotype $A$, is
@@
 
-We now describe a model in which fluctuations in selection and transmission occur in space, rather then time; that is, we model a population divided to two demes.
+We now describe a model in which fluctuations in selection and transmission occur in space, rather then time; we model a population divided to two demes.
 Selection (e.g.
 reproduction) and transmission (e.g.
-learning, development) occur within the demes, followed by migration of sub-adults -- individuals that already acquired their phenotype but have yet to reproduce.
+learning, development) occur within the demes, followed by migration of sub-adults---individuals that have already acquired their phenotype but have yet to reproduce.
 The frequency of phenotype $A$ in deme $j$ is denoted by $x_j$, and therefore after selection and transmission the frequencies are 
@@
 
-\paragraph{Unconditional migration and symmetric selection.}
+\paragraph{Unconditional migration and symmetric selection.}%[MWF: Why call it unconditional?]
 
-For the simple case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} and symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+For the simple case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and fixation is stable.
+%[MWF: Are the roots real?]
 
@@
 \end{equation}
-In addition, $2a+b$ is linear in $m$ and equal to 2 when $m=1$; it is equal to zero when $m=(w-1)(w\rho-1)/(w(w-1)\rho+w+1) < \hat m$, and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_m} is sufficient for stability of fixation of $A$.
+In addition, $2a+b$ is linear in $m$ and equal to 2 when $m=1$; it is equal to zero when $m=(w-1)(w\rho-1)/(w(w-1)\rho+w+1) < \hat m$, and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_m}
+is sufficient for stability of fixation of $A$.
+%[MWF: (22, also 27)]
 
@@
 
-\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection such that $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
+\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection where $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
 Notably, in the absence of oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
@@
 The polymorphism $(x_1^*, x_2^*)$ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, in which case fixation of phenotype $A$ is stable.
-For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the stronger migration is, the less oblique transmission is required to fix $A$, see shaded areas in \autoref{fig:migration_rho}.
-Similarly, with enough migration ($m > \hat m$), phenotype $A$ fixes, and the more oblique transmission in deme 2, the less migration is needed to fix $A$; see shaded area in \autoref{fig:migration_m}.
+For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the stronger migration is, the less oblique transmission is required to fix $A$ (see shaded areas in \autoref{fig:migration_rho}).
+Similarly, with enough migration ($m > \hat m$), phenotype $A$ fixes, and the more oblique transmission in deme 2, the less migration is needed to fix $A$ (see shaded area in \autoref{fig:migration_m}).
 
@@
 
-Most models of cultural transmission assume a fixed relative rate at which different modes of transmission -- vertical, horizontal, or oblique -- occur.
+Most models of cultural transmission assume a fixed relative rate at which different modes of transmission---vertical, horizontal, or oblique---occur.
 Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly, or over space.
@@
 
-One caveat of our model is the use of the ``phenotypic gambit'' -- the assumption that the transmission mode itself is strictly vertically transmitted.
+One caveat of our model is the use of the ``phenotypic gambit''---the assumption that the transmission mode itself is strictly vertically transmitted.
 Although there is evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
@@
 An extension of our model could incorporate more sophisticated oblique transmission schemes.
-For example, conformity -- preference for learning a frequent phenotype -- has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
+For example, conformity---preference for learning a frequent phenotype---has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
 We suggest that the specific mode of oblique transmission can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions.
@@
 %\bibliography{/Users/yoavram/Documents/library}
-\bibliography{ms}
+\bibliography{ms_sunnyvale}
 

### 2018-05-29 ms/ms_sunnyvale.tex
real roots for some cases, need to finish

@@
-\documentclass[14pt]{extarticle}
+\documentclass[12pt]{extarticle}
 \usepackage{geometry}
@@
 
-Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from non-parental adults to offspring) transmission. 
+Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from adults in the parental generation to offspring) transmission. 
 \citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific trait is transmitted either vertically or obliquely.
@@
 Several of these models included fluctuating selection, which could be due to environmental changes~\citep[reviewed in][]{Aoki2014}. 
-We recently studied a model in which each individual can learn a dichotomous phenotype either from a parent, with probability $\rho$, or from a non-parental adult, with probability $1-\rho$ (\autoref{fig:transmission}).
+We recently studied a model in which each individual can learn a dichotomous phenotype either from a parent, with probability $\rho$, or from an adult in the parental generation, with probability $1-\rho$ (\autoref{fig:transmission}).
 We found that if selection fluctuates between favoring each of the two phenotypes, but on average favored both phenotypes for similar time periods, then a phenotypic polymorphism may be maintained~\citep{Ram2018}.
@@
 For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases. 
-%[MWF-NOTE: non-parental: adults in the parental generation]
 %For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
 
-Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from non-parental adults with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space~(\autoref{fig:transmission}).
+Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space~(\autoref{fig:transmission}).
 We find that $\ldots$ (to be continued)  %TODO
@@
 
-\paragraph{Unconditional migration and symmetric selection.}%[MWF: Why call it unconditional?]
+\paragraph{Unconditional migration and symmetric selection.}
 
-For the simple case of unconditional migration $m_1=m_2=m$ \citep{McPeek1992} after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+For the simple case of unconditional migration $m_1=m_2=m<1/2$ \citep{McPeek1992} after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 
-For the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $a \lambda^2 + b \lambda + c$ with
+For the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $f(\lambda) = a \lambda^2 + b \lambda + c$ with
 \begin{equation}
@@
 \end{equation}
-Since $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is greater than one, both equilibria are unstable, and a protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
+First, the discriminant $\Delta = b^2-4ac$ is positive for any real $m$, because, when writing it as a quadratic polynomial in $m$, $\Delta(m)=\frac{1}{w^{2}} \big(m^{2} (w^2+1)^2 - 2 m (w^2-1)^2 + (w^2-1)^2\big)$, this polynomial is positive at $m=0$ and $m=1/2$ and has a negative discriminant $-16w^2(w^2-1)^2$.
+Therefore, the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two positive roots. \\
+Second, as $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is greater than one, both equilibria are unstable, and a protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
 
-When $\rho_1=1$ and $\rho_2=\rho$, the stability of $(0,0)$ (i.e. fixation in $B$) is determined by the characteristic polynomial with coefficients
+When $\rho_1=1$ and $\rho_2=\rho$, the stability of $(0,0)$ (i.e. fixation in $B$) is determined by the characteristic polynomial $f(\lambda)$ with coefficients
 \begin{equation}
@@
 a &= 1, \quad
-b &= -\frac{(1-m)(w+1-\rho+\rho/w)}{w}, \quad
+b &= -(1 - m) (w + 1-\rho + \rho/w), \quad
 c &=  (1-2m)(w(1-\rho)+\rho).
@@
 \end{equation}
-Again, $a>0$ and $c>0$ and stability can be analyzed by the sign of $a+b+c$.
+
+First, the discriminant is 
+$$
+\Delta = (1 - m)^2 (w + 1-\rho + \rho/w)^2 - 4(1-2m)(w(1-\rho)+\rho),
+$$
+which is positive for $m=0$ and $m=1/2$:
+\begin{equation}\begin{aligned}
+\Delta(m=0) = \Big(\frac{(w + \rho) (w-1)}{w}\Big)^2 \\
+\Delta(m=1/2) = \Big(\frac{w^2 + w(1-\rho) + \rho}{2 w}\Big)^2.
+\end{aligned}\end{equation}
+In addition, if we write $\Delta$ as a polynomial in $m$, its discriminant is negative -- $- 16 (w + \rho)^2 (w - 1)^2 \big(w (1 - \rho) + \rho \big) / w^2$, and therefore the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two real roots.
+\\
+Second, $a>0$, $b<0$, and $c>0$ and stability can be analyzed by the sign of $a+b+c$.
 For $\rho=1$ we already have $a+b+c<0$.
@@
 a &= 1, \quad
-b &= -\frac{1-m}{w}(1 + w(1-\rho) +w^2 \rho), \quad
-c &= \frac{1-2m}{w} (w\rho+1-\rho),
+b &= -(1-m)(w \rho + 1 -\rho + 1/w), \quad
+c &= (1-2m) \big((1-\rho)/w + \rho\big),
 \end{aligned}
 \end{equation}
-and real roots. 
+%[MWF: Are the roots real?]
 Again, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
 However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and fixation is stable.
-%[MWF: Are the roots real?]
+%
 
@@
 The values of $a+b+c$ at $m=0$ and $m=1$ are $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when 
-\begin{equation}
+\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m2}
 m > \hat m = \frac{\rho (w-1)}{\rho (w-1) +1-\rho}.
 \end{equation}
-In addition, $2a+b$ is linear in $m$ and equal to 2 when $m=1$; it is equal to zero when $m=(w-1)(w\rho-1)/(w(w-1)\rho+w+1) < \hat m$, and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_m}
+In addition, $2a+b$ is linear in $m$ and equal to 2 when $m=1$; it is equal to zero when $m=(w-1)(w\rho-1)/(w(w-1)\rho+w+1) < \hat m$, and therefore (\ref{eq:migration_model_unconditional_symmetric_condition_m}, also \ref{eq:migration_model_unconditional_symmetric_condition_m2})
 is sufficient for stability of fixation of $A$.
-%[MWF: (22, also 27)]
 

### 2018-05-30 ms/ms_sunnyvale.tex
symmetric migartion

@@
-\documentclass[12pt]{extarticle}
+\documentclass[14pt]{extarticle}
 \usepackage{geometry}
@@
 
-\paragraph{Unconditional migration and symmetric selection.}
+\paragraph{Symmetric migration and selection.}
 
-For the simple case of unconditional migration $m_1=m_2=m<1/2$ \citep{McPeek1992} after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+For the simple case of symmetric migration $m_1=m_2=m<1/2$ after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}

### 2018-06-03 ms/ms_sunnyvale.tex
finished real roots; todos and revisions discussed with MWF

@@
 \end{figure*}
+
+% TODO what can of spatial differences can we imagine in transmission?
+% two populations, one has first cousin marriages and one doesn't (endogmia)
+% orthodox religious communities
+% boarding schools for one population
+% foraging in groups or families
+% pigmies women 'migration' - adoption by husband family
+% interesting: migration correlated with transmission rate, future work
+
+% matrilineal primates - vertical transmission stronger in females than males
+% similar differences in ultra religious human societies
 
@@
 
-For the simple case of symmetric migration $m_1=m_2=m<1/2$ after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+For the simple case of symmetric migration $m_1=m_2=m<1/2$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 \end{equation}
-First, the discriminant $\Delta = b^2-4ac$ is positive for any real $m$, because, when writing it as a quadratic polynomial in $m$, $\Delta(m)=\frac{1}{w^{2}} \big(m^{2} (w^2+1)^2 - 2 m (w^2-1)^2 + (w^2-1)^2\big)$, this polynomial is positive at $m=0$ and $m=1/2$ and has a negative discriminant $-16w^2(w^2-1)^2$.
+First, the discriminant $\Delta = b^2-4ac$ is positive for any real $m$, because, when writing it as a quadratic polynomial in $m$, $\Delta(m)=\frac{1}{w^{2}} \big(m^{2} (w^2+1)^2 - 2 m (w^2-1)^2 + (w^2-1)^2\big)$, this polynomial is positive at $m=0$ and $m=1/2$ and has a negative discriminant $\big(-16w^2(w^2-1)^2\big)$.
 Therefore, the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two positive roots. \\
@@
 a &= 1, \quad
-b &= -(1 - m) (w + 1-\rho + \rho/w), \quad
-c &=  (1-2m)(w(1-\rho)+\rho).
+b &= -(1 - m) (w + 1 - \rho + \rho/w), \quad
+c &=  (1 - 2m)(w(1-\rho)+\rho).
 \end{aligned}
@@
 \end{aligned}\end{equation}
-In addition, if we write $\Delta$ as a polynomial in $m$, its discriminant is negative -- $- 16 (w + \rho)^2 (w - 1)^2 \big(w (1 - \rho) + \rho \big) / w^2$, and therefore the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two real roots.
+In addition, if we write $\Delta$ as a polynomial in $m$, its discriminant is negative $\big(- 16 (w + \rho)^2 (w - 1)^2 \big(w (1 - \rho) + \rho \big) / w^2\big)$, and therefore the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two real roots.
 \\
@@
 
-The stability of $(1,1)$ (i.e. fixation in $A$) is determined by the characteristic polynomial with coefficients
+The stability of $(1,1)$ (i.e. fixation in $A$) is determined by the characteristic polynomial $f(\lambda)$ with coefficients
 \begin{equation}
@@
 a &= 1, \quad
-b &= -(1-m)(w \rho + 1 -\rho + 1/w), \quad
-c &= (1-2m) \big((1-\rho)/w + \rho\big),
+b &= -(1 - m)(w \rho + 1 - \rho + 1/w), \quad
+c &= (1 - 2m) \big((1-\rho)/w + \rho\big),
 \end{aligned}
 \end{equation}
-%[MWF: Are the roots real?]
-Again, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
+First, the discriminant is 
+$$
+\Delta = (1 - m)^2(w \rho + 1 - \rho + 1/w)^2 - 4 (1 - 2m) \big((1-\rho)/w + \rho\big),
+$$
+which is positive for $m=0$ and $m=1/2$:
+\begin{equation}\begin{aligned}
+\Delta(m=0) = \Big(\frac{(w \rho + 1) (w-1)}{w}\Big)^2 \\
+\Delta(m=1/2) = \Big(\frac{w^2 \rho + w(1-\rho) + 1}{2 w}\Big)^2.
+\end{aligned}\end{equation}
+In addition, if we write $\Delta$ as a polynomial in $m$, its discriminant is negative $\big(- 16 (w\rho + 1)^2 (w - 1)^2 \big(\rho w + 1 - \rho \big) / w^3\big)$, and therefore the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two real roots.
+\\
+Second, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
 However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and fixation is stable.
@@
 % National service - Israel, Singapore
+% Boarding schools
 % Immigrtion - draught

### 2018-06-10 ms/ms_sunnyvale.tex
finished migration proof

--- 2018-05-30 ms/ms_sunnyvale.tex
@@
 
-For the simple case of symmetric migration $m_1=m_2=m<1/2$ after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+For the case of symmetric migration $m_1=m_2=m$ ($0\le m \le1/2$) occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 \begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
-\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)},
+\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)} < 1,
 \end{equation}
@@
 \begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
-m > \hat m = \frac{\rho (w-1)}{\rho (w-1) +1-\rho}.
+m > \hat m = \frac{\rho w - 1}{\rho w + 1}.
 \end{equation}
@@
 
-For the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $f(\lambda) = a \lambda^2 + b \lambda + c$ with
+%For the stability of the equilibria when $\rho_1=\rho_2=1$, the characteristic polynomial of the Jacobian is the same for both fixation equilibria $(0,0)$ and $(1,1)$ and is given by $f(\lambda) = a \lambda^2 + b \lambda + c$ with
+%\begin{equation}
+%\begin{aligned}
+%a &= 1, \quad
+%b &= -\frac{(1-m)(w^2+1)}{w}, \quad
+%c &=  1-2m.
+%\end{aligned}
+%\end{equation}
+%First, the discriminant $\Delta = b^2-4ac$ is positive for any real $m$, because, when writing it as a quadratic polynomial in $m$, $\Delta(m)=\frac{1}{w^{2}} \big(m^{2} (w^2+1)^2 - 2 m (w^2-1)^2 + (w^2-1)^2\big)$, this polynomial is positive at $m=0$ and $m=1/2$ and has a negative discriminant $\big(-16w^2(w^2-1)^2\big)$.
+%Therefore, the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two positive roots. \\
+%Second, as $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is greater than one, both equilibria are unstable, and a protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
+
+When $\rho_1=1$ and $\rho_2=\rho$, with $0 \le \rho \le 1$, the stability of $(0,0)$ (i.e.
+fixation in $B$) is determined by the characteristic polynomial of the Jacobian of~\eqref{eq:migration_model_unconditional_symmetric}, $f_0(x)=ax^2+bx+c$, with coefficients
 \begin{equation}
@@
 a &= 1, \quad
-b &= -\frac{(1-m)(w^2+1)}{w}, \quad
-c &=  1-2m.
+b &= -(1 - m) (w + 1 - \rho + \rho/w), \quad
+c &=  (1 - 2m)(\rho + w(1-\rho)).
 \end{aligned}
 \end{equation}
-First, the discriminant $\Delta = b^2-4ac$ is positive for any real $m$, because, when writing it as a quadratic polynomial in $m$, $\Delta(m)=\frac{1}{w^{2}} \big(m^{2} (w^2+1)^2 - 2 m (w^2-1)^2 + (w^2-1)^2\big)$, this polynomial is positive at $m=0$ and $m=1/2$ and has a negative discriminant $-16w^2(w^2-1)^2$.
-Therefore, the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two positive roots. \\
-Second, as $a>0$, $b<0$, and $c>0$, and because $a+b+c=-\frac{1-m}{w}(w-1)^2<0$, the leading eigenvalue is greater than one, both equilibria are unstable, and a protected polymorphism exists and is given by~\eqref{eq:migration_model_unconditional_symmetric_polymorphism}.
 
-When $\rho_1=1$ and $\rho_2=\rho$, the stability of $(0,0)$ (i.e. fixation in $B$) is determined by the characteristic polynomial $f(\lambda)$ with coefficients
+First, the discriminant $\Delta = b^2-4ac$ for $m=0$ and $m=1/2$ is
+\begin{equation}\begin{aligned}
+\Delta(m=0) = \Big(\frac{(w + \rho) (w-1)}{w}\Big)^2, \text{and} \\
+\Delta(m=1/2) = \Big(\frac{w^2 + w(1-\rho) + \rho}{2 w}\Big)^2,
+\end{aligned}\end{equation}
+which are both positive.
+Writing $\Delta$ as a polynomial in $m$, $g(m)$, the discriminant of $g(m)$ is 
+$$
+\big(- 16 (w + \rho)^2 (w - 1)^2 \big(w (1 - \rho) + \rho \big) / w^2\big),
+$$
+which is negative, and therefore the discriminant $\Delta$ of $f_0(x)$ is positive for any $0 \le m \le 1/2$ and $f_0(x)$ has two real roots. \\
+Second, $a>0$, $b<0$, and $c>0$ so $f_0$ is positive with a negative derivative at $x=0$ and has a positive derivative at infinity.
+Therefore stability of $(0,0)$ can be determined by the sign of $f_0(1)=a+b+c$.
+For $\rho=1$ we have $a+b+c=-\frac{1-m}{w}(w-1)^2<0$.
+For $\rho=0$ we have $a+b+c=-m(w-1)<0$.
+Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any $0 \le \rho \le 1$, so $f(x)$ has a real root greater than one, and $(0,0)$ is unstable.
+
+The stability of $(1,1)$ (i.e. fixation in $A$) is determined by the characteristic polynomial $f_1(x)=ax^2+bx+c$ with coefficients
 \begin{equation}
@@
 a &= 1, \quad
-b &= -(1 - m) (w + 1-\rho + \rho/w), \quad
-c &=  (1-2m)(w(1-\rho)+\rho).
+b &= -(1 - m) (w\rho + 1 - \rho + 1/w), \quad
+c &=  (1 - 2m)(\rho + (1-\rho)/w).
 \end{aligned}
@@
 
-First, the discriminant is 
+First, the discriminant $\Delta = b^2-4ac$ for $m=0$ and $m=1/2$ is
+\begin{equation}\begin{aligned}
+\Delta(m=0) = \Big(\frac{(w \rho + 1) (w-1)}{w}\Big)^2, \text{and} \\
+\Delta(m=1/2) = \Big(\frac{w^2 \rho + w(1-\rho) + 1}{2 w}\Big)^2,
+\end{aligned}\end{equation}
+which are both positive. 
+Writing the discriminant $\Delta$ of $f_1(x)$ as a polynomial in $m$, $g(m)$, the discriminant of $g(m)$ is
 $$
-\Delta = (1 - m)^2 (w + 1-\rho + \rho/w)^2 - 4(1-2m)(w(1-\rho)+\rho),
+\big(- 16 (w\rho + 1)^2 (w - 1)^2 \big(\rho w + 1 - \rho \big) / w^3\big),
 $$
-which is positive for $m=0$ and $m=1/2$:
-\begin{equation}\begin{aligned}
-\Delta(m=0) = \Big(\frac{(w + \rho) (w-1)}{w}\Big)^2 \\
-\Delta(m=1/2) = \Big(\frac{w^2 + w(1-\rho) + \rho}{2 w}\Big)^2.
-\end{aligned}\end{equation}
-In addition, if we write $\Delta$ as a polynomial in $m$, its discriminant is negative -- $- 16 (w + \rho)^2 (w - 1)^2 \big(w (1 - \rho) + \rho \big) / w^2$, and therefore the discriminant $\Delta$ is positive for any $0 \le m \le 1/2$ and $f(\lambda)$ has two real roots.
-\\
-Second, $a>0$, $b<0$, and $c>0$ and stability can be analyzed by the sign of $a+b+c$.
-For $\rho=1$ we already have $a+b+c<0$.
-For $\rho=0$ we have $a+b+c=-m(w-1)<0$.
-Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any valid $\rho$, the leading eigenvalue is greater than one, and $(0,0)$ is unstable.
-
-The stability of $(1,1)$ (i.e. fixation in $A$) is determined by the characteristic polynomial with coefficients
-\begin{equation}
-\begin{aligned}
-a &= 1, \quad
-b &= -(1-m)(w \rho + 1 -\rho + 1/w), \quad
-c &= (1-2m) \big((1-\rho)/w + \rho\big),
-\end{aligned}
-\end{equation}
-%[MWF: Are the roots real?]
-Again, $a>0$, $b<0$, and $c>0$, and if $\rho=1$ then $a+b+c<0$, as before.
-However, for $\rho=0$ we have $a+b+c = \frac{m}{w}(w-1)>0$, in which case the stability is determined by $2a+b=\frac{m(w+1)+w-1}{w}$, the derivative of the polynomial at $\lambda=1$, which is positive, and therefore both eigenvalues occur in $(0,1)$ and fixation is stable.
-%
-
-For any $\rho$, $a+b+c>0$ if and only if
-\begin{equation}
-\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)},
-\end{equation}
-and then $(1,1)$ is stable if and only if $2a+b>0$.
-We have already seen that when $\rho=0$ we have $2a+b>0$.
-When $\rho = \hat \rho$, we have $2a+b=\frac{m(w+1)+w-1}{w(1-m)(w-1)}>0$.
-Since $2a+b$ is linear in $\rho$, we find that $2a+b>0$ for $\hat \rho \ge \rho \ge 0$ and therefore \eqref{eq:migration_model_unconditional_symmetric_condition_rho} guarantees the stable fixation of $A$.
-
-The values of $a+b+c$ at $m=0$ and $m=1$ are $-\frac{\rho(w-1)^2}{w}<0$ and $\frac{(1-\rho)(w-1)}{w}>0$, respectively, and $a+b+c$ is linear in $m$, so that we have $a+b+c>0$ when 
-\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m2}
-m > \hat m = \frac{\rho (w-1)}{\rho (w-1) +1-\rho}.
-\end{equation}
-In addition, $2a+b$ is linear in $m$ and equal to 2 when $m=1$; it is equal to zero when $m=(w-1)(w\rho-1)/(w(w-1)\rho+w+1) < \hat m$, and therefore (\ref{eq:migration_model_unconditional_symmetric_condition_m}, also \ref{eq:migration_model_unconditional_symmetric_condition_m2})
-is sufficient for stability of fixation of $A$.
+which is negative, and therefore $\Delta$ is positive for any $0 \le m \le 1/2$ and $f_1(x)$ has two real roots. \\
+Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and has a positive derivative at infinity.
+Therefore, fixation of $A$ is stable iff both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
+$f_1(1)>0$ occurs if $\rho < \frac{m}{m+(1-m)(w-1)}$.
+$f_1'(1)>0$ occurs if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}$, which is larger than $\frac{m}{m+(1-m)(w-1)}$ for any $w>1$ and $0 \le m \le 1/2$ (the difference is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $w=2$, $m=0$).
+Therefore, fixation of $A$ is stable iff $\rho < \frac{m}{m+(1-m)(w-1)}$ \eqref{eq:migration_model_unconditional_symmetric_condition_rho}, or, by rearranging this inequality, we can also get an expression for $m$ \eqref{eq:migration_model_unconditional_symmetric_condition_m}.
 
@@
 % Acknowledgements
-%{\small
-%\section*{Acknowledgements}
-%
-%This work was supported in part by 
-%the Stanford Center for Computational, Evolutionary and Human Genomics, 
-%and the Morrison Institute for Population and Resources Studies, Stanford University.
-%}
+{\small
+\section*{Acknowledgements}
+
+This work was supported in part by 
+the Stanford Center for Computational, Evolutionary and Human Genomics, 
+and the Morrison Institute for Population and Resources Studies, Stanford University.
+}
 

### 2018-06-10 ms/ms_sunnyvale.tex
fixed math typos suggested by MWF in symmetric periods section

@@
 F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, \\
-F_B(x) = \rho_B \frac{1}{1+s-sx} x' + (1-\rho_B)x.
+F_B(x) = \rho_B \frac{1}{1+s-sx} x + (1-\rho_B)x.
 \end{aligned}\end{equation}
@@
 \end{aligned}\end{equation}
-and if neither of these conditions are met then there exists a \emph{protected polymorphism}.
-Thus for a protected polymorphism, we require
-% TODO check RHS
+A \emph{protected polymorphism} exists if neither of these conditions are met.
+Therefore, for a protected polymorphism to exist, we require
 \begin{equation}
 1 < F'_A(1) F'_B(1) = \Big(1-\rho_A\frac{s}{1+s}\Big)\Big(1+\rho_B s\Big) 
-= 1+\frac{s}{1+s}\big[\rho_B(1+s-\rho_A)-\rho_A\big],
+= 1+\frac{s}{1+s}\big[\rho_B(1+s-s\rho_A)-\rho_A\big],
 \end{equation}

### 2018-06-17 ms/ms_sunnyvale.tex
UL comments

@@
 % Title page
-\title{Vertical and oblique transmission with fluctuating transmission}
+\title{Vertical and oblique transmission fluctuating in time and in space}
 
@@
 \begin{equation}\begin{aligned} \label{eq:recurrence_two_generations}
-x' = \rho_1 \frac{1+s}{\overline w} x + (1-\rho_1)x, \\
-x'' = \rho_2 \frac{1+s}{\overline w} x' + (1-\rho_2)x'.
+x' = \rho_1 \frac{1+s}{\overline w} x + (1-\rho_1)x, \quad \overline w = 1 + xs, \\
+x'' = \rho_2 \frac{1+s}{\overline w'} x' + (1-\rho_2)x', \quad \overline w' = 1 + x's.
 \end{aligned}\end{equation}
@@
 1 < F'_A(0) F'_B(0) = \Big(1+\rho_A s\Big)\Big(1-\rho_B \frac{s}{1+s}\Big) 
-= 1+\frac{s}{1+s}\big[\rho_A(1+s) -\rho_B(1+\rho_A s)\big],
+= 1+\frac{s}{1+s}\big[\rho_A(1+s-s\rho_B)-\rho_B\big],
 \end{equation}
@@
 \begin{equation}\begin{aligned}
-1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b - \Delta \rho s/(1+s)\big)^l,\\
-1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a + \Delta \rho s)^l,
+1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b - s \Delta \rho /(1+s)\big)^l,\\
+1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a + s \Delta \rho)^l,
 \end{aligned}\end{equation}
@@
 \end{equation}
-where $s_t$ are i.i.d (independent and identically distributed), $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$, $\rho_t$ are i.i.d, and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
-Therefore, $z_t = \rho_t s_t$  are independent and identically distributed and $P(-1+C < z_t < D)$. 
+where $z_t=\rho_t s_t$, $s_t$ are i.i.d (independent and identically distributed), $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$, $\rho_t$ are i.i.d, and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
+Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$. 
 From \citet[][results~6 and 7]{Ram2018} we have:
@@
 \paragraph{Examples.}
-First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)=0$) and $s_t$ is symmetric around zero, then $E[z_t]=0$ and $E[-z_t/(1+s_t)]>0$ (because $E[s_t/(1+s_t)] < E[s_t]$).
+First, if $s_t$ and $\rho_t$ are independent ($cov(s_t, \rho_t)=0$) and $s_t$ is symmetric around zero, then $E[-z_t/(1+s_t)] > 0$ (because $E[-z_t/(1+s_t)] = - E[\rho_t] \cdot E[s_t/(1+s_t)]$ and $E[s_t/(1+s_t)] < E[s_t] = 0$).
 Therefore, fixation of $B$ is stochastically locally stable and fixation of $A$ almost surely does not occur.
@@
 
-% TODO full correlation between rho and s?
-
 % Finite population size
@@
 \end{equation}
-where $S_n = \sum_{t=1}^{n}{z_t}$.
+where $S_n = \sum_{t=1}^{n}{z_t}$ and $z_t = \rho_t s_t$.
 Furthermore, combining \eqref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the fixation probability of phenotype $A$, starting from a frequency $x$ of phenotype $A$, is
@@
 
-\autoref{fig:migration_rho} and \autoref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection where $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
+Figures \ref{fig:migration_rho} and \ref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection where $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
 Notably, in the absence of oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
@@
 
-\newpage
-
 \bibliographystyle{agsm}

### 2018-06-20 src/collect_figures.py
update filename

@@ -11,7 +11,7 @@ if __name__ == '__main__':
 	figures_dir = '../figures'
 	ms_figures_dir = join_path(ms_dir, 'figures')
 	zip_fname = '/Users/yoavram/Dropbox/Sunnyvale/figures.zip'
-	tex_files = ['ms.tex']
+	tex_files = ['ms_sunnyvale.tex']
 	pattern = re.compile(r'\\includegraphics(?:\[.*\])?\{(.*\.\w{3})\}')
 
 	if os.path.exists(ms_figures_dir):

### 2018-06-20 ms/ms_sunnyvale.tex
MWF changes via LB

--- 2018-06-10 ms/ms_sunnyvale.tex
@@
 
-Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from adults in the parental generation to offspring) transmission. 
-\citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific trait is transmitted either vertically or obliquely.
+In evolutionary genetics, the properties of transmission are known to be important, e.g., uniparental vs.\ biparental; haploid vs.\ diploid; with or without recombination; sexual or asexual.
+For cultural evolution, the mode of transmission between individuals of the same or different generations is also central to the dynamics of cultural traits.
+%Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from adults in the parental generation to offspring) transmission. 
+\citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, or obliquely, that is, from non-parental members of the parent's generation to an offspring.
 Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time.
-For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-k, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
+For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-$k$, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
 In these studies, the cultural transmission rule did not change during the evolutionary process.
@@
 Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space~(\autoref{fig:transmission}).
-We find that $\ldots$ (to be continued)  %TODO
-We suggest that cultural evolution can be affected by environmental changes in both fitness and transmission fluctuations, and that further empirical and theoretical studies are required to disentangle these effects.
+We find that fluctuations in $\rho$ without corresponding fluctuations in fitness  cannot maintain a polymorphism.
+When both the rate of vertical transmission and fitness fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that  acts on a polymorphism.
+
+We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two demes.
+Stability of fixation points  and a polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.
+We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either fitness or transmission or both, and that further empirical and theoretical studies are required to disentangle these effects on these important evolutionary forces.
 
@@
 
-Suppose that the vertical transmission rate $\rho$ fluctuates, such that $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
+Suppose that the vertical transmission rate $\rho$ fluctuates, with $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
 From \eqref{eq:recurrence}, the recurrence equations for two generations is then
@@
 x' = \rho_1 \frac{1+s}{\overline w} x + (1-\rho_1)x, \\
-x'' = \rho_2 \frac{1+s}{\overline w} x' + (1-\rho_2)x'.
+x'' = \rho_2 \frac{1+s}{\overline w'} x' + (1-\rho_2)x'.
 \end{aligned}\end{equation}
@@
 Suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$, and when $B$ is favored the transmission rate is $\rho_B$.
-The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively,
+The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively, where
 \begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
@@
 \end{equation}
-Surprisingly, the stronger the selection (i.e. the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.
+We can state this as Result 1.
+
+{\bf Result 1.} {\sl Suppose, vertical transmission $A$ occurs at rate $\rho_A$ when its fitness is $1+s$ relative to $1$ for type $B$, and vertical transmission of $B$ occurs at rate $\rho_B$ when the fitness of $B$ is $1+s$ relative to $1$ for $A$.
+Then the stronger the selection (i.e.
+the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.}
 
 \paragraph{Period of one---$A1B1$.}
-In the case of $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
+In the case  $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
 Set $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
@@
 \end{equation}
-\autoref{fig:rho1_rho2_phase_k=1} shows $x^*$, highlighting the area of the parameter space in which a protected polymorphism exists.
+\autoref{fig:rho1_rho2_phase_k=1} shows $x^*$ and highlights the area of the parameter space in which a protected polymorphism exists.
 The figure demonstrates that the stronger the selection (x-axis), the greater the fluctuations in $\rho$  (y-axis) that still allow a polymorphic population (area between the dashed lines).
@@
 Therefore, if we take the i.i.d fitness random variables for $A$ and $B$ to be $w_{A,t}, w_{B,t} \sim U(0,1)$, respectively, and define $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$, then neither $A$ nor $B$ has an advantage, on average (i.e.
-$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to converge to a polymorphic distribution.
+$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to approach a polymorphic distribution.
 
@@
 Therefore, if $s_t$ and $\rho_t$ are positively (negatively) correlated, $S_n/n$ increases (decreases), and the fixation probability $u(x)$ increases (decreases). 
-This occurs because selection only affects those individuals that transmit their phenotype to their own offspring (i.e.
+This occurs because selection  affects only those individuals that transmit their phenotype to their own offspring (i.e.
 vertically), and a fraction $1-\rho_t$ of the population individuals is effectively masked (for better or worse) from selection at each generation.
@@
 
-We now describe a model in which fluctuations in selection and transmission occur in space, rather then time; we model a population divided to two demes.
+We now describe a model in which fluctuations in selection and transmission occur in space, rather then time. Consider a population divided to two demes.
 Selection (e.g.
@@
 
-For the case of symmetric migration $m_1=m_2=m$ ($0\le m \le1/2$) occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+In the case of symmetric migration $m_1=m_2=m$ ($0\le m \le1/2$) occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
-x_1'' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big), \\
-x_2'' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_w} x_2 + 1-\rho_2 \Big).
+x_1'' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_2} x_2 + 1-\rho_2 \Big), \\
+x_2'' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_2} x_2 + 1-\rho_2 \Big).
 \end{aligned}
@@
 which is negative, and therefore the discriminant $\Delta$ of $f_0(x)$ is positive for any $0 \le m \le 1/2$ and $f_0(x)$ has two real roots. \\
-Second, $a>0$, $b<0$, and $c>0$ so $f_0$ is positive with a negative derivative at $x=0$ and has a positive derivative at infinity.
+Second, $a>0$, $b<0$, and $c>0$ so $f_0$ is positive with a negative derivative at $x=0$ and  a positive derivative at infinity.
 Therefore stability of $(0,0)$ can be determined by the sign of $f_0(1)=a+b+c$.
@@
 
-First, the discriminant $\Delta = b^2-4ac$ for $m=0$ and $m=1/2$ is
+First, the discriminant $\Delta$ of $f_1(x)=0$ is $b^2-4ac$ for $m=0$ and $m=1/2$ is
 \begin{equation}\begin{aligned}
@@
 Therefore, fixation of $A$ is stable iff both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
-$f_1(1)>0$ occurs if $\rho < \frac{m}{m+(1-m)(w-1)}$.
-$f_1'(1)>0$ occurs if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}$, which is larger than $\frac{m}{m+(1-m)(w-1)}$ for any $w>1$ and $0 \le m \le 1/2$ (the difference is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $w=2$, $m=0$).
-Therefore, fixation of $A$ is stable iff $\rho < \frac{m}{m+(1-m)(w-1)}$ \eqref{eq:migration_model_unconditional_symmetric_condition_rho}, or, by rearranging this inequality, we can also get an expression for $m$ \eqref{eq:migration_model_unconditional_symmetric_condition_m}.
+Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}$, and
+$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}$, which is larger than $\frac{m}{m+(1-m)(w-1)}$ for any $w>1$ and $0 \le m \le 1/2$ (the difference is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $w=2$, $m=0$).
+Therefore, fixation of $A$ is stable if $\rho < \frac{m}{m+(1-m)(w-1)}$ \eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
+By rearranging this inequality, we can also obtain an expression for $m$ \eqref{eq:migration_model_unconditional_symmetric_condition_m}.
 
@@
 The polymorphism $(x_1^*, x_2^*)$ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, in which case fixation of phenotype $A$ is stable.
-For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the stronger migration is, the less oblique transmission is required to fix $A$ (see shaded areas in \autoref{fig:migration_rho}).
+For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the more migration, the less oblique transmission is required to fix $A$ (see shaded areas in \autoref{fig:migration_rho}).
 Similarly, with enough migration ($m > \hat m$), phenotype $A$ fixes, and the more oblique transmission in deme 2, the less migration is needed to fix $A$ (see shaded area in \autoref{fig:migration_m}).
@@
 Migration rate $m$ is 0.05, 0.1, or 0.25 I the left, middle, and right columns, respectively.
-The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
+The shaded area denotes stable fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
 Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
@@
 Most models of cultural transmission assume a fixed relative rate at which different modes of transmission---vertical, horizontal, or oblique---occur.
-Here we explored a model in which the relative rate of vertical and oblique transmission fluctuates over time, either periodically or randomly, or over space.
-Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ (to be continued) % TODO
+Here we explored a model in which the relative rates of vertical and oblique transmission fluctuate over time, either periodically or randomly, or over space.
+Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ (TO BE CONTINUED) % TODO
 %Specifically, if fluctuations are not small then a single phenotype is likely to fix in the population~(\autoref{eq:poly_condition_periodic_fluc}, \autoref{fig:rho1_rho2_k=1}, \autoref{fig:rho1_rho2_stoch_p}), and
@@
 the Stanford Center for Computational, Evolutionary and Human Genomics, 
-and the Morrison Institute for Population and Resources Studies, Stanford University.
+the Morrison Institute for Population and Resources Studies, Stanford University, and the John Templeton Foundation.
 }
-
-\newpage
 

### 2018-06-20 ms/ms_sunnyvale.tex

@@
 \begin{equation}\begin{aligned} \label{eq:recurrence_two_generations}
-x' = \rho_1 \frac{1+s}{\overline w} x + (1-\rho_1)x, \\
-x'' = \rho_2 \frac{1+s}{\overline w'} x' + (1-\rho_2)x'.
+x' = \rho_1 \frac{1+s}{\overline w} x + (1-\rho_1)x, \quad \overline w = 1 + xs, \\
+x'' = \rho_2 \frac{1+s}{\overline w'} x' + (1-\rho_2)x', \quad \overline w' = 1 + x's.
 \end{aligned}\end{equation}

### 2018-06-20 ms/ms_sunnyvale.tex
added line numbers; revised all

@@
 
+% line numbers
+\usepackage[displaymath, mathlines]{lineno}
+\renewcommand\linenumberfont{\normalfont\small\sffamily}
+\linenumbers
+\modulolinenumbers[2]
+
 % Title page
@@
 For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases.
-%For example, it may be the case that during cold seasons offspring are more likely to stay close to their parents and therefore learn vertically, whereas during warm seasons they are more likely to wander and interact with other adults and therefore learn obliquely.
 
-Here, we analyze a model in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space~(\autoref{fig:transmission}).
-We find that fluctuations in $\rho$ without corresponding fluctuations in fitness  cannot maintain a polymorphism.
-When both the rate of vertical transmission and fitness fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that  acts on a polymorphism.
+Here, we analyze a model~(\autoref{fig:transmission}) in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space.
+We find that fluctuations in $\rho$ without corresponding fluctuations in selection cannot maintain a polymorphism.
+When both the rate of vertical transmission and selection fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that maintains a polymorphism.
+We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two populations.
+Stability of fixation points  and a polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.
 
-We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two demes.
-Stability of fixation points  and a polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.
-We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either fitness or transmission or both, and that further empirical and theoretical studies are required to disentangle these effects on these important evolutionary forces.
+We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either selection or transmission or both, and that further empirical and theoretical studies are required to disentangle these effects.
 
@@
 
-% TODO what can of spatial differences can we imagine in transmission?
+% TODO what kind of spatial differences can we imagine in transmission?
 % two populations, one has first cousin marriages and one doesn't (endogmia)
@@
 
-% Model and Results
-\section*{Model and Results}
+% Models and Results
+\section*{Models and Results}
 
@@
 Suppose that the vertical transmission rate $\rho$ fluctuates, with $\rho = \rho_1$ in odd generations and $\rho = \rho_2$ in even generations.
-From \eqref{eq:recurrence}, the recurrence equations for two generations is then
+From \eqref{eq:recurrence}, the recurrence equations for two generations is
 \begin{equation}\begin{aligned} \label{eq:recurrence_two_generations}
@@
 Second, when $s>0$ ($s<0$) because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ ($<1$) and $x''/x' = 1+\frac{\rho_2 s (1-x')}{1+sx'} > 1$ ($<1$), the frequency of $A$ ($B$) increases every generation and $x^*=1$ ($x^*=0$) is globally stable.
-Therefore, fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype, and there can be no stable polymorphisms.
+Therefore, fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype, and there cannot be a stable polymorphisms.
 
@@
 
-Suppose that both the vertical transmission rate $\rho$ and the selection coefficient $s$ fluctuate together, so that when $A$ is favored the transmission rate is $\rho_A$, and when $B$ is favored the transmission rate is $\rho_B$.
-The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ and $F_B(x)$, respectively, where
+Suppose that both transmission and selection fluctuate together, so that when $A$ is favored, with $w_A=1+s$ and $w_B=1$, the transmission rate is $\rho_A$, and when $B$ is favored, with $w_A=1$ and $w_B=1+s$, the transmission rate is $\rho_B$.
+The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ or $F_B(x)$, respectively, where
 \begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
-F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, \\
-F_B(x) = \rho_B \frac{1}{1+s-sx} x + (1-\rho_B)x.
+&F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, &\quad \text{and} \\
+&F_B(x) = \rho_B \frac{1}{1+s-sx} x + (1-\rho_B)x.
 \end{aligned}\end{equation}
@@
 \begin{equation}\begin{aligned}
-\big[F'_A(1) F'_B(1)\big]^k < 1, \\
+\big[F'_A(1) F'_B(1)\big]^k < 1,  &\quad \text{and}\\
 \big[F'_A(0) F'_B(0)\big]^k < 1. \\
 \end{aligned}\end{equation}
-A \emph{protected polymorphism} exists if neither of these conditions are met.
+By definition, a \emph{protected polymorphism} exists if neither fixation is stable~\citep{Prout1968}, that is, if neither of these conditions are met.
 Therefore, for a protected polymorphism to exist, we require
@@
 
-{\bf Result 1.} {\sl Suppose, vertical transmission $A$ occurs at rate $\rho_A$ when its fitness is $1+s$ relative to $1$ for type $B$, and vertical transmission of $B$ occurs at rate $\rho_B$ when the fitness of $B$ is $1+s$ relative to $1$ for $A$.
-Then the stronger the selection (i.e.
-the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.}
+{\bf Result 1.} {\sl Suppose that vertical transmission occurs at rate $\rho_A$ when the fitness of phenotype $A$ is $1+s$ relative to $1$ for phenotype $B$, and at rate $\rho_B$ when the fitness of phenotype $B$ is $1+s$ relative to $1$ for phenotype $A$.
+Then the stronger the selection (i.e. the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.}
 
 \paragraph{Period of one---$A1B1$.}
-In the case  $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
+In the case  $k=1$ we can find $x^*$ the stable frequency of $A$ at the protected polymorphism.
 Set $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
 \begin{equation} \begin{aligned} \label{eq:xstar_periodic_fluc_k=1}
-G(x) &= ax^2+bx+c, \\
-\gamma &= -\frac{1 + sx}{s} \Big[ 1 + s + s^2 x (1-\rho)(1+x)\Big], \\
-a &= s^2 \rho_A (1-\rho_A) (1 - \rho_B + \rho_B/\rho_A)) > 0, \\
-b &= s(1-\rho_A)(2\rho_B - s\rho_A(1-\rho_B)), \\
-c &= \rho_B - \rho_A - s\rho_A(1-\rho_B).
+&G(x) = ax^2+bx+c, \\
+&\gamma = -\frac{1 + sx}{s} \Big[ 1 + s + s^2 x (1-\rho)(1+x)\Big], \\
+&a = s^2 \rho_A (1-\rho_A) (1 - \rho_B + \rho_B/\rho_A)) > 0, \\
+&b = s(1-\rho_A)(2\rho_B - s\rho_A(1-\rho_B)), \\
+& c = \rho_B - \rho_A - s\rho_A(1-\rho_B).
 \end{aligned} \end{equation}
-% $a+b+c>0$.
-Then $x^*$ is a solution of the quadratic $G(x)=0$. %, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
+Then $x^*$ is a solution of the quadratic $G(x)=0$.
 The condition \eqref{eq:poly_condition_periodic_fluc}, which guarantees that $0 < x^* < 1$, is equivalent to $c<0$, and therefore it also guarantees that $\sqrt{b^2-4ac} > b$.
@@
 \textbf{Protected polymorphism.}
-The stable equilibrium of the frequency of phenotype $A$ (from eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
-Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from~\eqref{eq:poly_condition_periodic_fluc} that allow protected polymorphism.
+The stable equilibrium of the frequency of phenotype $A$ (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
+Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from eq.~\ref{eq:poly_condition_periodic_fluc} that permit a protected polymorphism.
 Here, $\rho_A=0.5$.}
@@
 \begin{equation}\begin{aligned}
-1 &< (F'_A(0))^k (F'_B(0))^l = a^k \big(b - s \Delta \rho /(1+s)\big)^l,\\
-1 &< (F'_A(1))^k (F'_B(1))^l = b^k (a + s \Delta \rho)^l,
+& 1 < (F'_A(0))^k (F'_B(0))^l = a^k \big(b - s \Delta \rho /(1+s)\big)^l, \quad \text{and} \\
+& 1 < (F'_A(1))^k (F'_B(1))^l = b^k (a + s \Delta \rho)^l,
 \end{aligned}\end{equation}
@@
 \end{equation}
-Therefore, for a given value of $\rho_A$, if $\rho_B>\rho_A$ then increasing the vertical transmission rate of $B$ decreases the environmental period ratio $l/k$ that permits protected polymorphism; decreasing $\rho_B$  will have the opposite effect; it increases the ratio $l/k$ that permits a protected polymorphism.
+Therefore, for a given value of $\rho_A$, if $\rho_B>\rho_A$ then increasing the vertical transmission rate of $B$ decreases the environmental period ratio $l/k$ that permits a protected polymorphism; decreasing $\rho_B$  will have the opposite effect: it increases the ratio $l/k$ that permits a protected polymorphism.
 
@@
 \begin{equation}
-x_t = x_t \frac{1 + z_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
+x_{t+1} = x_t \frac{1 + z_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
 \end{equation}
-where $z_t=\rho_t s_t$, $s_t$ are i.i.d (independent and identically distributed), $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$, $\rho_t$ are i.i.d, and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
+where $z_t=\rho_t s_t$; $s_t$ are i.i.d (independent and identically distributed); $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$; $\rho_t$ are i.i.d; and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
 Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$.
-From \citet[][results~6 and 7]{Ram2018} we have:
+
+{\it Definition: ``stochastic local stability''.}
+A constant equilibrium state $x^*$ is said to be \emph{stochastically locally stable} if for any $\epsilon>0$ there exists a $\delta>0$ such that $|x_0-x^*|<\delta$ implies
+\begin{equation}
+P(\lim_{t \to \infty}x_t = x^*) \ge 1-\epsilon.
+\end{equation}
+Thus stochastic local stability holds for $x^*$ provided for any initial $x_0$ sufficiently near $x^*$ the process $x_t$ converges to $x^*$ with high probability.
+
+In our case there are two constant equilibria $x^*=0$ and $x^*=1$ corresponding to fixation in $B$ and $A$, respectively.
+From \citet[][results~6 and 7]{Ram2018} we can characterize the stochastic local stability of these fixations as follows.
+
 \begin{itemize}
@@
 
-Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, there can be fixation.
+Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, a fixation can occur.
 Let $w_{A,t}, w_{B,t} \sim U(0,1)$, $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$ and $\rho_t \sim Beta(1+s_t, 1)$ (a beta distribution with parameters $1+s_t$ and $1$). 
-The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$; estimated by averaging $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
+The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$; estimated by averaging over $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
 $s_t>0$) and oblique transmission is more  likely when $B$ is favored (i.e.
@@
 
-Third, it is also possible that both fixations are not stochastically locally stable even if $s_t$ and $\rho_t$ covary, but, as in the case of periodic fluctuations, this can only occur if the fluctuations in $\rho_t$ are small.
+Third, it is also possible that both fixations are not stochastically locally stable even if $s_t$ and $\rho_t$ covary, but, as in the case of periodic fluctuations, this can only occur if fluctuations in $\rho_t$ are small.
 For example, \autoref{fig:rho1_rho2_stoch_p} shows the expected outcome when $s_t=s$ and $\rho_t=\rho_1$ with probability $p=0.505$, while $s_t=-s$ and $\rho_t=\rho_2$ with probability $1-p=0.495$.
@@
 \begin{equation} \label{eq:drift_diffusion_terms}
-\mu(x) = S_n x(1-x)
+\mu(x) = S_n x(1-x),
 \quad \text{and} \quad
@@
 where $S_n = \sum_{t=1}^{n}{z_t}$ and $z_t = \rho_t s_t$.
-Furthermore, combining \eqref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the fixation probability of phenotype $A$, starting from a frequency $x$ of phenotype $A$, is
+Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability for fixation of phenotype $A$ when its initial frequency if $x$, is
 \begin{equation}
@@
 \end{equation}
-From result 10 in \citep{Ram2018}, $u(x)$ is monotone increasing in $S_n/n$ which is the average selection coefficient of $A$ weighted by the vertical transmission rates $\rho_t$.
+From result 10 in \citet{Ram2018}, $u(x)$ is monotone increasing in $S_n/n$ which is the average selection coefficient of $A$ weighted by the vertical transmission rates $\rho_t$.
 Therefore, if $s_t$ and $\rho_t$ are positively (negatively) correlated, $S_n/n$ increases (decreases), and the fixation probability $u(x)$ increases (decreases). 
 This occurs because selection  affects only those individuals that transmit their phenotype to their own offspring (i.e.
-vertically), and a fraction $1-\rho_t$ of the population individuals is effectively masked (for better or worse) from selection at each generation.
+vertically), and a fraction $1-\rho_t$ of the population is effectively masked (for better or worse) from selection at each generation.
 
@@
 \end{equation}
-where $w_j$ is the fitness of phenotype $A$ in deme $j$ relative to the fitness of phenotype $B$; $\overline{w}_j=w_j x_j + (1-x_j)$ is the population mean fitness in deme $j$; and $\rho_j$ is the vertical transmission rate in deme $j$.
+where $w_j$ is the fitness of phenotype $A$ in deme $j$ relative to the fitness of phenotype $B$; $\overline{w}_j=w_j x_j + (1-x_j)$ is the mean fitness in deme $j$; and $\rho_j$ is the vertical transmission rate in deme $j$.
 
 Following migration, the frequencies of $A$ in the two demes are
-\begin{equation} \label{eq:migration_model}
+\begin{equation} \label{eq:migration_model_migration}
 \begin{aligned}
@@
 
-In the case of symmetric migration $m_1=m_2=m$ ($0\le m \le1/2$) occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model} become
+In the case of symmetric migration $m_1=m_2=m$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 Here we explored a model in which the relative rates of vertical and oblique transmission fluctuate over time, either periodically or randomly, or over space.
-Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that $\ldots$ (TO BE CONTINUED) % TODO
-%Specifically, if fluctuations are not small then a single phenotype is likely to fix in the population~(\autoref{eq:poly_condition_periodic_fluc}, \autoref{fig:rho1_rho2_k=1}, \autoref{fig:rho1_rho2_stoch_p}), and
 
-One caveat of our model is the use of the ``phenotypic gambit''---the assumption that the transmission mode itself is strictly vertically transmitted.
+Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that a protected polymorphism can be maintained only if fluctuations in the rate of vertical transmission are small, and that stronger selection permits larger fluctuations while still maintaining a protected polymorphism.
+In the case of fluctuations in space, the greater the separation between the two populations, the larger the fluctuations can be while still maintaining polymorphism; however, as migration becomes more frequent, even small differences in the vertical transmission rate will eliminate the polymorphism.
+When fluctuations are stochastic, we find that if vertical transmission covaries with selection, the phenotype that is more likely to be vertically transmitted specifically when it is favored will likely fix in the population, eventually.
+However, if transmission and selection are independent, then a polymorphism can be maintained if selection doesn't, on average, favor one phenotype over the other.
+
+One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
 Although there is evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
@@
 
-An extension of our model could incorporate more sophisticated oblique transmission schemes.
+An extension of our model could incorporate more sophisticated oblique transmission schemes~\citep[][Figure~3]{Creanza2017}.
 For example, conformity---preference for learning a frequent phenotype---has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.

### 2018-06-20 ms/ms_sunnyvale.tex
MWF changes to results summary in discussion

@@
 
-Comparing our results with previous results from a similar model with a fixed rate~\citep{Ram2018}, we find that a protected polymorphism can be maintained only if fluctuations in the rate of vertical transmission are small, and that stronger selection permits larger fluctuations while still maintaining a protected polymorphism.
+Comparing our results with previous results from a similar model with a fixed rate of vertical transmission~\citep{Ram2018}, we find that a protected polymorphism can be maintained only if fluctuations in the rate of vertical transmission are small, and that stronger selection on the transmitted trait permits larger fluctuations in the rate of transmission while still maintaining a protected polymorphism.
 In the case of fluctuations in space, the greater the separation between the two populations, the larger the fluctuations can be while still maintaining polymorphism; however, as migration becomes more frequent, even small differences in the vertical transmission rate will eliminate the polymorphism.
-When fluctuations are stochastic, we find that if vertical transmission covaries with selection, the phenotype that is more likely to be vertically transmitted specifically when it is favored will likely fix in the population, eventually.
-However, if transmission and selection are independent, then a polymorphism can be maintained if selection doesn't, on average, favor one phenotype over the other.
+When fluctuations are stochastic, we find that if vertical transmission covaries with selection, the phenotype that has a higher probability of being vertically transmitted  when it is favored will likely eventually fix in the population.
+However, if transmission and selection are independent, then a polymorphism can be maintained if selection does not, on average, favor one phenotype over the other.
 

### 2018-06-26 ms/ms_sunnyvale.tex
fix general migration model eq. 19

@@
 
-We now describe a model in which fluctuations in selection and transmission occur in space, rather then time. Consider a population divided to two demes.
+We now describe a model in which fluctuations in selection and transmission occur in space, rather then time.
+Consider a population divided to two demes.
 Selection (e.g.
@@
 \begin{aligned}
-x_1'' &= (1-m_1) x_1' + m_2 x_2', \\
-x_2'' &= m_1 x_1' + (1-m_2) x_2',
+\gamma_1 x_1'' &= (1-m) x_1' + c m x_2', \quad \gamma_1 = (1-m) + c m, \\
+\gamma_2 x_2'' &= m x_1' + c (1-m) x_2', \quad \gamma_2 = m + c (1-m),
 \end{aligned}
 \end{equation}
-where $0 \le m_j \le 1/2$ is the migration rate out of deme $j$.
+where $0 \le m \le 1/2$ is the migration rate, i.e.
+the fraction of the population that is replaced by migrants; $c>0$ is the relative size of deme 2 compared to deme 1 (at least where migration is concerned); and $\gamma_1$ is a normalizing factor.
+This is a \emph{two-deme stepping-stone migration scheme}~\citep[][eqs.~2.15-16]{Karlin1982}.
 
@@
 
-In the case of symmetric migration $m_1=m_2=m$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
+%todo
+%asymmetric migration, check x*=(0,0) and (1,1)
+%do simulations to find other stable points
+%how do m1/m2 and rho1/rho2 balance each other in terms of polymorphism etc.
+
+In the case of symmetric migration $c=1$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
@@
 \end{equation}
+This is the \emph{homogeneous stepping-stone migration scheme}~\citep[][eq.~2.14]{Karlin1982}.
 
@@
 The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
-Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
+Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
 The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2$ (x-axis) in deme 2.
-Migration rate $m$ is 0.05, 0.1, or 0.25 I the left, middle, and right columns, respectively.
+Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
 The shaded area denotes stable fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
@@
 The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
-Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype 2 is $1$ in both demes).
+Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
 The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.4$, and $0.2$, in the left, middle, and right columns, respectively, in deme 2.

### 2018-06-27 ms/ms_sunnyvale.tex
cite karlin1981
fixed typos MWF found; changed todo

@@
 
-Following migration, the frequencies of $A$ in the two demes are
+Following migration~\citep{Karlin1981}, the frequencies of $A$ in the two demes are
 \begin{equation} \label{eq:migration_model_migration}
@@
 where $0 \le m \le 1/2$ is the migration rate, i.e.
-the fraction of the population that is replaced by migrants; $c>0$ is the relative size of deme 2 compared to deme 1 (at least where migration is concerned); and $\gamma_1$ is a normalizing factor.
+the fraction of the population that is replaced by migrants; $c>0$ is the relative size of deme 2 compared to deme 1 (at least where migration is concerned); and $\gamma_j$ is a normalizing factor.
 This is a \emph{two-deme stepping-stone migration scheme}~\citep[][eqs.~2.15-16]{Karlin1982}.
 
+%todo
+%do simulations to find stable polymorphisms
+%how do m1/m2 and rho1/rho2 balance each other in terms of polymorphism
 \paragraph{Symmetric migration and selection.}
-
-%todo
-%asymmetric migration, check x*=(0,0) and (1,1)
-%do simulations to find other stable points
-%how do m1/m2 and rho1/rho2 balance each other in terms of polymorphism etc.
 
@@
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
-x_1'' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_2} x_2 + 1-\rho_2 \Big), \\
-x_2'' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} x_1 + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_2} x_2 + 1-\rho_2 \Big).
+x_1'' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_2} + 1-\rho_2 \Big), \\
+x_2'' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_2} + 1-\rho_2 \Big).
 \end{aligned}

### 2018-07-02 ms/ms_sunnyvale.tex
changed superscripts of x in migration model

@@
 \begin{equation} \label{eq:migration_model_selection_transmission}
-x_j' = \rho_j \frac{w_j}{\overline{w}_j} x_j + (1-\rho_j) x_j,
+x_j^s = \rho_j \frac{w_j}{\overline{w}_j} x_j + (1-\rho_j) x_j,
 \end{equation}
@@
 \begin{aligned}
-\gamma_1 x_1'' &= (1-m) x_1' + c m x_2', \quad \gamma_1 = (1-m) + c m, \\
-\gamma_2 x_2'' &= m x_1' + c (1-m) x_2', \quad \gamma_2 = m + c (1-m),
+\gamma_1 x_1' &= (1-m) x_1^s + c m x_2^s, \quad \gamma_1 = (1-m) + c m, \\
+\gamma_2 x_2' &= m x_1^s + c (1-m) x_2^s, \quad \gamma_2 = m + c (1-m),
 \end{aligned}
@@
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}
-x_1'' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_2} + 1-\rho_2 \Big), \\
-x_2'' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_2} + 1-\rho_2 \Big).
+x_1' &= (1-m)x_1\Big(\rho_1 \frac{w}{\overline w_1} + 1-\rho_1 \Big) + m x_2\Big(\rho_2 \frac{1/w}{\overline w_2} + 1-\rho_2 \Big), \\
+x_2' &= m x_1\Big(\rho_1 \frac{w}{\overline w_1} + 1-\rho_1 \Big) + (1-m) x_2\Big(\rho_2 \frac{1/w}{\overline w_2} + 1-\rho_2 \Big).
 \end{aligned}

### 2018-07-02 ms/ms_sunnyvale.tex
c->tau because c is already used for poly coef

@@
 \begin{aligned}
-\gamma_1 x_1' &= (1-m) x_1^s + c m x_2^s, \quad \gamma_1 = (1-m) + c m, \\
-\gamma_2 x_2' &= m x_1^s + c (1-m) x_2^s, \quad \gamma_2 = m + c (1-m),
+\gamma_1 x_1' &= (1-m) x_1^s + \tau m x_2^s, \quad \gamma_1 = (1-m) + \tau m, \\
+\gamma_2 x_2' &= m x_1^s + \tau (1-m) x_2^s, \quad \gamma_2 = m + \tau (1-m),
 \end{aligned}
@@
 where $0 \le m \le 1/2$ is the migration rate, i.e.
-the fraction of the population that is replaced by migrants; $c>0$ is the relative size of deme 2 compared to deme 1 (at least where migration is concerned); and $\gamma_j$ is a normalizing factor.
+the fraction of the population that is replaced by migrants; $\tau>0$ is the relative size of deme 2 compared to deme 1 (at least where migration is concerned); and $\gamma_j$ is a normalizing factor.
 This is a \emph{two-deme stepping-stone migration scheme}~\citep[][eqs.~2.15-16]{Karlin1982}.
@@
 
-In the case of symmetric migration $c=1$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
+In the case of symmetric migration $\tau=1$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}

### 2018-07-02 ms/ms_sunnyvale.tex
stability for asymetric migration

--- 2018-07-02 ms/ms_sunnyvale.tex
@@
 \begin{aligned}
-\gamma_1 x_1' &= (1-m) x_1^s + c m x_2^s, \quad \gamma_1 = (1-m) + c m, \\
-\gamma_2 x_2' &= m x_1^s + c (1-m) x_2^s, \quad \gamma_2 = m + c (1-m),
+\gamma_1 x_1' &= (1-m_1) x_1^s + m_1 x_2^s, \\
+\gamma_2 x_2' &= m_2 x_1^s + (1-m_2) x_2^s,
 \end{aligned}
 \end{equation}
-where $0 \le m \le 1/2$ is the migration rate, i.e.
-the fraction of the population that is replaced by migrants; $c>0$ is the relative size of deme 2 compared to deme 1 (at least where migration is concerned); and $\gamma_j$ is a normalizing factor.
-This is a \emph{two-deme stepping-stone migration scheme}~\citep[][eqs.~2.15-16]{Karlin1982}.
+where $0 \le m_1, m_2 \le 1/2$ are the migration rates, such that $m_1$ is the fraction of the population of deme 1 replaced by individuals from deme 2, and vice versa for $m_2$.
+This is a \emph{two-deme stepping-stone migration scheme}~\citep[][eq.~2.17]{Karlin1982}.
+Analysis of this general model is difficult, though some analytical results may be attained~\citep[see~Principle~6.1]{Karlin1982}.
 
-%todo
-%do simulations to find stable polymorphisms
-%how do m1/m2 and rho1/rho2 balance each other in terms of polymorphism
+\paragraph{Asymmetric migration and symmetric selection.}
+
+\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection symmetric selection $w_1=1/w_2=w>1$.
+We focus on composite parameters of the model:
+$m_1/m_2$, on the x-axis, is the ratio of migration rates into deme 1 and deme 2; when this ratio is large, deme 1 accepts more migrants than deme 2.
+$\rho_1/\rho_2$, on the y-axis the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.
+
+The results demonstrate that fixation of phenotype $A$ is stable if migration to and oblique transmission within deme 1, where $A$ is favored, are higher than in deme 2.
+The vice-versa is true for phenotype $B$.
+A protected polymorphism exists if neither fixation is stable: if migration ratios are positively correlated -- vertical transmission occurs more often in the deme that accepts more migrants -- or if both ratios are close to unity, that is, the differences between the demes in terms of migration and transmission are small (\autoref{fig:asym_migration}).
+
+\begin{figure*}[hbt]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/asym_migration.pdf}
+\caption{
+\textbf{Oblique transmission and asymmetric migration.}
+Classification of the stable equilibrium of the system in eqs.~\ref{eq:migration_model_selection_transmission}-\ref{eq:migration_model_migration} for different ratios of the migration rates (x-axis) and vertical transmission rates (y-axis) in the two demes.
+Stability was determined for 10,000 random choices of $m_1$, $m_2$, $\rho_1$, and $\rho_2$ by calculating the leading eigenvalue of the Jacobian of the system.
+Blue markers denote cases in which the leading eigenvalue of the Jacobian at $x_1=x_2=0$ was less than 1, leading to fixation of $B$.
+Red markers denote cases in which the leading eigenvalue of the Jacobian at $x_1=x_2=1$ was less than 1, leading to fixation of $A$.
+Greens markers denote cases in which both leading eigenvalues were larger than 1, leading to a protected polymorphism.
+Here, the fitness values are $w_1=1/w_2=2$.
+}
+\label{fig:asym_migration}
+\end{figure*}
+
 \paragraph{Symmetric migration and selection.}
 
-In the case of symmetric migration $c=1$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
+In the case of symmetric migration $m_1=m_2=m$ occurring after symmetric selection $w_1=1/w_2=w>1$, the recursions~\eqref{eq:migration_model_selection_transmission}~and~\eqref{eq:migration_model_migration} become
 \begin{equation}\begin{aligned} \label{eq:migration_model_unconditional_symmetric}

### 2018-07-05 ms/ms_sunnyvale.tex
removed gammas from eq 19
MWF changes

@@
 a4paper,
-total={170mm,257mm},
+total={170mm,252mm},
 left=20mm,
@@
 \usepackage[round,colon,authoryear]{natbib}
+\usepackage{xcolor}
 
@@
 % Title page
-\title{Vertical and oblique transmission fluctuating in time and in space}
+\title{Vertical and oblique {\color{red}cultural} transmission fluctuating in time and in space}
 
@@
 %Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from adults in the parental generation to offspring) transmission. 
-\citet[ch.~3]{Cavalli-Sforza1981} first introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, or obliquely, that is, from non-parental members of the parent's generation to an offspring.
+{\color{red}Using analysis with epidemic theory}, \citet[ch.~3]{Cavalli-Sforza1981}  introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, obliquely, that is, from non-parental members of the parent's generation to an offspring{\color{red}, or horizontally, that is, among members of the same generational cohort.}
 Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time.
@@
 Phenotypes are transmitted vertically with probability $\rho$ or obliquely with probability $1-\rho$~(\autoref{fig:transmission}).
-Given $x$ the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is
+Given $x$, the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is
 \begin{equation} \label{eq:recurrence}
@@
 Second, when $s>0$ ($s<0$) because $x'/x = 1+\frac{\rho_1 s (1-x)}{1+sx} > 1$ ($<1$) and $x''/x' = 1+\frac{\rho_2 s (1-x')}{1+sx'} > 1$ ($<1$), the frequency of $A$ ($B$) increases every generation and $x^*=1$ ($x^*=0$) is globally stable.
-Therefore, fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype, and there cannot be a stable polymorphisms.
+Therefore, fluctuations in the mode of transmission ($\rho$) without fluctuations in selection lead to fixation of the favored phenotype, and there cannot be a stable polymorphism.
 
@@
 The stable equilibrium of the frequency of phenotype $A$ (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
-Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from eq.~\ref{eq:poly_condition_periodic_fluc} that permit a protected polymorphism.
+Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from {\color{red}inequalities~(\ref{eq:poly_condition_periodic_fluc})} that permit a protected polymorphism.
 Here, $\rho_A=0.5$.}
@@
 \item In particular, if $E[z_t] = cov(\rho_t, s_t) + E[\rho_t] E[s_t] \le 0$ then $x^*=0$ is stochastically locally stable, and similarly if $E[-z_t/(1+s_t)] = cov(\rho_t, -s_t/(1+s_t)) - E[\rho_t] E[s_t/(1+s_t)] \le 0$ then $x^*=1$ is stochastically locally stable.
-\item It is not possible that $E[log(1+z_t)]$ and $E[log(1-z_t/(1+s_t)]$ are both negative, as their sum is positive
-\begin{multline}
-E[log(1+z_t)] + E[log(1-z_t/(1+s_t)] = \\
-E[log(1+z_t) + log(1-z_t/(1+s_t)]= \\
-E[log\big((1+z_t)(1-z_t/(1+s_t)\big)]= \\
-E[log\big( 1+\rho_t(1-\rho_t)s_t^2/(1+s_t) \big)] > 0,
-\end{multline}
+\item It is not possible that $E[log(1+z_t)]$ and $E[log(1-z_t/(1+s_t)]$ are both negative, as their sum is positive:
+\begin{equation}\begin{aligned}
+{}&E[log(1+z_t)] + E[log(1-z_t/(1+s_t)]  \\
+={} &E[log(1+z_t) + log(1-z_t/(1+s_t)] \\
+={} &E[log\big((1+z_t)(1-z_t/(1+s_t)\big)] \\
+={} &E[log\big( 1+\rho_t(1-\rho_t)s_t^2/(1+s_t) \big)] > 0,
+\end{aligned}\end{equation}
 and therefore it is not possible that both fixations are stochastically locally stable.
@@
 Let $w_{A,t}, w_{B,t} \sim U(0,1)$, $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$ and $\rho_t \sim Beta(1+s_t, 1)$ (a beta distribution with parameters $1+s_t$ and $1$). 
-The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$; estimated by averaging over $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
+The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$ {\color{red}as} estimated by averaging over $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
 $s_t>0$) and oblique transmission is more  likely when $B$ is favored (i.e.
@@
 where $S_n = \sum_{t=1}^{n}{z_t}$ and $z_t = \rho_t s_t$.
-Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability for fixation of phenotype $A$ when its initial frequency if $x$, is
+Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability {\color{red}of} fixation of phenotype $A$ when its initial frequency if $x$, is
 \begin{equation}
@@
 \begin{aligned}
-\gamma_1 x_1' &= (1-m_1) x_1^s + m_1 x_2^s, \\
-\gamma_2 x_2' &= m_2 x_1^s + (1-m_2) x_2^s,
+x_1' &= (1-m_1) x_1^s + m_1 x_2^s, \\
+x_2' &= m_2 x_1^s + (1-m_2) x_2^s,
 \end{aligned}
@@
 The results demonstrate that fixation of phenotype $A$ is stable if migration to and oblique transmission within deme 1, where $A$ is favored, are higher than in deme 2.
-The vice-versa is true for phenotype $B$.
+The {\color{red}opposite} is true for phenotype $B$.
 A protected polymorphism exists if neither fixation is stable: if migration ratios are positively correlated -- vertical transmission occurs more often in the deme that accepts more migrants -- or if both ratios are close to unity, that is, the differences between the demes in terms of migration and transmission are small (\autoref{fig:asym_migration}).
@@
 Red markers denote cases in which the leading eigenvalue of the Jacobian at $x_1=x_2=1$ was less than 1, leading to fixation of $A$.
-Greens markers denote cases in which both leading eigenvalues were larger than 1, leading to a protected polymorphism.
+Green markers denote cases in which both leading eigenvalues were larger than 1, leading to a protected polymorphism.
 Here, the fitness values are $w_1=1/w_2=2$.
@@
 \begin{itemize}
-\item With only oblique transmission ($\rho_1=\rho_2=0$), the only equilibria are the neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
+\item With only oblique transmission ($\rho_1=\rho_2=0$), {\color{red}there are only} neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
 \item With only vertical transmission ($\rho_1=\rho_2=1$), the fixation equilibria  $(0,0)$, $(1,1)$ are unstable and there exists a protected polymorphism
@@
 \end{itemize}
+{\color{red}The proofs of (22) and (23) are in Appendix I.}
+\paragraph{Examples.}
 
-\paragraph{Proof.}
+Figures \ref{fig:migration_rho} and \ref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection where $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
+Notably, in the absence of oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
+With some oblique transmission, but equal in both demes, results are similar (not shown).
+However, if oblique transmission is stronger in deme 2 than in deme 1 (\autoref{fig:migration_m}, middle and right columns), the stable frequency of $A$ increases in both demes.
+Therefore, the mean fitness in deme 1 decreases to a lesser extent than in deme 2, and even increases when the migration rate is high enough.
+
+The polymorphism $(x_1^*, x_2^*)$ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, in which case fixation of phenotype $A$ is stable.
+For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the more migration, the less oblique transmission is required to fix $A$ (see shaded areas in \autoref{fig:migration_rho}).
+Similarly, with enough migration ($m > \hat m$), phenotype $A$ fixes, and the more oblique transmission in deme 2, the less migration is needed to fix $A$ (see shaded area in \autoref{fig:migration_m}).
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/migration_rho.pdf}
+\caption{
+\textbf{Oblique transmission and migration: effect of transmission.} 
+The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
+Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
+The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2$ (x-axis) in deme 2.
+Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
+The shaded area denotes stable fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
+Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
+}
+\label{fig:migration_rho}
+\end{figure*}
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/migration_m.pdf}
+\caption{
+\textbf{Oblique transmission and migration: effect of migration.} 
+The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\bar{w}^*_i$, the population mean fitnesses (bottom row), in the two demes.
+Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
+The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.4$, and $0.2$, in the left, middle, and right columns, respectively, in deme 2.
+Migration rate $m$ is on the x-axis.
+The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
+Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
+}
+\label{fig:migration_m}
+\end{figure*}
+
+% Discussion
+\section*{Discussion}
+
+Most models of cultural transmission assume a fixed relative rate at which different modes of transmission---vertical, horizontal, or oblique---occur.
+Here we explored a model in which the relative rates of vertical and oblique transmission fluctuate over time, either periodically or randomly, or over space.
+
+Comparing our results with previous results from a similar model with a fixed rate of vertical transmission~\citep{Ram2018}, we find that a protected polymorphism can be maintained only if fluctuations in the rate of vertical transmission are small, and that stronger selection on the transmitted trait permits larger fluctuations in the rate of transmission while still maintaining a protected polymorphism.
+In the case of fluctuations in space, the greater the separation between the two populations {\color{red}(i.e., the smaller the migration rates)} the larger {\color{red} can be the fluctuations that maintain} polymorphism; however, as migration becomes more frequent, even small differences in the vertical transmission rate will eliminate the polymorphism.
+When fluctuations are stochastic, we find that if vertical transmission covaries with selection, the phenotype that has a higher probability of being vertically transmitted  when it is favored will likely eventually fix in the population.
+However, if transmission and selection are independent, then a polymorphism can be maintained if selection does not, on average, favor one phenotype over the other.
+
+One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
+Although there is {\color{red}some} evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
+Indeed, it has been demonstrated that guppy fish are more likely to learn from others if previous social experiences provided benefits~\citep{Leris2016}.
+It is also possible that the transmission mode is regulated.
+For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
+Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only some individuals do so.
+
+An extension of our model could incorporate more sophisticated oblique transmission schemes~\citep[][Figure~3]{Creanza2017}.
+For example, conformity---preference for learning a frequent phenotype---has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
+We suggest that the specific mode of oblique transmission can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions.
+Additional work will be required to understand how such fluctuations affect the population dynamics.
+
+% Facebook rural internet 
+% National service - Israel, Singapore
+% Boarding schools
+% Immigrtion - draught
+% Opening schools, universities
+% sending students abroad, bringing experts from abroad
+% foreign labor
+% refugees
+% hunter-gatherers/chimps increase habitat in years with less food production
+% religion - dark ages etc.
+
+% Acknowledgements
+{\small
+\section*{Acknowledgements}
+
+This work was supported in part by 
+the Stanford Center for Computational, Evolutionary and Human Genomics, 
+the Morrison Institute for Population and Resources Studies, Stanford University, and the John Templeton Foundation.
+}
+
+\bibliographystyle{agsm}
+%\bibliography{/Users/yoavram/Documents/library}
+\bibliography{ms_sunnyvale}
+
+\section*{Appendix I}
+\paragraph{Proof of (22) and (23).}
 
@@
 which is negative, and therefore $\Delta$ is positive for any $0 \le m \le 1/2$ and $f_1(x)$ has two real roots. \\
-Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and has a positive derivative at infinity.
+Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and  a positive derivative at infinity.
 Therefore, fixation of $A$ is stable iff both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
 Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}$, and
-$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}$, which is larger than $\frac{m}{m+(1-m)(w-1)}$ for any $w>1$ and $0 \le m \le 1/2$ (the difference is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $w=2$, $m=0$).
+$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}$, which is larger than $\frac{m}{m+(1-m)(w-1)}$ for any $w>1$ and $0 \le m \le 1/2$ (the difference{[\color{red}WHAT DIFFERENCE?]} is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $w=2$, $m=0$).
 Therefore, fixation of $A$ is stable if $\rho < \frac{m}{m+(1-m)(w-1)}$ \eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
@@
 
-\paragraph{Examples.}
-
-Figures \ref{fig:migration_rho} and \ref{fig:migration_m} show the stable frequencies of phenotype $A$~ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) and the stable population mean fitnesses in the two demes with symmetric selection where $A$ is favored in deme 1 and $B$ is favored in deme 2 with similar selection intensities.
-Notably, in the absence of oblique transmission (\autoref{fig:migration_m}, left column), migration decreases the differences between the demes and reduces the population mean fitnesses.
-With some oblique transmission, but equal in both demes, results are similar (not shown).
-However, if oblique transmission is stronger in deme 2 than in deme 1 (\autoref{fig:migration_m}, middle and right columns), the stable frequency of $A$ increases in both demes.
-Therefore, the mean fitness in deme 1 decreases to a lesser extent than in deme 2, and even increases when the migration rate is high enough.
-
-The polymorphism $(x_1^*, x_2^*)$ (eq.~\ref{eq:migration_model_unconditional_symmetric_polymorphism}) is protected when transmission rates are equal, but not when transmission rates differ enough and migration is strong enough, in which case fixation of phenotype $A$ is stable.
-For example, with enough oblique transmission ($\rho_2<\hat \rho$) in deme 2, phenotype $A$ fixes, and the more migration, the less oblique transmission is required to fix $A$ (see shaded areas in \autoref{fig:migration_rho}).
-Similarly, with enough migration ($m > \hat m$), phenotype $A$ fixes, and the more oblique transmission in deme 2, the less migration is needed to fix $A$ (see shaded area in \autoref{fig:migration_m}).
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=0.75\linewidth]{../figures/migration_rho.pdf}
-\caption{
-\textbf{Oblique transmission and migration: effect of transmission.} 
-The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
-Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
-The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2$ (x-axis) in deme 2.
-Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
-The shaded area denotes stable fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
-Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
-}
-\label{fig:migration_rho}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=0.75\linewidth]{../figures/migration_m.pdf}
-\caption{
-\textbf{Oblique transmission and migration: effect of migration.} 
-The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
-Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
-The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.4$, and $0.2$, in the left, middle, and right columns, respectively, in deme 2.
-Migration rate $m$ is on the x-axis.
-The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
-Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
-}
-\label{fig:migration_m}
-\end{figure*}
-
-% Discussion
-\section*{Discussion}
-
-Most models of cultural transmission assume a fixed relative rate at which different modes of transmission---vertical, horizontal, or oblique---occur.
-Here we explored a model in which the relative rates of vertical and oblique transmission fluctuate over time, either periodically or randomly, or over space.
-
-Comparing our results with previous results from a similar model with a fixed rate of vertical transmission~\citep{Ram2018}, we find that a protected polymorphism can be maintained only if fluctuations in the rate of vertical transmission are small, and that stronger selection on the transmitted trait permits larger fluctuations in the rate of transmission while still maintaining a protected polymorphism.
-In the case of fluctuations in space, the greater the separation between the two populations, the larger the fluctuations can be while still maintaining polymorphism; however, as migration becomes more frequent, even small differences in the vertical transmission rate will eliminate the polymorphism.
-When fluctuations are stochastic, we find that if vertical transmission covaries with selection, the phenotype that has a higher probability of being vertically transmitted  when it is favored will likely eventually fix in the population.
-However, if transmission and selection are independent, then a polymorphism can be maintained if selection does not, on average, favor one phenotype over the other.
-
-One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
-Although there is evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
-Indeed, it has been demonstrated that guppy fish are more likely to learn from others if previous social experiences provided benefits~\citep{Leris2016}.
-It is also possible that the transmission mode is regulated.
-For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
-Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only some individuals do so.
-
-An extension of our model could incorporate more sophisticated oblique transmission schemes~\citep[][Figure~3]{Creanza2017}.
-For example, conformity---preference for learning a frequent phenotype---has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
-We suggest that the specific mode of oblique transmission can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions.
-Additional work will be required to understand how such fluctuations affect the population dynamics.
-
-% Facebook rural internet 
-% National service - Israel, Singapore
-% Boarding schools
-% Immigrtion - draught
-% Opening schools, universities
-% sending students abroad, bringing experts from abroad
-% foreign labor
-% refugees
-% hunter-gatherers/chimps increase habitat in years with less food production
-% religion - dark ages etc.
-
-% Acknowledgements
-{\small
-\section*{Acknowledgements}
-
-This work was supported in part by 
-the Stanford Center for Computational, Evolutionary and Human Genomics, 
-the Morrison Institute for Population and Resources Studies, Stanford University, and the John Templeton Foundation.
-}
-
-\bibliographystyle{agsm}
-%\bibliography{/Users/yoavram/Documents/library}
-\bibliography{ms_sunnyvale}
-
 \end{document}

### 2018-07-05 ms/ms_sunnyvale.tex
small corrections; fixed phrasing at end of appendix

@@
 % Title page
-\title{Vertical and oblique {\color{red}cultural} transmission fluctuating in time and in space}
+\title{Vertical and oblique cultural transmission fluctuating in time and in space}
 
@@
 For cultural evolution, the mode of transmission between individuals of the same or different generations is also central to the dynamics of cultural traits.
-%Important modes of cultural transmission are vertical (from parent to offspring), horizontal (between peers), and oblique (from adults in the parental generation to offspring) transmission. 
-{\color{red}Using analysis with epidemic theory}, \citet[ch.~3]{Cavalli-Sforza1981}  introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, obliquely, that is, from non-parental members of the parent's generation to an offspring{\color{red}, or horizontally, that is, among members of the same generational cohort.}
-Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time.
+Using analysis with epidemic theory, \citet[ch.~3]{Cavalli-Sforza1981}  introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, obliquely, that is, from non-parental members of the parent's generation to an offspring, or horizontally, that is, among members of the same generational cohort.
+Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time {\color{red}or space}.
 For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-$k$, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
@@
 
+% figure
 \begin{figure*}[h]
@@
 \end{figure*}
-
-% TODO what kind of spatial differences can we imagine in transmission?
-% two populations, one has first cousin marriages and one doesn't (endogmia)
-% orthodox religious communities
-% boarding schools for one population
-% foraging in groups or families
-% pigmies women 'migration' - adoption by husband family
-% interesting: migration correlated with transmission rate, future work
-
-% matrilineal primates - vertical transmission stronger in females than males
-% similar differences in ultra religious human societies
 
@@
 
+% Symmetric periods
 \paragraph{Symmetric periods---$AkBk$.}
@@
 
+% Period of one
 \paragraph{Period of one---$A1B1$.}
@@
 
-%\begin{align}
-%G(0) = c > 0  \Leftrightarrow \rho_B > \frac{\rho_A(1+s)}{1+\rho_As}, \\
-%G'(0) = b > 0 \Leftrightarrow \rho_B > \frac{\rho_A s}{2 + \rho_As}, \\
-%G(1) = a + b + c = (1+s)[\rho_B - \rho_A + s \rho_B (1-\rho_A)], \\
-%G'(1) = 2a + b = s(1-\rho_A)(2(1+s)\rho_B + s\rho_A(1-\rho_B)) > 0.
-%\end{align}
-
+% figure
 \begin{figure*}[htb]
@@
 The stable equilibrium of the frequency of phenotype $A$ (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
-Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from {\color{red}inequalities~(\ref{eq:poly_condition_periodic_fluc})} that permit a protected polymorphism.
+Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from inequalities~\eqref{eq:poly_condition_periodic_fluc} that permit a protected polymorphism.
 Here, $\rho_A=0.5$.}
@@
 
+% Asymmetric period
 \paragraph{Asymmetric period---$AkBl$.}
@@
 Let $w_{A,t}, w_{B,t} \sim U(0,1)$, $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$ and $\rho_t \sim Beta(1+s_t, 1)$ (a beta distribution with parameters $1+s_t$ and $1$). 
-The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$ {\color{red}as} estimated by averaging over $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
+The covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 4$ as estimated by averaging over $10^8$ random values of $s_t$ and $\rho_t$); that is, vertical transmission is more likely when $A$ is favored (i.e.
 $s_t>0$) and oblique transmission is more  likely when $B$ is favored (i.e.
@@
 
+% figure
 \begin{figure*}[hbt]
@@
 where $S_n = \sum_{t=1}^{n}{z_t}$ and $z_t = \rho_t s_t$.
-Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability {\color{red}of} fixation of phenotype $A$ when its initial frequency if $x$, is
+Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability of fixation of phenotype $A$ when its initial frequency if $x$, is
 \begin{equation}
@@
 
-Following migration~\citep{Karlin1981}, the frequencies of $A$ in the two demes are
+Following migration, the frequencies of $A$ in the two demes are
 \begin{equation} \label{eq:migration_model_migration}
@@
 
+% Asymmetric migration and symmetric selection
 \paragraph{Asymmetric migration and symmetric selection.}
 
-\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection symmetric selection $w_1=1/w_2=w>1$.
+\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection{\color{red},} $w_1=1/w_2=w>1$.
 We focus on composite parameters of the model:
 $m_1/m_2$, on the x-axis, is the ratio of migration rates into deme 1 and deme 2; when this ratio is large, deme 1 accepts more migrants than deme 2.
-$\rho_1/\rho_2$, on the y-axis the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.
+$\rho_1/\rho_2$, on the y-axis{\color{red},} the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.
 
 The results demonstrate that fixation of phenotype $A$ is stable if migration to and oblique transmission within deme 1, where $A$ is favored, are higher than in deme 2.
-The {\color{red}opposite} is true for phenotype $B$.
+The opposite is true for phenotype $B$.
 A protected polymorphism exists if neither fixation is stable: if migration ratios are positively correlated -- vertical transmission occurs more often in the deme that accepts more migrants -- or if both ratios are close to unity, that is, the differences between the demes in terms of migration and transmission are small (\autoref{fig:asym_migration}).
 
+% figure
 \begin{figure*}[hbt]
@@
 
+% Symmetric migration and selection
 \paragraph{Symmetric migration and selection.}
@@
 \begin{itemize}
-\item With only oblique transmission ($\rho_1=\rho_2=0$), {\color{red}there are only} neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
+\item With only oblique transmission ($\rho_1=\rho_2=0$), there are only neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
 \item With only vertical transmission ($\rho_1=\rho_2=1$), the fixation equilibria  $(0,0)$, $(1,1)$ are unstable and there exists a protected polymorphism
@@
 
+% figure
 \begin{figure*}[ht]
@@
 \textbf{Oblique transmission and migration: effect of transmission.} 
-The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\bar{w}^*_i$ the population mean fitnesses (bottom row) in the two demes.
+The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\overline{w}^*_i$ the {\color{red}stable} population mean fitnesses (bottom row) in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
@@
 Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
-The shaded area denotes stable fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
-Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
+The shaded area denotes stable fixation of phenotype $A$ according to {\color{red}inequality}~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
+Lines are drawn by iterating eqs.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
 }
@@
 \textbf{Oblique transmission and migration: effect of migration.} 
-The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\bar{w}^*_i$, the population mean fitnesses (bottom row), in the two demes.
+The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\overline{w}^*_i$, the {\color{red}stable} population mean fitnesses (bottom row), in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
@@
 Migration rate $m$ is on the x-axis.
-The shaded area denotes stables fixation of phenotype $A$ according to eq.~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
+The shaded area denotes stables fixation of phenotype $A$ according to {\color{red}inequality}~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
 Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
@@
 Comparing our results with previous results from a similar model with a fixed rate of vertical transmission~\citep{Ram2018}, we find that a protected polymorphism can be maintained only if fluctuations in the rate of vertical transmission are small, and that stronger selection on the transmitted trait permits larger fluctuations in the rate of transmission while still maintaining a protected polymorphism.
-In the case of fluctuations in space, the greater the separation between the two populations {\color{red}(i.e., the smaller the migration rates)} the larger {\color{red} can be the fluctuations that maintain} polymorphism; however, as migration becomes more frequent, even small differences in the vertical transmission rate will eliminate the polymorphism.
+In the case of fluctuations in space, the greater the separation between the two populations (i.e., the smaller the migration rates) the larger can be the fluctuations that maintain polymorphism; however, as migration becomes more frequent, even small differences in the vertical transmission rate will eliminate the polymorphism.
 When fluctuations are stochastic, we find that if vertical transmission covaries with selection, the phenotype that has a higher probability of being vertically transmitted  when it is favored will likely eventually fix in the population.
@@
 One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
-Although there is {\color{red}some} evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
+Although there is some evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
 Indeed, it has been demonstrated that guppy fish are more likely to learn from others if previous social experiences provided benefits~\citep{Leris2016}.
@@
 For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
-Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only some individuals do so.
+Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only {\color{red}specific (e.g. stressed)} individuals do so.
 
@@
 Additional work will be required to understand how such fluctuations affect the population dynamics.
-
-% Facebook rural internet 
-% National service - Israel, Singapore
-% Boarding schools
-% Immigrtion - draught
-% Opening schools, universities
-% sending students abroad, bringing experts from abroad
-% foreign labor
-% refugees
-% hunter-gatherers/chimps increase habitat in years with less food production
-% religion - dark ages etc.
 
@@
 
+% bibliography
 \bibliographystyle{agsm}
-%\bibliography{/Users/yoavram/Documents/library}
 \bibliography{ms_sunnyvale}
 
+% Appendix
 \section*{Appendix I}
@@
 \begin{equation}\begin{aligned}
-\Delta(m=0) = \Big(\frac{(w + \rho) (w-1)}{w}\Big)^2, \text{and} \\
-\Delta(m=1/2) = \Big(\frac{w^2 + w(1-\rho) + \rho}{2 w}\Big)^2,
+&\Delta(m=0) = \Big(\frac{(w + \rho) (w-1)}{w}\Big)^2, \quad \text{and} \\
+&\Delta(m=1/2) = \Big(\frac{w^2 + w(1-\rho) + \rho}{2 w}\Big)^2,
 \end{aligned}\end{equation}
@@
 For $\rho=0$ we have $a+b+c=-m(w-1)<0$.
-Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any $0 \le \rho \le 1$, so $f(x)$ has a real root greater than one, and $(0,0)$ is unstable.
+Finally, $a+b+c$ is a linear function of $\rho$ and therefore $a+b+c<0$ for any $0 \le \rho \le 1$, so $f_0(x)$ has a real root greater than one, and $(0,0)$ is unstable.
 
@@
 \begin{equation}\begin{aligned}
-\Delta(m=0) = \Big(\frac{(w \rho + 1) (w-1)}{w}\Big)^2, \text{and} \\
-\Delta(m=1/2) = \Big(\frac{w^2 \rho + w(1-\rho) + 1}{2 w}\Big)^2,
+&\Delta(m=0) = \Big(\frac{(w \rho + 1) (w-1)}{w}\Big)^2, \quad \text{and} \\
+&\Delta(m=1/2) = \Big(\frac{w^2 \rho + w(1-\rho) + 1}{2 w}\Big)^2,
 \end{aligned}\end{equation}
@@
 Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and  a positive derivative at infinity.
-Therefore, fixation of $A$ is stable iff both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
-Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}$, and
-$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}$, which is larger than $\frac{m}{m+(1-m)(w-1)}$ for any $w>1$ and $0 \le m \le 1/2$ (the difference{[\color{red}WHAT DIFFERENCE?]} is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $w=2$, $m=0$).
-Therefore, fixation of $A$ is stable if $\rho < \frac{m}{m+(1-m)(w-1)}$ \eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
-By rearranging this inequality, we can also obtain an expression for $m$ \eqref{eq:migration_model_unconditional_symmetric_condition_m}.
+Therefore, fixation of $A$ is stable {\color{red}if and only if} both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
+{\color{red}Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}=M_1$, and
+$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}=M_2$.
+The difference $M_1-M_2$ is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $(w=2, m=0)$, so $M_1<M_2$ for any $w>1$ and $0 \le m \le 1/2$.}
+
+Therefore, fixation of $A$ is stable if {\color{red}$\rho < M_1 = \frac{m}{m+(1-m)(w-1)}$}~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
+By rearranging this inequality, we can also obtain an expression for $m$~\eqref{eq:migration_model_unconditional_symmetric_condition_m}.
 

### 2018-07-13 ms/ms_sunnyvale.tex
first submission to TPB

@@
 Using analysis with epidemic theory, \citet[ch.~3]{Cavalli-Sforza1981}  introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, obliquely, that is, from non-parental members of the parent's generation to an offspring, or horizontally, that is, among members of the same generational cohort.
-Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time {\color{red}or space}.
+Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time or space.
 For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-$k$, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
@@
 When both the rate of vertical transmission and selection fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that maintains a polymorphism.
+Versions of the model with random fitnesses and transmission rates, and with random drift due to finite population size are also analyzed.
 We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two populations.
@@
 
-Suppose that both transmission and selection fluctuate together, so that when $A$ is favored, with $w_A=1+s$ and $w_B=1$, the transmission rate is $\rho_A$, and when $B$ is favored, with $w_A=1$ and $w_B=1+s$, the transmission rate is $\rho_B$.
+Suppose that both transmission and selection fluctuate together, so that when $A$ is favored, with $w_A=1+s$ and $w_B=1$ with $s>0$, the transmission rate is $\rho_A$, and when $B$ is favored, with $w_A=1$ and $w_B=1+s$, the transmission rate is $\rho_B$.
 The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ or $F_B(x)$, respectively, where
@@
 
-{\bf Result 1.} {\sl Suppose that vertical transmission occurs at rate $\rho_A$ when the fitness of phenotype $A$ is $1+s$ relative to $1$ for phenotype $B$, and at rate $\rho_B$ when the fitness of phenotype $B$ is $1+s$ relative to $1$ for phenotype $A$.
+{\bf Result 1.} {\sl Suppose that vertical transmission occurs at rate $\rho_A$ when the fitness of phenotype $A$ is $1+s$ relative to $1$ for phenotype $B$, and at rate $\rho_B$ when the fitness of phenotype $B$ is $1+s$ relative to $1$ for phenotype $A$, where $s>0$.
 Then the stronger the selection (i.e. the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.}
@@
 \textbf{Protected polymorphism.}
-The stable equilibrium of the frequency of phenotype $A$ (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
+The frequency of phenotype $A$ at the stable equilibrium   (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
 Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from inequalities~\eqref{eq:poly_condition_periodic_fluc} that permit a protected polymorphism.
@@
 where $z_t=\rho_t s_t$; $s_t$ are i.i.d (independent and identically distributed); $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$; $\rho_t$ are i.i.d; and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
-Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$.
+Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$.  Following \citet{Karlin1974}, we have:
 
@@
 
-In our case there are two constant equilibria $x^*=0$ and $x^*=1$ corresponding to fixation in $B$ and $A$, respectively.
+In our case there are two constant equilibria $x^*=0$ and $x^*=1$ corresponding to fixation in $B$ and in $A$, respectively.
 From \citet[][results~6 and 7]{Ram2018} we can characterize the stochastic local stability of these fixations as follows.
@@
 Therefore, if we take the i.i.d fitness random variables for $A$ and $B$ to be $w_{A,t}, w_{B,t} \sim U(0,1)$, respectively, and define $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$, then neither $A$ nor $B$ has an advantage, on average (i.e.
-$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to approach a polymorphic distribution.
+$E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}a), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to approach a polymorphic distribution.
 
@@
 \centering
-\includegraphics[width=\linewidth]{../figures/{beta}.png}
+\includegraphics[width=\linewidth]{../figures/{beta}.pdf}
 \caption{
 \textbf{Covariance of selection and transmission.}
-\textbf{(A)} Histogram of $w_{A,t}/w_{B,t}$ where $w_{A,t}$ and $w_{B,t}$ are identically and independently distributed uniform random variables $U(0,1)$.
-\textbf{(B)} Histogram of $s_t = (w_{A,t}-w_{B,t})/w_{B,t}$.
-\textbf{(C)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
-\textbf{(D)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
+\textbf{(a)} Histogram of $w_{A,t}/w_{B,t}$ where $w_{A,t}$ and $w_{B,t}$ are identically and independently distributed uniform random variables $U(0,1)$.
+\textbf{(b)} Histogram of $s_t = (w_{A,t}-w_{B,t})/w_{B,t}$.
+\textbf{(c)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
+\textbf{(d)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
 }
@@
 
-\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection{\color{red},} $w_1=1/w_2=w>1$.
+\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection, $w_1=1/w_2=w>1$.
 We focus on composite parameters of the model:
 $m_1/m_2$, on the x-axis, is the ratio of migration rates into deme 1 and deme 2; when this ratio is large, deme 1 accepts more migrants than deme 2.
-$\rho_1/\rho_2$, on the y-axis{\color{red},} the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.
+$\rho_1/\rho_2$ on the y-axis is the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.
 
@@
 \end{itemize}
-{\color{red}The proofs of (22) and (23) are in Appendix I.}
+The proofs of (22) and (23) are in Appendix I.
 \paragraph{Examples.}
@@
 \textbf{Oblique transmission and migration: effect of transmission.} 
-The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\overline{w}^*_i$ the {\color{red}stable} population mean fitnesses (bottom row) in the two demes.
+The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\overline{w}^*_i$ the stable population mean fitnesses (bottom row) in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
@@
 Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
-The shaded area denotes stable fixation of phenotype $A$ according to {\color{red}inequality}~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
+The shaded area denotes stable fixation of phenotype $A$ according to inequality~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
 Lines are drawn by iterating eqs.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
@@
 \textbf{Oblique transmission and migration: effect of migration.} 
-The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\overline{w}^*_i$, the {\color{red}stable} population mean fitnesses (bottom row), in the two demes.
+The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\overline{w}^*_i$, the stable population mean fitnesses (bottom row), in the two demes.
 Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
@@
 Migration rate $m$ is on the x-axis.
-The shaded area denotes stables fixation of phenotype $A$ according to {\color{red}inequality}~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
+The shaded area denotes stables fixation of phenotype $A$ according to inequality~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
 Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
@@
 For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
-Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only {\color{red}specific (e.g. stressed)} individuals do so.
+Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only specific (e.g. stressed) individuals do so.
 
@@
 Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and  a positive derivative at infinity.
-Therefore, fixation of $A$ is stable {\color{red}if and only if} both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
-{\color{red}Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}=M_1$, and
+Therefore, fixation of $A$ is stable if and only if both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
+Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}=M_1$, and
 $f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}=M_2$.
-The difference $M_1-M_2$ is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $(w=2, m=0)$, so $M_1<M_2$ for any $w>1$ and $0 \le m \le 1/2$.}
+The difference $M_1-M_2$ is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $(w=2, m=0)$, so $M_1<M_2$ for any $w>1$ and $0 \le m \le 1/2$.
 
-Therefore, fixation of $A$ is stable if {\color{red}$\rho < M_1 = \frac{m}{m+(1-m)(w-1)}$}~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
+Therefore, fixation of $A$ is stable if $\rho < M_1 = \frac{m}{m+(1-m)(w-1)}$~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}.
 By rearranging this inequality, we can also obtain an expression for $m$~\eqref{eq:migration_model_unconditional_symmetric_condition_m}.

### 2018-10-29 ms/ms_sunnyvale.tex
first draft of response

@@
 
-Here, we analyze a model~(\autoref{fig:transmission}) in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space.
-We find that fluctuations in $\rho$ without corresponding fluctuations in selection cannot maintain a polymorphism.
-When both the rate of vertical transmission and selection fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that maintains a polymorphism.
+We previously analyzed a model~\citep{Ram2018} in which individuals learn a cultural trait from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission; see \autoref{fig:transmission}).
+Here, we generalize this model to allow for the vertical transmission rate $\rho$ to fluctuate over time or space.
+We find that fluctuations in $\rho$ without corresponding fluctuations in selection cannot maintain trait polymorphism.
+When both the rate of vertical transmission and selection fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that maintains trait polymorphism.
 Versions of the model with random fitnesses and transmission rates, and with random drift due to finite population size are also analyzed.
 We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two populations.
-Stability of fixation points  and a polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.
+Stability of fixation points  and trait polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.
 
-We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either selection or transmission or both, and that further empirical and theoretical studies are required to disentangle these effects.
+We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either selection or transmission or both, and that such variation has a significant effect of trait polymorphism.
+%Further empirical and theoretical studies are required to disentangle these effects.
 
@@
 
-Consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ and fitness values $w_A=1+s$ and $w_B=1$, respectively.
-Phenotypes are transmitted vertically with probability $\rho$ or obliquely with probability $1-\rho$~(\autoref{fig:transmission}).
-Given $x$, the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is
-\begin{equation} \label{eq:recurrence}
-x' = \rho \frac{1+s}{\overline w} x + (1-\rho)x,
-\end{equation}
+Following \citet{Ram2018}, we consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ that determine the reproductive success of adults, with fitness values $w_A=1+s$ and $w_B=1$, respectively.
+Phenotypes are transmitted from adults to juveniles: vertically from parent to offspring with probability $\rho$, or obliquely from random adults to offspring with probability $1-\rho$~(\autoref{fig:transmission}).
+Given~$x$, the frequency of phenotype $A$ at the current generation, the frequency of $A$ in the next generation is (see eq.~2 in~\citet{Ram2018})
+\begin{equation}\begin{aligned} \label{eq:recurrence}
+x' &= \rho \frac{1+s}{\overline w} x + (1-\rho)x \\
+&= \frac{1+s}{\overline w} x \big[(1-\rho)x + \rho\big] + \frac{1}{\overline w} (1-x) [(1-\rho)x],
+\end{aligned}\end{equation}
 where $\overline w = 1 + xs$ is the population mean fitness.
@@
 where $z_t=\rho_t s_t$; $s_t$ are i.i.d (independent and identically distributed); $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$; $\rho_t$ are i.i.d; and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
-Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$.  Following \citet{Karlin1974}, we have:
+Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)=1$.  Following \citet{Karlin1974}, we have:
 
@@
 
+An interesting model with a similar structure was studied by \citet{Bisin2001}: it included both vertical (i.e.~``socialization in the family'') and non-vertical (i.e.~``imitation from society'') transmission.
+They focused on a frequency-dependent transmission mode called \emph{cultural substitution}, in which a parent invests more in child socialization if the parental trait is rare.
+Their model (eq.
+12) is a continuous-time equivalent of our model (eq.~\ref{eq:recurrence}) that does not distinguish between horizontal and oblique transmission; considers constant, rather than fluctuating, selection; and includes endogenously fluctuating transmission due to changes in trait frequency.
+They have found that under some selection schemes, in which parents choose how many offspring to produce, either an unstable polymorphism (Proposition 4) or a stable polymorphism (Proposition 5) exists, depending on the costs associated with the transmission mode.
+Furthermore, in the absence of selection trait polymorphism is neutral in our model (set $s=0$ in eq.~\ref{eq:recurrence}), but \citet{Bisin2001} have found that frequency-dependent transmission leads to a globally stable polymorphism~(Proposition 1).
+
+An important topic in cultural evolution is the evolutionary dilemma known as Rogers' paradox~\citep{Rogers1988}, which is similar to the problem of the maintenance of cooperation: social learners (aka \emph{cheaters}) benefit from new cultural traits invented by individual learners (aka \emph{cooperators}) without paying the cost associated with individual learning.
+Several authors used population structure to help explain the maintenance of individual learning in this scenario, paralleling the use of population structure for explaining the maintenance of cooperation.
+For example, \citet{Kobayashi2012} modeled both individual learning and social learning via oblique transmission in an infinite island model.
+Their results show that spatial subdivision favors individual learning over social learning.
+Maintenance of individual learning is also crucial for cumulative culture, in which cultural innovations made by individual learning are transmitted and accumulated via social learning.
+Similarly to the effect of cheating in public goods games, social learning can cause the collapse of cumulative culture~\citep{Wakano2014}.
+To resolve this dilemma, \citet{Ohtsuki2017} modeled individual and social learning in a structured population, but allowed social learning to be at least partially vertical, rather than completely horizontal or oblique.
+They have demonstrated that if kin-selection has a significant effect on the dynamics, then increases in inclusive fitness due to vertical transmission can negate the cost of individual learning, at least to some extent, and allow for high levels of cumulative culture to evolve.
+\citet{Mullon2017} have also studied a model of individual and social learning that includes both vertical and oblique transmission in a cumulative culture setting. 
+Their results show that the effect of vertical transmission on maintenance of individual learning is stronger if vertical transmission is more efficient than oblique transmission, i.e.
+``if offspring learn better from relatives, or if parents devote more teaching effort towards offspring''.
+Following our hypothesis that the transmission mode fluctuates over time and space, we suggest that such fluctuations could facilitate the maintenance of individual learning, and therefore cultural accumulation: when vertical transmission is high, innovation by individual learning will be common due to increased inclusive fitness; when oblique transmission is high, new cultural traits will quickly spread via social learning.
+
 One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
@@
 
+Our spatial model (eqs.
+\ref{eq:migration_model_selection_transmission}-\ref{eq:migration_model_migration}) assumes migration by sub-adults, or more specifically, that migration follows cultural transmission, which occurs at the native deme.
+This is a necessary assumption for modeling vertical transmission, which cannot occur after offspring migrate out of their native deme, leaving their parents behind.
+Other authors have taken a different approach when focusing on oblique transmission: for example, \citet{Kobayashi2012} assumed migration by juveniles, followed by a combination of individual and social learning at the new deme.
+Further work could investigate the evolution of vertical vs. horizontal transmission, in which case migration could occur before transmission, rather than after.
+
 An extension of our model could incorporate more sophisticated oblique transmission schemes~\citep[][Figure~3]{Creanza2017}.
 For example, conformity---preference for learning a frequent phenotype---has been demonstrated in wild monkeys~\citep{VanDeWaal2013} and birds~\citep{Aplin2015}.
-We suggest that the specific mode of oblique transmission can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions.
+We suggest that the specific mode of oblique transmission can also fluctuate over time, so that individuals can, for example, conform to the frequent phenotype under benign conditions, and prefer a rare phenotype under stressful conditions; more generally, the mode of transmission could be plastic and viewed as a reaction norm to variable environments.
 Additional work will be required to understand how such fluctuations affect the population dynamics.
@@
 \bibliographystyle{agsm}
-\bibliography{ms_sunnyvale}
+\bibliography{/Users/yoavram/Documents/library}
+%\bibliography{ms_sunnyvale}
 

### 2018-10-30 ms/ms_sunnyvale.tex
new par by MWF

@@
 
+The interaction between spatially varying selection and transmission is shown in Figures \ref{fig:migration_rho} and \ref{fig:migration_m} to depend on migration in a complex way.
+However, our two-deme model does not exhibit the kind of complexity seen in \citet{Ram2018}, where the period of environmental fluctuation determined whether the vertical transmission rate $\rho$ that maximized the population mean fitness was the same as the evolutionarily stable value.
+We can speculate that a model with more sub-populations and a more geographically complicated fitness regime might show less predictable relationships between transmission and selection.
+
 An interesting model with a similar structure was studied by \citet{Bisin2001}: it included both vertical (i.e.~``socialization in the family'') and non-vertical (i.e.~``imitation from society'') transmission.

### 2018-10-31 ms/ms_sunnyvale.tex
MWF comments

@@
 % line numbers
-\usepackage[displaymath, mathlines]{lineno}
-\renewcommand\linenumberfont{\normalfont\small\sffamily}
-\linenumbers
-\modulolinenumbers[2]
+%\usepackage[displaymath, mathlines]{lineno}
+%\renewcommand\linenumberfont{\normalfont\small\sffamily}
+%\linenumbers
+%\modulolinenumbers[2]
 
@@
 Nevertheless, if we assume that social learning can be affected by ecological and demographic factors, such as the frequency of interactions~\citep{VanSchaik2003}, weather~\citep{Phithakkitnukoon2012}, population size or density~\citep{Fischer2015,Aureli1997a}, or stress~\citep{Farine2015}, then it is reasonable that the mode and rate of transmission may fluctuate over time and/or space.
-For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish), increase their reliance on social learning when predation risk increases.
+For example, \citet{Webster2008} have demonstrated that minnows (small freshwater shoaling fish) increase their reliance on social learning when predation risk increases, whereas \citet{Hewlett2011} found that the balance between vertical and non-vertical cultural transmission depends on core cultural values and therefore differs between hunter-gatherers and farmers.
 
@@
 
-We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either selection or transmission or both, and that such variation has a significant effect of trait polymorphism.
+We suggest that cultural evolution can be affected by environmental changes that cause temporal or spatial variation in either selection or transmission or both, and that such variation has a significant effect on trait polymorphism.
 %Further empirical and theoretical studies are required to disentangle these effects.
@@
 Following \citet{Ram2018}, we consider a very large population whose members are characterized by a single dichotomous cultural trait with phenotypes $A$ and $B$ that determine the reproductive success of adults, with fitness values $w_A=1+s$ and $w_B=1$, respectively.
+Example traits include foraging and hunting techniques, such as lobtail feeding in humpback whales~\citep{Allen2013} and fruit processing in capuchin monkeys~\citep{Barrett2017}, child caring, mating skills, and dancing in human hunter-gatherers~\citep{Hewlett1986}, and even canoe design in Polynesia~\citep{Rogers2008}.
 Phenotypes are transmitted from adults to juveniles: vertically from parent to offspring with probability $\rho$, or obliquely from random adults to offspring with probability $1-\rho$~(\autoref{fig:transmission}).
@@
 12) is a continuous-time equivalent of our model (eq.~\ref{eq:recurrence}) that does not distinguish between horizontal and oblique transmission; considers constant, rather than fluctuating, selection; and includes endogenously fluctuating transmission due to changes in trait frequency.
-They have found that under some selection schemes, in which parents choose how many offspring to produce, either an unstable polymorphism (Proposition 4) or a stable polymorphism (Proposition 5) exists, depending on the costs associated with the transmission mode.
-Furthermore, in the absence of selection trait polymorphism is neutral in our model (set $s=0$ in eq.~\ref{eq:recurrence}), but \citet{Bisin2001} have found that frequency-dependent transmission leads to a globally stable polymorphism~(Proposition 1).
+They found that under some selection schemes, in which parents choose how many offspring to produce, either an unstable polymorphism (Proposition 4) or a stable polymorphism (Proposition 5) exists, depending on the costs associated with the transmission mode.
+Furthermore, in the absence of selection trait polymorphism is neutral in our model (set $s=0$ in eq.~\ref{eq:recurrence}), but \citet{Bisin2001} found that frequency-dependent transmission leads to a globally stable polymorphism~(Proposition 1).
 
@@
 To resolve this dilemma, \citet{Ohtsuki2017} modeled individual and social learning in a structured population, but allowed social learning to be at least partially vertical, rather than completely horizontal or oblique.
-They have demonstrated that if kin-selection has a significant effect on the dynamics, then increases in inclusive fitness due to vertical transmission can negate the cost of individual learning, at least to some extent, and allow for high levels of cumulative culture to evolve.
+They demonstrated that if kin-selection has a significant effect on the dynamics, then increases in inclusive fitness due to vertical transmission can negate the cost of individual learning, at least to some extent, and allow for high levels of cumulative culture to evolve.
 \citet{Mullon2017} have also studied a model of individual and social learning that includes both vertical and oblique transmission in a cumulative culture setting. 
@@
 Our spatial model (eqs.
-\ref{eq:migration_model_selection_transmission}-\ref{eq:migration_model_migration}) assumes migration by sub-adults, or more specifically, that migration follows cultural transmission, which occurs at the native deme.
-This is a necessary assumption for modeling vertical transmission, which cannot occur after offspring migrate out of their native deme, leaving their parents behind.
+\ref{eq:migration_model_selection_transmission}-\ref{eq:migration_model_migration}) assumes migration by sub-adults, or more specifically, that migration follows cultural transmission, which occurs at the natal deme.
+This is a necessary assumption for modeling vertical transmission, which cannot occur after offspring migrate out of their natal deme, leaving their parents behind.
 Other authors have taken a different approach when focusing on oblique transmission: for example, \citet{Kobayashi2012} assumed migration by juveniles, followed by a combination of individual and social learning at the new deme.

### 2018-11-07 ms/ms_sunnyvale.tex
submitted revision to tpb

@@
 % line numbers
-%\usepackage[displaymath, mathlines]{lineno}
-%\renewcommand\linenumberfont{\normalfont\small\sffamily}
-%\linenumbers
-%\modulolinenumbers[2]
+\usepackage[displaymath, mathlines]{lineno}
+\renewcommand\linenumberfont{\normalfont\small\sffamily}
+\linenumbers
+\modulolinenumbers[2]
 
@@
 
-\author[a]{Yoav Ram}
+\author[ac*]{Yoav Ram}
 \author[b]{Uri Liberman}
-\author[a]{Marcus W. Feldman}
+\author[a\dagger]{Marcus W. Feldman}
 \affil[a]{Department of Biology, Stanford University, Stanford, CA}
 \affil[b]{School of Mathematical Sciences, Tel Aviv University, Israel}
+\affil[c]{The Interdisciplinary Center, Herzliya, Israel}
+\affil[*]{{\it Current address}}
+\affil[$\dagger$]{{\it Corresponding author}}
 
@@
 \end{aligned}\end{equation}
+If the environment fluctuates between favoring $A$ for $k$ generations and $B$ for $l$ generations constituting a period of $(k+l)$ generations, we describe this as $AkBl$.
 
@@
 % Period of one
-\paragraph{Period of one---$A1B1$.}
-In the case  $k=1$ we can find $x^*$ the stable frequency of $A$ at the protected polymorphism.
-Set $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
+\paragraph{The case $A1B1$.}
+In the case  $k=1$ we can find $x^*$ the frequency of $A$ at the protected polymorphism.
+Set $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma(x)$ with
 \begin{equation} \begin{aligned} \label{eq:xstar_periodic_fluc_k=1}
-&G(x) = ax^2+bx+c, \\
-&\gamma = -\frac{1 + sx}{s} \Big[ 1 + s + s^2 x (1-\rho)(1+x)\Big], \\
-&a = s^2 \rho_A (1-\rho_A) (1 - \rho_B + \rho_B/\rho_A)) > 0, \\
-&b = s(1-\rho_A)(2\rho_B - s\rho_A(1-\rho_B)), \\
-& c = \rho_B - \rho_A - s\rho_A(1-\rho_B).
+G(x) &= ax^2+bx+c, \\
+\gamma(x) &= -\frac{1 + sx}{s} \Big[ 1 + s + s^2 x (1-\rho)(1+x)\Big], \\
+a &= s^2 \rho_A (1-\rho_A) (1 - \rho_B + \rho_B/\rho_A)) > 0, \\
+b &= s(1-\rho_A)(2\rho_B - s\rho_A(1-\rho_B)), \\
+c &= \rho_B - \rho_A - s\rho_A(1-\rho_B).
 \end{aligned} \end{equation}
@@
 The condition \eqref{eq:poly_condition_periodic_fluc}, which guarantees that $0 < x^* < 1$, is equivalent to $c<0$, and therefore it also guarantees that $\sqrt{b^2-4ac} > b$.
-So, if $0<s<1$ (so that $b$ is guaranteed to be positive if LHS of eq.~\ref{eq:poly_condition_periodic_fluc} holds), then 
+So, if $0<s<1$ ($b$ is then guaranteed to be positive if LHS of eq.~\ref{eq:poly_condition_periodic_fluc} holds), then 
 \begin{equation}
@@
 \autoref{fig:rho1_rho2_phase_k=1} shows $x^*$ and highlights the area of the parameter space in which a protected polymorphism exists.
-The figure demonstrates that the stronger the selection (x-axis), the greater the fluctuations in $\rho$  (y-axis) that still allow a polymorphic population (area between the dashed lines).
+The figure demonstrates that the stronger the selection, the greater the fluctuations in $\rho$ that still allow a polymorphic population (area between the dashed lines).
 
@@
 % Asymmetric period
-\paragraph{Asymmetric period---$AkBl$.}
+\paragraph{General period---$AkBl$.}
 More generally, phenotype $A$ could be favored for $k$ generations and phenotype $B$ for $l$ generations, with the transmission rate following the same cycle with $\rho=\rho_A$ when $A$ is favored and $\rho=\rho_B$ when $B$ is favored.
@@
 
-Second, if $s_t$ and $\rho_t$ are not independent such that $cov(\rho_t, s_t) \ne 0$, a fixation can occur.
+Second, if $s_t$ and $\rho_t$ are not independent and $cov(\rho_t, s_t) \ne 0$, a fixation can occur.
 Let $w_{A,t}, w_{B,t} \sim U(0,1)$, $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$ and $\rho_t \sim Beta(1+s_t, 1)$ (a beta distribution with parameters $1+s_t$ and $1$). 
@@
 \centering
-\includegraphics[width=\linewidth]{../figures/{beta}.pdf}
+\includegraphics[width=\linewidth]{../figures/{beta}.png}
 \caption{
@@
 where $S_n = \sum_{t=1}^{n}{z_t}$ and $z_t = \rho_t s_t$.
-Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability of fixation of phenotype $A$ when its initial frequency if $x$, is
+Furthermore, combining eq.~\ref{eq:drift_diffusion_terms} with eqs.~46-47 from \citet{Ram2018}, we find that the probability of fixation of phenotype $A$ when its initial frequency is $x$, is
 \begin{equation}
@@
 learning, development) occur within the demes, followed by migration of sub-adults---individuals that have already acquired their phenotype but have yet to reproduce.
-The frequency of phenotype $A$ in deme $j$ is denoted by $x_j$, and therefore after selection and transmission the frequencies are 
+The frequency of phenotype $A$ in deme $j$ is denoted by $x_j$, and therefore after selection and transmission the frequencies $x_j^s$ are 
 \begin{equation} \label{eq:migration_model_selection_transmission}
@@
 \centering
-\includegraphics[width=0.75\linewidth]{../figures/asym_migration.pdf}
+\includegraphics[width=0.75\linewidth]{../figures/asym_migration.png}
 \caption{
@@
 They found that under some selection schemes, in which parents choose how many offspring to produce, either an unstable polymorphism (Proposition 4) or a stable polymorphism (Proposition 5) exists, depending on the costs associated with the transmission mode.
-Furthermore, in the absence of selection trait polymorphism is neutral in our model (set $s=0$ in eq.~\ref{eq:recurrence}), but \citet{Bisin2001} found that frequency-dependent transmission leads to a globally stable polymorphism~(Proposition 1).
+Furthermore, in the absence of selection, trait polymorphism is neutral in our model (set $s=0$ in eq.~\ref{eq:recurrence}), but \citet{Bisin2001} found that frequency-dependent transmission leads to a globally stable polymorphism~(Proposition 1).
 
@@
 \bibliographystyle{agsm}
-\bibliography{/Users/yoavram/Documents/library}
-%\bibliography{ms_sunnyvale}
+%\bibliography{/Users/yoavram/Documents/library}
+\bibliography{ms_sunnyvale}
 
